### 2010-09-08 JSS/article.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@13 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article]{jss}

\usepackage{amsthm, amsmath, amssymb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Taylor B. Arnold \\ Yale University \And 
        John W. Emerson\\ Yale University}
\title{Nonparametric Goodness-of-Fit Tests \\ for Discrete Null Distributions}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Taylor B. Arnold, John W. Emerson} %% comma-separated
\Plaintitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% without formatting
\Shorttitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
   to users. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this need in the \proglang{R} language for
   the two most popular nonparametric tests. The structure of these two packages are explained and examples
   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
   in their implementation.
}
\Keywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, \proglang{R}}
\Plainkeywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, R} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Taylor B. Arnold \\
  24 Hillhouse Ave. \\
  New Haven, CT 06511, USA\\
  E-mail: \email{taylor.arnold@yale.edu}\\
  URL: \url{http://euler.stat.yale.edu/~tba3}\\

  John W. Emerson \\
  24 Hillhouse Ave. \\
  New Haven, CT 06511, USA\\
  E-mail: \email{john.emerson@yale.edu}\\
  URL: \url{http://euler.stat.yale.edu/~jay}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section{Introduction}

%General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
Given an observed sequence of data, the typical question of statistical inference is to determine
the underlying process from which it came. Given a particular null distribution, goodness-of-fit tests 
are used in order to test whether it is likely that the data were generated via the null distribution.

%Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
While almost any hypothesis test can be viewed as a variant on a goodness-of-fit test, the term is typically
applied to those tests which are nonparametric in nature. That is, they do not conduct the hypothesis testing 
through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
calculate differences between the empirical distribution of the observed data and the overall distribution 
of the null model. In many cases these tests are preferential since they tend to have an increased power for 
interesting deviations from the null model, whilst allowing for more robustness due to uninteresting deviations
such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and Smirnov,
and several variants on the estimated proposed by Cramer von Mises.

% Discrete versions exist; just no code
While the original aim of nonparametric goodness of fit tests was more continuous null distributions, discrete versions
have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
apply the functions meant for continuous distributions. As shown in section 5, the latter can be particularly dangerous
in the small sample setting. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this void for users
operating in the \proglang{R} environment.

\section{Kolmogorov-Smirnov Test}

% Most popular choice of test; explain the basic algorithm and purpose for continuous distributions (maybe a few P's).

\subsection{Overview}

Of all the methods for nonparametric goodness-of-fit tests, by far the most popular is the method devised by Kolmogorov and Smirnov.
It is the only such test built into the base of \proglang{R}. The idea behind it is fairly simple. Given the cumulative distribution
function $F_0(x)$ of the continuous null distribution, and the empirical distribution function $F_{data}(x)$ of the observed data, 
one constructs the statistic:
\begin{align}
D = \sup_x \left| F_0(x)- F_{data}(x) \right|
\end{align}
The difficult part of the theory is showing that under the null model $D$ has a fixed distribution. Perhaps most interestingly is the fact 
that this distribution does not depend on which null model was specified. Given this, it is easy to read p-values for the test statistic 
off of a table or computer function.

The extension of this result to non-continuous null distributions does not have such a clean solution. While the calculation of
the test statistic $D$ is still just as easy, the distribution of the testing statistic is much more difficult. Unlike in the 
continuous case, it ultimately depends on which null model was chosen; this makes it impossible to simply read p-values directly 
off of a fixed table. It was known since at least the 1950's that using the tables for continuous distributions resulted in 
conservative p-values; it was not until \citet{Conover1972} that a method for computing exact p-values in this case was developed.

Two common alternatives of the above test statistic exit, and are also encoded in \proglang{R}. The absolute value is discarded and the rest is either
left alone (the `greater' testing alternative) or the supremum is replaces with a negative infimum (the `lesser' hypothesis alternative). These can
be helpful depending on the nature of the alternative hypotheses for which the test is desired to be powerful towards.

\subsection{Implementation}

The implementation of the discrete Kolmogorov-Smirnov function consists of two parts. First the particular test statistic needs to be calculated,
and then the p-value for that particular statistic must be computed. These are the same general procedures as for the default function 
\pkg{ks.test} in \proglang{R}. Therefore, the basic implementation here attempts to incorporate the discrete test into the default function. There is
a default built-in class for step-functions and therefore it is easy to write code which checks for the input null distribution being a step-function
and to divert these cases to the implementation below. Other null distributions are handled as they were previously.

Given that the test statistic is, theoretically, the same as in the
continuous case it would seem that the first part could be directly taken from the existing code. This is unfortunately not the case. Given the
quantity $x$, which is a vector consisting of the difference between the cumulative null distribution and cumulative empirical distribution at each
observed data point, the default in \proglang{R} to determine the Kolmogorov-Smirnov test statistic is (with modifications made for readability):
\begin{CodeChunk}
\begin{CodeInput}
  STATISTIC <- max(c(abs(x), abs(x-1/n)))
\end{CodeInput}
\end{CodeChunk}
This method utilizes the fact that the supremum in equation 1 is the absolute difference between a continuous function and stepwise constant
function; this supremum can be reached either at an observed data point, or for a sequence of points approaching an observed data point from the 
left. This particular fact is still true if both functions are stepwise constant, but the problem comes from the fact that this code assumes, with the code
\code{abs(x-1/n)}, that the cumulative null distribution is continuous at the observed data points. This should never be the case for discrete null
distributions. This gives the incorrect default behavior:
\begin{CodeChunk}
\begin{CodeInput}
  R>  ks.test(c(1,2), ecdf(c(1,2))) 
\end{CodeInput}
\begin{CodeOutput}
	One-sample Kolmogorov-Smirnov test

  data:  c(1, 2) 
  D = 0.5, p-value = 0.5
  alternative hypothesis: two-sided 
\end{CodeOutput}
\end{CodeChunk}
Where clearly the test statistic should have been $0$, since the null and empirical cumulative distributions are the same.

Getting around this
issue is easy to do by computing the test statistic as \code{max(abs(x))} in the discrete case. The important point here is that \proglang{R}'s
default behavior for computing the test statistic was dependent on the continuity of the null distribution, even though in theory the discreteness
should not have affected the calculations.

Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
continuous case. The default options for \pkg{ks.test} already allow the user to specify if they would like approximate asymptotic distributions (since these exist
in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the \proglang{R} language, or in any other open source options, did not previously
exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
the original Conover paper.

%Numerical issues arise from the summation; need to do something else for high $n$. Simulation? Asymptotic approximation? 
The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.

\section{Cramer von Mises Tests}

\subsection{Algorithm}

%General idea for continuous distributions. In many cases actually more powerful of the two. There again exists a distribution free 
%distribution for the testing statistic.
While the Kolmogorov-Smirnov test is the most well known of the non-parametric goodness of fit tests, there is another family of 
tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
Harald Cramer and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted in \cite{anderson1952}, and 
\cite{Watson1961}. The test statistics are, respectively, given as:
\begin{align*}
W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
A^2 &= n \cdot \int_{-\infty}^{\infty} \left[F_0(x) -F_0(x)^2 \right]^{-1} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
\end{align*}
Where $F$ is either the cumulative distribution of the null model or the empirical cumulative distribution of the observed data.
As in the Kolmogorov-Smirnov test statistic, these all have distribution free null distributions in the continuous case.

It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations. As they all involve
integration over the whole range of data, rather than one supremum, it is not surprising that they are generally best when the
true alternative distribution deviates a little over the whole range of data rather than deviating a lot over a small range. For
a complete analysis of the relative powers of these tests see \cite{stephens1974}.

%Discuss the discrete version. How to calculate W2, U2, and A2. 
Generalizations of the Cramer-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
the original papers shows that this approximation

\subsection{Implementation}

For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
should also originally implement the continuous case even though this is not what we concentrate on here.

Calculation of the three test statistics is done by straightforward matrix algebra as given in \cite{choulakian1994}. Determining 
the form of the asymptotic null distribution is also easy using the built-in eigenvalue decomposition functions. The only 
difficulty in the process involves actually calculating the percentiles for these weighted chi-squares.

%Show how to compute the distribution of a weighted sum of chi squared's.
The method used for calculating the distribution of a weighted sum of independent chi-squared variables is given in \cite{imhof1961}.
A general method for computing any quadratic form of normals is presented, which is easily adapted for our case since
each chi-squared variable has only one degree of freedom. The exact formula given for the distribution function of $Q$, the weighted
sum of chi squares, is:
\begin{align*}
\mathbb{P}\{Q \geq x \} = \frac{1}{2} + \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
\end{align*}
For continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ which depend on the actual weights used.

There is no analytic solution to the integral, but integration can be carried out using numerical techniques. This seems
fine in most situations, but numerical issues do become a problem in the regime of large test statistics $x$. The function
$\theta(\cdot, x)$ is linear in $x$, and thus as the the test statistic grows the corresponding period of the integrand 
decreases. As the function acquires too many inflection points, the approximation becomes unstable. This is further
magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
elsewhere are quite prominent. Figure ? shows the non-monotonicity of the function as the test statistic grows.

There is fortunately a simple conservative approximation which can get around this numerical problem. Given the
following inequality:
\begin{align*}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
\end{align*}
We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
of a higher degree of freedom. Calculations for the cumulative distribution of a chi-squared distribution are stable
and already implemented in \proglang{R}, which makes this an even easier fix. Therefore, whenever the above approximation
yields a small p-values, we suspect numerical issues and use the preceding conservative approximation instead.

\section{Discussion}

Go into more detail about reasons for using these versus chi-squared type tests.

Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))

Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.

Benefits and drawbacks of using simulation to gather p-values.

Possibly go into a brief discussion of classes and class inheritance in R. The use of `htest' in our functions; benefits and issues with using it. Particularly troublesome in ks.test where we have a range of p-values.

Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure.

For further generalizations of tests see \cite{dewev1973}

\section{Examples}

%\bibliographystyle{plain}
\bibliography{jss}

\end{document}

### 2010-09-08 paper/outpaper.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@13 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-09-08 JSS/article.tex
@@
-\documentclass[article]{jss}
-
-\usepackage{amsthm, amsmath, amssymb}
+\documentclass[codesnippet]{jss}
 
@@
 %% almost as usual
-\author{Taylor B. Arnold \\ Yale University \And 
-        John W. Emerson\\ Yale University}
-\title{Nonparametric Goodness-of-Fit Tests \\ for Discrete Null Distributions}
+\author{Taylor B. Arnold\\Yale University \And 
+        John W. Emerson\\Yale University}
+\title{Non-parametric Goodness of Fit Tests for Discrete Distributions}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Taylor B. Arnold, John W. Emerson} %% comma-separated
-\Plaintitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% without formatting
-\Shorttitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% a short title (if necessary)
+\Plainauthor{Taylor Arnold, John Emerson} %% comma-separated
+\Plaintitle{Non-parametric Goodness of Fit Tests for Discrete Distributions} %% without formatting
+\Shorttitle{Non-parametric Goodness of Fit Tests for Discrete Distributions} %% a short title (if necessary)
 
@@
 \Abstract{
-   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
-   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
-   to users. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this need in the \proglang{R} language for
-   the two most popular nonparametric tests. The structure of these two packages are explained and examples
-   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
-   in their implementation.
+  Here should be the abstract.
 }
-\Keywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, \proglang{R}}
-\Plainkeywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, R} %% without formatting
+\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
+\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
 %% at least one keyword must be supplied
@@
 \Address{
-  Taylor B. Arnold \\
-  24 Hillhouse Ave. \\
+  Taylor B. Arnold\\
+  24 Hillhouse Ave.\\
   New Haven, CT 06511, USA\\
   E-mail: \email{taylor.arnold@yale.edu}\\
-  URL: \url{http://euler.stat.yale.edu/~tba3}\\
-
-  John W. Emerson \\
-  24 Hillhouse Ave. \\
-  New Haven, CT 06511, USA\\
-  E-mail: \email{john.emerson@yale.edu}\\
-  URL: \url{http://euler.stat.yale.edu/~jay}
+  URL: \url{http://euler.stat.yale.edu/~tba3}
 }
@@
 
-\section{Introduction}
-
-%General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
-Given an observed sequence of data, the typical question of statistical inference is to determine
-the underlying process from which it came. Given a particular null distribution, goodness-of-fit tests 
-are used in order to test whether it is likely that the data were generated via the null distribution.
-
-%Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
-While almost any hypothesis test can be viewed as a variant on a goodness-of-fit test, the term is typically
-applied to those tests which are nonparametric in nature. That is, they do not conduct the hypothesis testing 
-through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
-calculate differences between the empirical distribution of the observed data and the overall distribution 
-of the null model. In many cases these tests are preferential since they tend to have an increased power for 
-interesting deviations from the null model, whilst allowing for more robustness due to uninteresting deviations
-such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and Smirnov,
-and several variants on the estimated proposed by Cramer von Mises.
-
-% Discrete versions exist; just no code
-While the original aim of nonparametric goodness of fit tests was more continuous null distributions, discrete versions
-have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
-This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
-apply the functions meant for continuous distributions. As shown in section 5, the latter can be particularly dangerous
-in the small sample setting. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this void for users
-operating in the \proglang{R} environment.
-
-\section{Kolmogorov-Smirnov Test}
-
-% Most popular choice of test; explain the basic algorithm and purpose for continuous distributions (maybe a few P's).
-
-\subsection{Overview}
-
-Of all the methods for nonparametric goodness-of-fit tests, by far the most popular is the method devised by Kolmogorov and Smirnov.
-It is the only such test built into the base of \proglang{R}. The idea behind it is fairly simple. Given the cumulative distribution
-function $F_0(x)$ of the continuous null distribution, and the empirical distribution function $F_{data}(x)$ of the observed data, 
-one constructs the statistic:
-\begin{align}
-D = \sup_x \left| F_0(x)- F_{data}(x) \right|
-\end{align}
-The difficult part of the theory is showing that under the null model $D$ has a fixed distribution. Perhaps most interestingly is the fact 
-that this distribution does not depend on which null model was specified. Given this, it is easy to read p-values for the test statistic 
-off of a table or computer function.
-
-The extension of this result to non-continuous null distributions does not have such a clean solution. While the calculation of
-the test statistic $D$ is still just as easy, the distribution of the testing statistic is much more difficult. Unlike in the 
-continuous case, it ultimately depends on which null model was chosen; this makes it impossible to simply read p-values directly 
-off of a fixed table. It was known since at least the 1950's that using the tables for continuous distributions resulted in 
-conservative p-values; it was not until \citet{Conover1972} that a method for computing exact p-values in this case was developed.
-
-Two common alternatives of the above test statistic exit, and are also encoded in \proglang{R}. The absolute value is discarded and the rest is either
-left alone (the `greater' testing alternative) or the supremum is replaces with a negative infimum (the `lesser' hypothesis alternative). These can
-be helpful depending on the nature of the alternative hypotheses for which the test is desired to be powerful towards.
-
-\subsection{Implementation}
-
-The implementation of the discrete Kolmogorov-Smirnov function consists of two parts. First the particular test statistic needs to be calculated,
-and then the p-value for that particular statistic must be computed. These are the same general procedures as for the default function 
-\pkg{ks.test} in \proglang{R}. Therefore, the basic implementation here attempts to incorporate the discrete test into the default function. There is
-a default built-in class for step-functions and therefore it is easy to write code which checks for the input null distribution being a step-function
-and to divert these cases to the implementation below. Other null distributions are handled as they were previously.
-
-Given that the test statistic is, theoretically, the same as in the
-continuous case it would seem that the first part could be directly taken from the existing code. This is unfortunately not the case. Given the
-quantity $x$, which is a vector consisting of the difference between the cumulative null distribution and cumulative empirical distribution at each
-observed data point, the default in \proglang{R} to determine the Kolmogorov-Smirnov test statistic is (with modifications made for readability):
-\begin{CodeChunk}
-\begin{CodeInput}
-  STATISTIC <- max(c(abs(x), abs(x-1/n)))
-\end{CodeInput}
-\end{CodeChunk}
-This method utilizes the fact that the supremum in equation 1 is the absolute difference between a continuous function and stepwise constant
-function; this supremum can be reached either at an observed data point, or for a sequence of points approaching an observed data point from the 
-left. This particular fact is still true if both functions are stepwise constant, but the problem comes from the fact that this code assumes, with the code
-\code{abs(x-1/n)}, that the cumulative null distribution is continuous at the observed data points. This should never be the case for discrete null
-distributions. This gives the incorrect default behavior:
-\begin{CodeChunk}
-\begin{CodeInput}
-  R>  ks.test(c(1,2), ecdf(c(1,2))) 
-\end{CodeInput}
-\begin{CodeOutput}
-	One-sample Kolmogorov-Smirnov test
-
-  data:  c(1, 2) 
-  D = 0.5, p-value = 0.5
-  alternative hypothesis: two-sided 
-\end{CodeOutput}
-\end{CodeChunk}
-Where clearly the test statistic should have been $0$, since the null and empirical cumulative distributions are the same.
-
-Getting around this
-issue is easy to do by computing the test statistic as \code{max(abs(x))} in the discrete case. The important point here is that \proglang{R}'s
-default behavior for computing the test statistic was dependent on the continuity of the null distribution, even though in theory the discreteness
-should not have affected the calculations.
-
-Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
-continuous case. The default options for \pkg{ks.test} already allow the user to specify if they would like approximate asymptotic distributions (since these exist
-in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
-the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the \proglang{R} language, or in any other open source options, did not previously
-exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
-the original Conover paper.
-
-%Numerical issues arise from the summation; need to do something else for high $n$. Simulation? Asymptotic approximation? 
-The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.
-
-\section{Cramer von Mises Tests}
-
-\subsection{Algorithm}
-
-%General idea for continuous distributions. In many cases actually more powerful of the two. There again exists a distribution free 
-%distribution for the testing statistic.
-While the Kolmogorov-Smirnov test is the most well known of the non-parametric goodness of fit tests, there is another family of 
-tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
-Harald Cramer and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted in \cite{anderson1952}, and 
-\cite{Watson1961}. The test statistics are, respectively, given as:
-\begin{align*}
-W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
-A^2 &= n \cdot \int_{-\infty}^{\infty} \left[F_0(x) -F_0(x)^2 \right]^{-1} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
-U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
-\end{align*}
-Where $F$ is either the cumulative distribution of the null model or the empirical cumulative distribution of the observed data.
-As in the Kolmogorov-Smirnov test statistic, these all have distribution free null distributions in the continuous case.
-
-It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations. As they all involve
-integration over the whole range of data, rather than one supremum, it is not surprising that they are generally best when the
-true alternative distribution deviates a little over the whole range of data rather than deviating a lot over a small range. For
-a complete analysis of the relative powers of these tests see \cite{stephens1974}.
-
-%Discuss the discrete version. How to calculate W2, U2, and A2. 
-Generalizations of the Cramer-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
-the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
-The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
-suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
-chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation
-
-\subsection{Implementation}
-
-For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
-alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
-As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
-should also originally implement the continuous case even though this is not what we concentrate on here.
-
-Calculation of the three test statistics is done by straightforward matrix algebra as given in \cite{choulakian1994}. Determining 
-the form of the asymptotic null distribution is also easy using the built-in eigenvalue decomposition functions. The only 
-difficulty in the process involves actually calculating the percentiles for these weighted chi-squares.
-
-%Show how to compute the distribution of a weighted sum of chi squared's.
-The method used for calculating the distribution of a weighted sum of independent chi-squared variables is given in \cite{imhof1961}.
-A general method for computing any quadratic form of normals is presented, which is easily adapted for our case since
-each chi-squared variable has only one degree of freedom. The exact formula given for the distribution function of $Q$, the weighted
-sum of chi squares, is:
-\begin{align*}
-\mathbb{P}\{Q \geq x \} = \frac{1}{2} + \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
-\end{align*}
-For continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ which depend on the actual weights used.
-
-There is no analytic solution to the integral, but integration can be carried out using numerical techniques. This seems
-fine in most situations, but numerical issues do become a problem in the regime of large test statistics $x$. The function
-$\theta(\cdot, x)$ is linear in $x$, and thus as the the test statistic grows the corresponding period of the integrand 
-decreases. As the function acquires too many inflection points, the approximation becomes unstable. This is further
-magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
-elsewhere are quite prominent. Figure ? shows the non-monotonicity of the function as the test statistic grows.
-
-There is fortunately a simple conservative approximation which can get around this numerical problem. Given the
-following inequality:
-\begin{align*}
-\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
-&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
-\end{align*}
-We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
-of a higher degree of freedom. Calculations for the cumulative distribution of a chi-squared distribution are stable
-and already implemented in \proglang{R}, which makes this an even easier fix. Therefore, whenever the above approximation
-yields a small p-values, we suspect numerical issues and use the preceding conservative approximation instead.
-
-\section{Discussion}
-
-Go into more detail about reasons for using these versus chi-squared type tests.
-
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))
-
-Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
-
-Benefits and drawbacks of using simulation to gather p-values.
-
-Possibly go into a brief discussion of classes and class inheritance in R.
-The use of `htest' in our functions; benefits and issues with using it.
-Particularly troublesome in ks.test where we have a range of p-values.
-
-Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure.
-
-For further generalizations of tests see \cite{dewev1973}
-
-\section{Examples}
-
-%\bibliographystyle{plain}
-\bibliography{jss}
+\section[About Java]{About \proglang{Java}}
+%% Note: If there is markup in \(sub)section, then it has to be escape as above.
 

### 2010-09-08 code.R

@@ -1,0 +1,114 @@
+N <- seq(10, 500, by = 20)
+P <- seq(10,200, by=10)
+
+N <- seq(60, 85, by = 2)
+P <- seq(10,30, by=1)
+P <- rep(25, 15)
+
+Q <- matrix(NA, nrow=length(P), ncol=length(N))
+for(i in 1:length(N)) {
+  for(j in 1:length(P)) {
+    n <- N[i]
+    p <- P[j]
+    y <- ecdf(1:p)
+    x <- sample(1:p, n, replace=TRUE)
+    Q[j,i] <- ks.test(x,y, alternative="less")$p.value
+  }
+}
+rownames(Q) <- P
+colnames(Q) <- N
+
+plot(1, 0, xlim=c(0,500), ylim=c(-650,650), col="white", xlab="Sample Size", ylab="log(|P-value|) * sign(P-value)")
+
+for(i in 1:nrow(Q)) {
+  lines(N, log(abs(Q[i,]))*sign(Q[i,]))
+}
+
+Q <- matrix(NA, nrow=40, ncol=82) 
+A <- rep(NA, 40)
+
+for(i in 1:40 ) {
+   x <- sample(1:25, 82, replace=TRUE)
+   y <- ecdf(1:25)
+   Q[i,] <- x
+   A[i] <- ks.test(x,y, alternative="less")$p.value
+}
+
+
+
+
+# x = data points
+# y = stepfunction of cdf
+cvm.stat <- function(x,y, type=c(W2, U2, A2)) {
+  I <- knots(y)
+  N <- length(x)
+  e <- diff(c(0,N*y(I)))
+  obs <- rep(0, length(I))
+  for(j in 1:length(I)) {
+    obs[j] <- length(which(x == I[j]))
+  }
+  S <- cumsum(obs)
+  T <- cumsum(e)
+  H <- T/N
+  p <- e/N
+  t <- (p + p[c(2:length(p), 1)])/2
+  Z <- S - T
+  Zbar <- sum(Z*t)
+
+  # Test Statistics:
+  W2 <- sum(Z^2*t)/N
+  U2 <- sum((Z-Zbar)^2*t)/N
+  A2 <- sum((Z^2*t/(H*(1-H) ))[-length(I)])/N
+  return(c(W2, U2, A2))
+}
+
+cvm.dist <- function(STAT, x,y, type=c(W2, U2, A2)) {
+
[…]

### 2010-10-10 JSS/article.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@14 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@15 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-09-08 JSS/article.tex
@@
 \usepackage{amsthm, amsmath, amssymb}
+%\usepackage{Sweave}
 
@@
 
+\setkeys{Gin}{width=1.0\textwidth, height=1.0\textheight}
+
+<<setup,include=FALSE,echo=FALSE>>=
+options(keep.source = TRUE, width = 60)
+@
+
 %% include your article here, just as usual
@@
 
+Example code:
+<<>>=
+1+1
+@
+
 %General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
-Given an observed sequence of data, the typical question of statistical inference is to determine
-the underlying process from which it came. Given a particular null distribution, goodness-of-fit tests 
+%Given an observed sequence of data, the typical question of statistical inference is to determine
+%the underlying process from which it came. 
+Given a particular null distribution, goodness-of-fit tests 
 are used in order to test whether it is likely that the data were generated via the null distribution.
@@
 of the null model. In many cases these tests are preferential since they tend to have an increased power for 
-interesting deviations from the null model, whilst allowing for more robustness due to uninteresting deviations
-such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and Smirnov,
-and several variants on the estimated proposed by Cramer von Mises.
+interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
+factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
+Smirnov, followed closely by several variants on an estimation procedure proposed by Cramer von Mises.
 
 % Discrete versions exist; just no code
-While the original aim of nonparametric goodness of fit tests was more continuous null distributions, discrete versions
+While the original aim of nonparametric goodness of fit tests was meant for continuous null distributions, discrete versions
 have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
@@
 \end{align}
-The difficult part of the theory is showing that under the null model $D$ has a fixed distribution. Perhaps most interestingly is the fact 
-that this distribution does not depend on which null model was specified. Given this, it is easy to read p-values for the test statistic 
-off of a table or computer function.
+The distribution of $D$ under the null model does not depend on which null distribution is being used, making this a computationally
+attractive method. For a standard treatment of the test and its performance relative to other algorithms see \cite{slakter65}. Two common
+alternatives of the above test statistic exist. The absolute value is discarded and the rest is either left alone (the `greater' testing 
+alternative) or the supremum is replaces with a negative infimum (the `lesser' hypothesis alternative). These can be helpful depending 
+on the nature of the alternative hypotheses for which the test is desired to be powerful towards.
 
-The extension of this result to non-continuous null distributions does not have such a clean solution. While the calculation of
-the test statistic $D$ is still just as easy, the distribution of the testing statistic is much more difficult. Unlike in the 
+The extension of this result to non-continuous null distributions does not have such a clean solution. The formula of
+the test statistic $D$ remains unchanged, however the distribution of the testing statistic is much more difficult. Unlike in the 
 continuous case, it ultimately depends on which null model was chosen; this makes it impossible to simply read p-values directly 
@@
 
-Two common alternatives of the above test statistic exit, and are also encoded in \proglang{R}. The absolute value is discarded and the rest is either
-left alone (the `greater' testing alternative) or the supremum is replaces with a negative infimum (the `lesser' hypothesis alternative). These can
-be helpful depending on the nature of the alternative hypotheses for which the test is desired to be powerful towards.
-
 \subsection{Implementation}
@@
 The implementation of the discrete Kolmogorov-Smirnov function consists of two parts. First the particular test statistic needs to be calculated,
-and then the p-value for that particular statistic must be computed. These are the same general procedures as for the default function 
-\pkg{ks.test} in \proglang{R}. Therefore, the basic implementation here attempts to incorporate the discrete test into the default function. There is
-a default built-in class for step-functions and therefore it is easy to write code which checks for the input null distribution being a step-function
-and to divert these cases to the implementation below. Other null distributions are handled as they were previously.
+and then the p-value for that particular statistic must be computed.
 
 Given that the test statistic is, theoretically, the same as in the
-continuous case it would seem that the first part could be directly taken from the existing code. This is unfortunately not the case. Given the
-quantity $x$, which is a vector consisting of the difference between the cumulative null distribution and cumulative empirical distribution at each
-observed data point, the default in \proglang{R} to determine the Kolmogorov-Smirnov test statistic is (with modifications made for readability):
-\begin{CodeChunk}
-\begin{CodeInput}
-  STATISTIC <- max(c(abs(x), abs(x-1/n)))
-\end{CodeInput}
-\end{CodeChunk}
-This method utilizes the fact that the supremum in equation 1 is the absolute difference between a continuous function and stepwise constant
-function; this supremum can be reached either at an observed data point, or for a sequence of points approaching an observed data point from the 
-left. This particular fact is still true if both functions are stepwise constant, but the problem comes from the fact that this code assumes, with the code
-\code{abs(x-1/n)}, that the cumulative null distribution is continuous at the observed data points. This should never be the case for discrete null
-distributions. This gives the incorrect default behavior:
-\begin{CodeChunk}
-\begin{CodeInput}
-  R>  ks.test(c(1,2), ecdf(c(1,2))) 
-\end{CodeInput}
-\begin{CodeOutput}
-	One-sample Kolmogorov-Smirnov test
+continuous case it would seem that the first part could be directly taken from the existing procedures. This is, unfortunately, often not the case.
+Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
+is continuous. If we want to determine the supreme of the difference between these two functions notice that:
 
-  data:  c(1, 2) 
-  D = 0.5, p-value = 0.5
-  alternative hypothesis: two-sided 
-\end{CodeOutput}
-\end{CodeChunk}
-Where clearly the test statistic should have been $0$, since the null and empirical cumulative distributions are the same.
+\begin{align}
+\sup_x \left| f(x)- g(x) \right| &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
+&=  \max_i \left( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \right)
+\end{align}
 
-Getting around this
-issue is easy to do by computing the test statistic as \code{max(abs(x))} in the discrete case. The important point here is that \proglang{R}'s
-default behavior for computing the test statistic was dependent on the continuity of the null distribution, even though in theory the discreteness
-should not have affected the calculations.
+Computing the maximum over these $2N$ values (with $f$ equal to $F_{data}(x)$ and $g$ equal to $F_0(x)$ as defined above) is clearly the 
+most efficient way to compute the Kolmogorov-Smirnov test statistic when given a continuous null distribution. When the function $g$ is not
+continuous, notice that this formula no longer works since in general we cannot replace the limit of $g$ with its value at $x_i$. If it is known that $g$
+is also a step function, we could replace the formula for some small $\epsilon$ by:
+\begin{align}
+\sup_x \left| f(x)- g(x) \right| &=  \max_i \left( \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \right)
+\end{align}
+Where the discontinuities in $g$ are at least a distance $\epsilon$ apart.
 
-Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
-continuous case. The default options for \pkg{ks.test} already allow the user to specify if they would like approximate asymptotic distributions (since these exist
-in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
-the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the \proglang{R} language, or in any other open source options, did not previously
-exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
-the original Conover paper.
+%Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
+%continuous case. The default options for \pkg{ks.test} already allow the user to specify if they would like approximate asymptotic distributions (since these exist
+%in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
+%the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the \proglang{R} language, or in any other open source options, did not previously
+%exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
+%the original Conover paper.
 
 %Numerical issues arise from the summation; need to do something else for high $n$. Simulation? Asymptotic approximation? 
-The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.
+%The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.
 
@@
 
-\subsection{Algorithm}
+\subsection{Overview}
 
@@
 tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
-Harald Cramer and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted in \cite{anderson1952}, and 
+Harald Cramer and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted by \cite{anderson1952}, and 
 \cite{Watson1961}. The test statistics are, respectively, given as:
@@
 \subsection{Implementation}
-
-For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
-alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
-As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
-should also originally implement the continuous case even though this is not what we concentrate on here.
 
@@
 We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
-of a higher degree of freedom. Calculations for the cumulative distribution of a chi-squared distribution are stable
-and already implemented in \proglang{R}, which makes this an even easier fix. Therefore, whenever the above approximation
-yields a small p-values, we suspect numerical issues and use the preceding conservative approximation instead.
+of a higher degree of freedom.
+
+\section{Kolmogorov-Smirnov and Cramer von Mises in R}
+
+For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
+alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
+As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
+should also originally implement the continuous case even though this is not what we concentrate on here.
+
+\section{Examples}
 
@@
 
-Go into more detail about reasons for using these versus chi-squared type tests.
+Go into more detail about reasons for using these versus chi-squared type tests. (*)
 
@@
 
-Benefits and drawbacks of using simulation to gather p-values.
+Benefits and drawbacks of using simulation to gather p-values. (*)
 
@@
 
-Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure.
+Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure. (*)
 

### 2010-10-10 JSS/article.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@15 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\setkeys{Gin}{width=1.0\textwidth, height=1.0\textheight}
+
 %% include your article here, just as usual
@@
 
+Example code:
+\begin{Schunk}
+\begin{Sinput}
+> 1 + 1
+\end{Sinput}
+\begin{Soutput}
+[1] 2
+\end{Soutput}
+\end{Schunk}
+
 %General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
-Given an observed sequence of data, the typical question of statistical inference is to determine
-the underlying process from which it came. Given a particular null distribution, goodness-of-fit tests 
+%Given an observed sequence of data, the typical question of statistical inference is to determine
+%the underlying process from which it came. 
+Given a particular null distribution, goodness-of-fit tests 
 are used in order to test whether it is likely that the data were generated via the null distribution.
@@
 We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
-of a higher degree of freedom. Calculations for the cumulative distribution of a chi-squared distribution are stable
-and already implemented in \proglang{R}, which makes this an even easier fix. Therefore, whenever the above approximation
-yields a small p-values, we suspect numerical issues and use the preceding conservative approximation instead.
+of a higher degree of freedom.
 

### 2010-10-12 JSS/article.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@16 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
    in their implementation.
+
+*** try to clean up, avoid package/function issue, make it a general
+contribution if possible.
 }
@@
 options(keep.source = TRUE, width = 60)
+set.seed(1)
 @
@@
 operating in the \proglang{R} environment.
+
+** Ditto on wordsmithing the contribution (package/function, etc...).
 
@@
 
+*** Check Cramer-von Mises for hyphen, and let's get the accent done properly.
+
 \section{Cramer von Mises Tests}
@@
 
+*** New paragraph, perhaps here, on the differences between these three tests.
+
 It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations. As they all involve
@@
 chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation
+the original papers shows that this approximation (*** NOT FINISHED?)
 
@@
 
-\section{Kolmogorov-Smirnov and Cramer von Mises in R}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Kolmogorov-Smirnov and Cramer von Mises Tests in R}
 
-For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
-alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
-As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
-should also originally implement the continuous case even though this is not what we concentrate on here.
+Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
+\proglang{R}'s Kolmogorov-Smirnov function \code{ks.test()} from base
+package \pkg{stats}.
+Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
+function for Cramer von Mises tests.  Both are available from the authors
+as of October, 2010,
+or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
+they will be proposed for inclusion in \pkg{stats} in late 2010.
 
+The revised \pkg{ks.test()} function supports one-sample tests for discrete
+null distributions by allowing the second argument, \code{y}, to be
+an empirical cumulative distribution function (an \proglang{R} function
+with class \code{ecdf}) or an object of class \code{stepfun} specifying
+a discrete distribution.  As in the original version of \code{ks.test()},
+the presence of ties in the data (the first argument, \code{x}) generates a
+warning unless \code{y} describes a discrete distribution.  When a discrete
+distribution is specified, exact p-values are not available for 
+two-sided alternative hypotheses, but the reported p-values will be
+conservative.  For one-sided tests,
+exact p-values are calculated using Conover's
+method (when \code{exact = NULL} or \code{exact = TRUE})
+if the sample size is less than or equal to 30; otherwise, asymptotic distributions
+are used which are reliable in such cases (CITATION? Is this correct?).
+When \code{exact = FALSE} the asymptotic distribution is used which is known
+to be imprecise but conservative, even for small sample (CITATION).
+
+*** Discussion of what we might have broken: cases where the user
+provided a discrete distribution to the original ks.test() even though
+it wasn't intended to be supported. ***
+
+The function \pkg{cvm.test()} is similar to \pkg{ks.test()}.  Its first two
+arguments specify the data and null distribution; the only extra option,
+\code{type}, specifies the variant of the Cramer-von Mises test:
+\begin{itemize}
+\item \code{x}: a numerical vector of data values.
+\item \code{y}: an \code{ecdf} or step-function (\code{stepfun}) for specifying
+the null model
+\item \code{type}: the variant of the Cramer-von Mises test; \code{W2}
+is the default and most common method, \code{U2} is for cyclical data,
+and \code{A2} is the Anderson-Darling alternative.
+\end{itemize}
+As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
+\code{htest}.
+
+%%%%%%%%%%%%%%%%%%
 \section{Examples}
+
+Hmm... I thought the package had a trivial example where the standard
+\code{ks.test()} was a problem?
+<<ks.test>>=
+library(ks.test)
+x <- sample(1:10, 25, replace=TRUE)
+x
+ks.test(x, ecdf(1:10))
+ks.test(x, ecdf(1:10), alternative="g")
+ks.test(x, ecdf(1:10), alternative="g", exact=FALSE)
+@
+
+Discuss this, nice example.  Relate to new material that is about to go into
+the earlier section.
+<<>>=
+stats::ks.test(c(0,1), ecdf(c(0,1)))
+ks.test(c(0,1), ecdf(c(0,1)))
+@
+
+JAY: simulation showing original ks.test() was conservative, Conover exact.
+
+<<cvm.test>>=
+library(cvm.test)
+cvm.test(x, ecdf(1:10))
+cvm.test(x, ecdf(1:10), type='A2')
+@
+
+TAYLOR: good cyclical example, relating to new material about to go into
+earlier section.
 
@@
 
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))
+Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?
 
 Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
+Earlier, needed here?
 
@@
 Particularly troublesome in ks.test where we have a range of p-values.
+More precisely, explain that we
+could provide upper and lower bounds for the p-value but are limited by class
+\code{htest}.
 
@@
 
-\section{Examples}
+Future work: continuous CVM.
+
+*** Bibliography needs cleaning before we finish up, just don't forget.
 

### 2010-10-12 JSS/article.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@16 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+** Ditto on wordsmithing the contribution (package/function, etc...).
+
 \section{Kolmogorov-Smirnov Test}
@@
 
+*** Check Cramer-von Mises for hyphen, and let's get the accent done properly.
+
 \section{Cramer von Mises Tests}
 
-\subsection{Algorithm}
+\subsection{Overview}
 
@@
 chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation
+the original papers shows that this approximation (*** NOT FINISHED?)
 
 \subsection{Implementation}
-
-For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
-alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
-As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
-should also originally implement the continuous case even though this is not what we concentrate on here.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Kolmogorov-Smirnov and Cramer von Mises Tests in R}
+
+Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
+\proglang{R}'s Kolmogorov-Smirnov function \code{ks.test()} from base
+package \pkg{stats}.
+Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
+function for Cramer von Mises tests.  Both are available from the authors
+as of October, 2010,
+or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
+they will be proposed for inclusion in \pkg{stats} in late 2010.
+
+The revised \pkg{ks.test()} function supports one-sample tests for discrete
+null distributions by allowing the second argument, \code{y}, to be
+an empirical cumulative distribution function (an \proglang{R} function
+with class \code{ecdf}) or an object of class \code{stepfun} specifying
+a discrete distribution.  As in the original version of \code{ks.test()},
+the presence of ties in the data (the first argument, \code{x}) generates a
+warning unless \code{y} describes a discrete distribution.  When a discrete
+distribution is specified, exact p-values are not available for 
+two-sided alternative hypotheses, but the reported p-values will be
+conservative.  For one-sided tests,
+exact p-values are calculated using Conover's
+method (when \code{exact = NULL} or \code{exact = TRUE})
+if the sample size is less than or equal to 30; otherwise, asymptotic distributions
+are used which are reliable in such cases (CITATION? Is this correct?).
+When \code{exact = FALSE} the asymptotic distribution is used which is known
+to be imprecise but conservative, even for small sample (CITATION).
+
+*** Discussion of what we might have broken: cases where the user
+provided a discrete distribution to the original ks.test() even though
+it wasn't intended to be supported. ***
+
+The function \pkg{cvm.test()} is similar to \pkg{ks.test()}.  Its first two
+arguments specify the data and null distribution; the only extra option,
+\code{type}, specifies the variant of the Cramer-von Mises test:
+\begin{itemize}
+\item \code{x}: a numerical vector of data values.
+\item \code{y}: an \code{ecdf} or step-function (\code{stepfun}) for specifying
+the null model
+\item \code{type}: the variant of the Cramer-von Mises test; \code{W2}
+is the default and most common method, \code{U2} is for cyclical data,
+and \code{A2} is the Anderson-Darling alternative.
+\end{itemize}
+As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
+\code{htest}.
+
+%%%%%%%%%%%%%%%%%%
+\section{Examples}
+
+Hmm... I thought the package had a trivial example where the standard
+\code{ks.test()} was a problem?
+\begin{Schunk}
+\begin{Sinput}
+> library(ks.test)
+> x <- sample(1:10, 25, replace = TRUE)
+> x
+\end{Sinput}
+\begin{Soutput}
+ [1]  3  4  6 10  3  9 10  7  7  1  3  2  7  4  8  5  8 10
+[19]  4  8 10  3  7  2  3
+\end{Soutput}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10))
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D = 0.08, p-value = 1
+alternative hypothesis: two-sided 
+\end{Soutput}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10), alternative = "g")
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D^+ = 0.04, p-value = 0.7731
+alternative hypothesis: the CDF of x lies above the null hypothesis 
+\end{Soutput}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10), alternative = "g", exact = FALSE)
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D^+ = 0.04, p-value = 0.9231
+alternative hypothesis: the CDF of x lies above the null hypothesis 
+\end{Soutput}
+\end{Schunk}
+
+Discuss this, nice example.  Relate to new material that is about to go into
+the earlier section.
+\begin{Schunk}
+\begin{Sinput}
+> stats::ks.test(c(0, 1), ecdf(c(0, 1)))
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  c(0, 1) 
+D = 0.5, p-value = 0.5
+alternative hypothesis: two-sided 
+\end{Soutput}
+\begin{Sinput}
+> ks.test(c(0, 1), ecdf(c(0, 1)))
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  c(0, 1) 
+D = 0, p-value = 1
+alternative hypothesis: two-sided 
+\end{Soutput}
+\end{Schunk}
+
+JAY: simulation showing original ks.test() was conservative, Conover exact.
+
+\begin{Schunk}
+\begin{Sinput}
+> library(cvm.test)
+> cvm.test(x, ecdf(1:10))
+\end{Sinput}
+\begin{Soutput}
+	Cramer-von Mises - W2
+
+data:  x 
+W2 = 0.057, p-value = 0.8114
+alternative hypothesis: Two.sided 
+\end{Soutput}
+\begin{Sinput}
+> cvm.test(x, ecdf(1:10), type = "A2")
+\end{Sinput}
+\begin{Soutput}
+	Cramer-von Mises - A2
+
+data:  x 
+A2 = 0.3969, p-value = 0.75
+alternative hypothesis: Two.sided 
+\end{Soutput}
+\end{Schunk}
+
+TAYLOR: good cyclical example, relating to new material about to go into
+earlier section.
+
 \section{Discussion}
 
-Go into more detail about reasons for using these versus chi-squared type tests.
+Go into more detail about reasons for using these versus chi-squared type tests. (*)
 
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))
+Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?
 
 Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
+Earlier, needed here?
 
-Benefits and drawbacks of using simulation to gather p-values.
+Benefits and drawbacks of using simulation to gather p-values. (*)
 
@@
 Particularly troublesome in ks.test where we have a range of p-values.
+More precisely, explain that we
+could provide upper and lower bounds for the p-value but are limited by class
+\code{htest}.
 
-Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure.
+Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure. (*)
 

### 2010-10-13 JSS/article.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@17 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
-is continuous. If we want to determine the supreme of the difference between these two functions notice that:
+is continuous. If we want to determine the supremum of the difference between these two functions notice that:
 
@@
 \end{align}
-Where the discontinuities in $g$ are at least a distance $\epsilon$ apart.
+Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function. In the case of not knowing,
+or having a $g$ which is in neither a step function nor a continuous function, the only method for computing the supremum is to take the numerical limit in (1). 
+This clearly takes more computational time than simply cycling over $2N$ values.
+
+Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, the user must either indicate the points of discontinuity of the null
+distribution's cdf or live with a much smaller.
 

### 2010-10-13 JSS/article.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@17 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-10-10 JSS/article.tex
@@
 
+\section{Kolmogorov-Smirnov and Cramer von Mises in R}
+
+For the user, the function \pkg{cvm.test} was designed to work in exactly the same manner as \pkg{ks.test} in \proglang{R}. This
+alleviates the need to learn a new set of input and output instructions and hopefully will increase the usage of these tests. 
+As there is no default functionality in \proglang{R} for continuous Cramer-von Mises, subsequent versions of the package
+should also originally implement the continuous case even though this is not what we concentrate on here.
+
+\section{Examples}
+
 \section{Discussion}
 
-Go into more detail about reasons for using these versus chi-squared type tests.
+Go into more detail about reasons for using these versus chi-squared type tests. (*)
 
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))
+Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?
 
 Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
+Earlier, needed here?
 
-Benefits and drawbacks of using simulation to gather p-values.
+Benefits and drawbacks of using simulation to gather p-values. (*)
 

### 2010-10-14 JSS/article.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@18 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\begin{Schunk}
+\begin{Sinput}
+set.seed(1)
+REPS <- 100000
+asymp.pval <- rep(0, REPS)
+exact.pval <- rep(0, REPS)
+for (i in 1:REPS) {
+  x <- sample(1:10, 15, replace=TRUE)
+  asymp.pval[i] <- stats::ks.test(x, ecdf(1:10), alternative="g")$p.value
+  exact.pval[i] <- ks.test::ks.test(x, ecdf(1:10), alternative="g")$p.value
+}
+sum(asymp.pval<exact.pval)/REPS
+sum(asymp.pval<=0.05)/REPS
+sum(exact.pval<=0.05)/REPS
+sqrt(0.05*0.95/REPS)
+\end{Sinput}
+\end{Schunk}
+
+\begin{verbatim}
+#####################################################
+# 100000:
+
+> sum(asymp.pval<exact.pval)/REPS
+[1] 0
+> sum(asymp.pval<=0.05)/REPS
+[1] 0.01848
+> sum(exact.pval<=0.05)/REPS
+[1] 0.03763
+> 
+> sqrt(0.05*0.95/REPS)
+[1] 0.0006892024
+> 
+> REPS
+[1] 1e+05
+\end{verbatim}
+
 JAY: simulation showing original ks.test() was conservative, Conover exact.

### 2010-10-18 JSS/article_save.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@19 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-10-13 JSS/article.tex
@@
    in their implementation.
-
-*** try to clean up, avoid package/function issue, make it a general
-contribution if possible.
 }
@@
 
+<<setup,include=FALSE,echo=FALSE>>=
+options(keep.source = TRUE, width = 60)
+@
+
 %% include your article here, just as usual
@@
 Example code:
-\begin{Schunk}
-\begin{Sinput}
-> 1 + 1
-\end{Sinput}
-\begin{Soutput}
-[1] 2
-\end{Soutput}
-\end{Schunk}
+<<>>=
+1+1
+@
 
@@
 operating in the \proglang{R} environment.
-
-** Ditto on wordsmithing the contribution (package/function, etc...).
 
@@
 \end{align}
-Where the discontinuities in $g$ are at least a distance $\epsilon$ apart.
+Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function. In the case of not knowing,
+or having a $g$ which is in neither a step function nor a continuous function, the only method for computing the supremum is to take the numerical limit in (1). 
+This clearly takes more computational time than simply cycling over $2N$ values.
+
+Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, the user must either indicate the points of discontinuity of the null
+distribution's cdf or live with a much smaller.
+
+Once the test statistic has been determined
 
@@
 %The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.
-
-*** Check Cramer-von Mises for hyphen, and let's get the accent done properly.
 
@@
 
-*** New paragraph, perhaps here, on the differences between these three tests.
-
 It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations. As they all involve
@@
 chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation (*** NOT FINISHED?)
+the original papers shows that this approximation
 
@@
 
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?
+Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2))
 
 Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
-Earlier, needed here?
 
@@
 Particularly troublesome in ks.test where we have a range of p-values.
-More precisely, explain that we
-could provide upper and lower bounds for the p-value but are limited by class
-\code{htest}.
 
@@
 
-Future work: continuous CVM.
-
-*** Bibliography needs cleaning before we finish up, just don't forget.
+\section{Examples}
 

### 2010-10-18 JSS/codeforgraphics.R

@@ -1,0 +1,59 @@
+setwd('/home/tba3/Desktop/today/ks-test/JSS')
+
+  cvm.pval <- function(STAT, lambda, good=TRUE) {
+
+    x <- STAT
+
+    theta <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+         VAL <- VAL + 0.5*atan(lambda[i]*u)
+      }
+      return(VAL - 0.5*x*u)
+    }
+
+    rho <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+        VAL <- VAL + log(1 + lambda[i]^2*u^2)     
+      }
+      VAL <- exp(VAL*0.25)
+      return(VAL)
+    }
+
+    fun <- function(u) return(sin(theta(u))/(u*rho(u)))
+
+    pval <- 0
+    try(pval <- 0.5 + integrate(fun, 0, Inf, subdivisions=1e6)$value/pi,
+        silent=TRUE)
+    if(pval > 0.001 || good == FALSE) return(pval)
+    if(pval <= 0.001 & good == TRUE) {
+      df <- sum(lambda != 0)
+      return(dchisq(STAT/max(lambda),df))
+    }
+  } # End cvm.pval()
+
+lambda <- sort(runif(10))
+cvm.pval(1, lambda)
+
+I <- seq(0,35,length.out=1000)
+y <- rep(NA, length(I))
+for(i in 1:length(I)) {
+  y[i] <- cvm.pval(I[i], lambda, good=TRUE)
+}
+z <- rep(NA, length(I))
+for(i in 1:length(I)) {
+  z[i] <- cvm.pval(I[i], lambda, good=FALSE)
+}
+
+#pdf('fig1.pdf', height=6, width=8)
+plot(I[I > 18], z[I > 18], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
+abline(h=0, lty=2)
+lines(I[I > 18], y[I > 18], lty=2, col='blue', lwd=2)
+lines(I[I > 18], z[I > 18], col='salmon', lwd=2)
+#dev.off()
+
+
+
+
+

### 2010-11-05 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@20 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-10-14 JSS/article.tex
@@
-\documentclass[article]{jss}
+\title{Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions}
+\author{by Taylor B. Arnold and John W. Emerson}
 
-\usepackage{amsthm, amsmath, amssymb}
-%\usepackage{Sweave}
+\maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% almost as usual
-\author{Taylor B. Arnold \\ Yale University \And 
-        John W. Emerson\\ Yale University}
-\title{Nonparametric Goodness-of-Fit Tests \\ for Discrete Null Distributions}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Taylor B. Arnold, John W. Emerson} %% comma-separated
-\Plaintitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% without formatting
-\Shorttitle{Nonparametric Goodness-of-Fit Tests for Discrete Distributions} %% a short title (if necessary)
-
-%% an abstract and keywords
-\Abstract{
+\abstract{
    A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
    for several decades. Despite this, modern statistical software have generally failed to provide this methodology
-   to users. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this need in the \proglang{R} language for
+   to users. The packages ks.test and cvm.test serve to fill this need in the R language for
    the two most popular nonparametric tests. The structure of these two packages are explained and examples
@@
 }
-\Keywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, \proglang{R}}
-\Plainkeywords{Cramer von Mises, goodness-of-fit tests, Kolmogorov-Smirnov, R} %% without formatting
-%% at least one keyword must be supplied
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{13}
-%% \Issue{9}
-%% \Month{September}
-%% \Year{2004}
-%% \Submitdate{2004-09-29}
-%% \Acceptdate{2004-09-29}
-
-%% The address of (at least) one author should be given
-%% in the following format:
-\Address{
-  Taylor B. Arnold \\
-  24 Hillhouse Ave. \\
-  New Haven, CT 06511, USA\\
-  E-mail: \email{taylor.arnold@yale.edu}\\
-  URL: \url{http://euler.stat.yale.edu/~tba3}\\
-
-  John W. Emerson \\
-  24 Hillhouse Ave. \\
-  New Haven, CT 06511, USA\\
-  E-mail: \email{john.emerson@yale.edu}\\
-  URL: \url{http://euler.stat.yale.edu/~jay}
-}
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/1/31336-5053
-%% Fax: +43/1/31336-734
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\setkeys{Gin}{width=1.0\textwidth, height=1.0\textheight}
-
-%% include your article here, just as usual
-%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
 
 \section{Introduction}
-
-Example code:
-\begin{Schunk}
-\begin{Sinput}
-> 1 + 1
-\end{Sinput}
-\begin{Soutput}
-[1] 2
-\end{Soutput}
-\end{Schunk}
 
@@
 factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
-Smirnov, followed closely by several variants on an estimation procedure proposed by Cramer von Mises.
+Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r-von Mises.
 
@@
 in the small sample setting. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this void for users
-operating in the \proglang{R} environment.
+operating in the R environment.
 
@@
 Of all the methods for nonparametric goodness-of-fit tests, by far the most popular is the method devised by Kolmogorov and Smirnov.
-It is the only such test built into the base of \proglang{R}. The idea behind it is fairly simple. Given the cumulative distribution
+It is the only such test built into the base of R. The idea behind it is fairly simple. Given the cumulative distribution
 function $F_0(x)$ of the continuous null distribution, and the empirical distribution function $F_{data}(x)$ of the observed data, 
@@
 
-\begin{align}
-\sup_x \left| f(x)- g(x) \right| &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
-&=  \max_i \left( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \right)
-\end{align}
+\begin{align*}
+\sup_x &\left| f(x)- g(x) \right|   \\
+       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
+&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg)
+\end{align*}
 
@@
 is also a step function, we could replace the formula for some small $\epsilon$ by:
-\begin{align}
-\sup_x \left| f(x)- g(x) \right| &=  \max_i \left( \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \right)
-\end{align}
+\begin{align*}
+\sup_x &\left| f(x)- g(x) \right| =   \\
+        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
+\end{align*}
 Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function. In the case of not knowing,
@@
 %in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
-%the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the \proglang{R} language, or in any other open source options, did not previously
+%the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the R language, or in any other open source options, did not previously
 %exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
@@
 
-*** Check Cramer-von Mises for hyphen, and let's get the accent done properly.
+*** Check Cram\'{e}r-von Mises for hyphen, and let's get the accent done properly.
 
-\section{Cramer von Mises Tests}
+\section{Cram\'{e}r-von Mises Tests}
 
@@
 tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
-Harald Cramer and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted by \cite{anderson1952}, and 
+Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted by \cite{anderson1952}, and 
 \cite{Watson1961}. The test statistics are, respectively, given as:
@@
 W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
-A^2 &= n \cdot \int_{-\infty}^{\infty} \left[F_0(x) -F_0(x)^2 \right]^{-1} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
+A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \\
 U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
@@
 %Discuss the discrete version. How to calculate W2, U2, and A2. 
-Generalizations of the Cramer-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
+Generalizations of the Cram\'{e}r-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
 the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
@@
 magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
-elsewhere are quite prominent. Figure ? shows the non-monotonicity of the function as the test statistic grows.
+elsewhere are quite prominent. Figure \ref{cvmissues} shows the non-monotonicity of the function as the test statistic grows.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.4]{fig1.pdf}
+\end{center}
+\caption{Plot of p-value for given test statistics using numerical integration (pink) versus a conservative
+         chi-squared bound (blue). }
+\label{cvmissues}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Kolmogorov-Smirnov and Cramer von Mises Tests in R}
+\section{Kolmogorov-Smirnov and Cram\'{e}r-von Mises Tests in R}
 
 Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
-\proglang{R}'s Kolmogorov-Smirnov function \code{ks.test()} from base
+R's Kolmogorov-Smirnov function \code{ks.test()} from base
 package \pkg{stats}.
 Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
-function for Cramer von Mises tests.  Both are available from the authors
+function for Cram\'{e}r-von Mises tests.  Both are available from the authors
 as of October, 2010,
@@
 null distributions by allowing the second argument, \code{y}, to be
-an empirical cumulative distribution function (an \proglang{R} function
+an empirical cumulative distribution function (an R function
 with class \code{ecdf}) or an object of class \code{stepfun} specifying
@@
 arguments specify the data and null distribution; the only extra option,
-\code{type}, specifies the variant of the Cramer-von Mises test:
+\code{type}, specifies the variant of the Cram\'{e}r-von Mises test:
 \begin{itemize}
@@
 the null model
-\item \code{type}: the variant of the Cramer-von Mises test; \code{W2}
+\item \code{type}: the variant of the Cram\'{e}r-von Mises test; \code{W2}
 is the default and most common method, \code{U2} is for cyclical data,
@@
 
-\begin{Schunk}
-\begin{Sinput}
-set.seed(1)
-REPS <- 100000
-asymp.pval <- rep(0, REPS)
-exact.pval <- rep(0, REPS)
-for (i in 1:REPS) {
-  x <- sample(1:10, 15, replace=TRUE)
-  asymp.pval[i] <- stats::ks.test(x, ecdf(1:10), alternative="g")$p.value
-  exact.pval[i] <- ks.test::ks.test(x, ecdf(1:10), alternative="g")$p.value
-}
-sum(asymp.pval<exact.pval)/REPS
-sum(asymp.pval<=0.05)/REPS
-sum(exact.pval<=0.05)/REPS
-sqrt(0.05*0.95/REPS)
-\end{Sinput}
-\end{Schunk}
-
-\begin{verbatim}
-#####################################################
-# 100000:
-
-> sum(asymp.pval<exact.pval)/REPS
-[1] 0
-> sum(asymp.pval<=0.05)/REPS
-[1] 0.01848
-> sum(exact.pval<=0.05)/REPS
-[1] 0.03763
-> 
-> sqrt(0.05*0.95/REPS)
-[1] 0.0006892024
-> 
-> REPS
-[1] 1e+05
-\end{verbatim}
-
 JAY: simulation showing original ks.test() was conservative, Conover exact.
@@
 
-Benefits and drawbacks of using simulation to gather p-values. (*)
+The computational capabilities of modern computers provide an alternative to using a complex formula to
+calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
+cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
+This can be quite useful in some cases, but using the exact formulas from our methodology has several
+distinct advantages. First of all, if the function is called many times (say, within another simulation
+study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
+for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
+similar threshold). Additionally, while
+simulations are generally a well received accepted method, it is often important in applied data analysis
+to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
+run of a simulation.
+
+In the end, for methods such as those presented in this paper where it is possible
+there are enough positives to suggest that we attempt to implement exact $p$-value calculations. This does lead
+to an interesting direction for future research in non-parametric goodness of fit tests: the methods presented
+here were historically chosen because they could be shown to have fixed null distributions which could easily
+calculated in an era without fast computers for carrying out permutation tests. It is quite possible that
+variants without the property, which could be easily used today, have greater power to certain alternative 
+distributions.
 
@@
 
-*** Bibliography needs cleaning before we finish up, just don't forget.
+%\bibliography{example}
 
-%\bibliographystyle{plain}
-\bibliography{jss}
+\bibliography{paper}
 
-\end{document}
+\address{
+  Taylor B. Arnold \\
+  24 Hillhouse Ave. \\
+  New Haven, CT 06511
+  USA\\
+}
+\email{taylor.arnold@yale.edu}
+
+\address{
+  John W. Emerson \\
+  24 Hillhouse Ave. \\
+  New Haven, CT 06511
+  USA\\
+}
+\email{john.emerson@yale.edu}

### 2010-11-05 rjournal/codeforgraphics.R

@@ -1,0 +1,60 @@
+#setwd('/home/tba3/Desktop/today/ks-test/JSS')
+
+  cvm.pval <- function(STAT, lambda, good=TRUE) {
+
+    x <- STAT
+
+    theta <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+         VAL <- VAL + 0.5*atan(lambda[i]*u)
+      }
+      return(VAL - 0.5*x*u)
+    }
+
+    rho <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+        VAL <- VAL + log(1 + lambda[i]^2*u^2)     
+      }
+      VAL <- exp(VAL*0.25)
+      return(VAL)
+    }
+
+    fun <- function(u) return(sin(theta(u))/(u*rho(u)))
+
+    pval <- 0
+    try(pval <- 0.5 + integrate(fun, 0, Inf, subdivisions=1e6)$value/pi,
+        silent=TRUE)
+    if(pval > 0.001 || good == FALSE) return(pval)
+    if(pval <= 0.001 & good == TRUE) {
+      df <- sum(lambda != 0)
+      return(dchisq(STAT/max(lambda),df))
+    }
+  } # End cvm.pval()
+
+#lambda <- sort(runif(10))
+lambda <- sort(c(0.020837 , 0.342575 , 0.531838 , 0.604825 , 0.645453 , 0.706874 , 0.750545 , 0.776914 , 0.810034 , 0.880271))
+cvm.pval(1, lambda)
+
+I <- seq(0,35,length.out=1000)
+y <- rep(NA, length(I))
+for(i in 1:length(I)) {
+  y[i] <- cvm.pval(I[i], lambda, good=TRUE)
+}
+z <- rep(NA, length(I))
+for(i in 1:length(I)) {
+  z[i] <- cvm.pval(I[i], lambda, good=FALSE)
+}
+
+pdf('fig1.pdf', height=6, width=8)
+plot(I[I > 18], z[I > 18], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
+abline(h=0, lty=2)
+lines(I[I > 21], y[I > 21], lty=2, col='blue', lwd=2)
+lines(I[I > 18], z[I > 18], col='salmon', lwd=2)
+dev.off()
+
+
+
+
+

### 2010-11-07 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@21 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
-Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r-von Mises.
+Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.
 
@@
 This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
-apply the functions meant for continuous distributions. As shown in section 5, the latter can be particularly dangerous
-in the small sample setting. The packages \pkg{ks.test} and \pkg{cvm.test} serve to fill this void for users
+apply the functions meant for continuous distributions. As we will see, the latter can be particularly dangerous
+in the small sample setting. The functions \pkg{ks.test} and \pkg{cvm.test} presented here serve to fill this void for users
 operating in the R environment.
-
-** Ditto on wordsmithing the contribution (package/function, etc...).
 
@@
 
-Of all the methods for nonparametric goodness-of-fit tests, by far the most popular is the method devised by Kolmogorov and Smirnov.
+Of all the methods for nonparametric goodness-of-fit tests, the most popular is the method devised by Kolmogorov and Smirnov.
 It is the only such test built into the base of R. The idea behind it is fairly simple. Given the cumulative distribution
@@
 alternatives of the above test statistic exist. The absolute value is discarded and the rest is either left alone (the `greater' testing 
-alternative) or the supremum is replaces with a negative infimum (the `lesser' hypothesis alternative). These can be helpful depending 
+alternative) or the supremum is replaced with a negative infimum (the `lesser' hypothesis alternative). These can be helpful depending 
 on the nature of the alternative hypotheses for which the test is desired to be powerful towards.
@@
 \end{align*}
-Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function. In the case of not knowing,
+Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function and knowing something
+about its particular break-points. In the case of not knowing,
 or having a $g$ which is in neither a step function nor a continuous function, the only method for computing the supremum is to take the numerical limit in (1). 
@@
 
-Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, the user must either indicate the points of discontinuity of the null
-distribution's cdf or live with a much smaller.
+Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, we have forced the user to indicate the points of discontinuity
+of the null distribution's cumulative distribution function. It is often the case that the test is used inside of a long simulation, 
+and the added computational time would likely be prohibitive.
 
-%Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
-%continuous case. The default options for \pkg{ks.test} already allow the user to specify if they would like approximate asymptotic distributions (since these exist
-%in the continuous setting as well). So when this approximation flag is turned on, this distribution is used for all situations. When the user requires exact p-values,
-%the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the R language, or in any other open source options, did not previously
-%exist and is included in the revised \pkg{ks.test}. The calculations are complex but straightforward; the full details are contained in the package source and in 
-%the original Conover paper.
+Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
+continuous case. When the user requires exact p-values,
+the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the R language, or in any other open source options, did not previously
+exist and is included in the new function \pkg{ks.test}. The calculations are complex but straightforward; 
+the full details are contained in the package source and in the original Conover paper.
 
 %Numerical issues arise from the summation; need to do something else for high $n$. Simulation? Asymptotic approximation? 
-%The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that arise when calculated the p-values for larger sample sizes.
-
-*** Check Cram\'{e}r-von Mises for hyphen, and let's get the accent done properly.
+%The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that 
+%arise when calculated the p-values for larger sample sizes.
 
@@
 
-*** New paragraph, perhaps here, on the differences between these three tests.
+The relative powers of these tests to different alternatives are studied in depth in \cite{stephens1974}. In general, the $W^2$
+statistic is used unless there is a good reason not to. The $A^2$ statistic was developed by Anderson in order to be
+generalized to the two-sample case, but was shown to perform very similarly to the original statistic in the one-sample
+setting. Watson's $U^2$ statistic was developed for distributions which are cyclically distributed; that is they have some
+order to them but no natural starting point. A common example would be the months of the year. His statistic will be same
+if the data are cyclically reordered, since giving preference to one ordering would be unnatural.
 
@@
 chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation (*** NOT FINISHED?)
+the original papers shows that this approximation is conservative and asymptotically equivalent to the true null distribution.
 
@@
 magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
-elsewhere are quite prominent. Figure \ref{cvmissues} shows the non-monotonicity of the function as the test statistic grows.
+elsewhere are quite prominent.
 
@@
 We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
-of a higher degree of freedom.
+of a higher degree of freedom. While for higher p-values it is better to use the original formulation, for small
+p-values (we picked a cut-off of 0.001) this correction is useful and, given the small values, it should not
+greatly affect the interpretation of the results.
+
+Figure \ref{cvmissues} shows the non-monotonicity of the function as the test statistic grows, as well as how the
+conservative approximation performs. While it is slightly conservative, it has the nice property of being strictly
+monotone and not exhibiting other odd noisy behavior.
 
@@
 When \code{exact = FALSE} the asymptotic distribution is used which is known
-to be imprecise but conservative, even for small sample (CITATION).
+to be imprecise but conservative, even for small samples (CITATION).
 
@@
 
-Go into more detail about reasons for using these versus chi-squared type tests. (*)
-
-Issues revolving around incorrect usage of default ks.test() in R because of continuity error (i.e. ks.test(1:2, ecdf(1:2)).  Will be earlier, needed here?
-
-Further discuss numerical issues; ks -> n over 30 has problems, CvM -> integrate function can sometimes refuse to run because of failure to calculate integral.
-Earlier, needed here?
+The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
+It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
+number of data points observed at each value and the expected number of data points that should be observed at each point. 
+This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
+at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
+have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
+test is generally not used unless there are at least five data points in each element of the support. On the other hand,
+chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
+(e.g. a set of ethnicities).
+For a more complete comparison of the two tests' power see \cite{slakter65}.
 
@@
 
-In the end, for methods such as those presented in this paper where it is possible
+In the end, for methods such as those presented in this paper where it is possible,
 there are enough positives to suggest that we attempt to implement exact $p$-value calculations. This does lead
@@
 
-Possibly go into a brief discussion of classes and class inheritance in R.
+%Possibly go into a brief discussion of classes and class inheritance in R.
 The use of `htest' in our functions; benefits and issues with using it.
@@
 More precisely, explain that we
-could provide upper and lower bounds for the p-value but are limited by class
-\code{htest}.
+%could provide upper and lower bounds for the p-value but are limited by class
+%\code{htest}.
 
-Two-sample cases; no theory for discrete distributions. Why? Distribution dependence messes things up. Discuss this for sure. (*)
+In the continuous setting, both the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two sample analogues. Here data are
+observed from two processes, and the hypothesis tested is whether they came from the same (but unspecified) distribution. There
+does not exist an analogous theory for discrete distributions. This comes from the fact that the discrete null distributions
+of the test statistics depend on the exact null distribution; therefore the two sample case would surely have to depend
+on the exact distributions used as well, which are generally not even stated in the two-sample case.
 
-For further generalizations of tests see \cite{dewev1973}
-
-Future work: continuous CVM.
+While we have implemented the two most popular variants of goodness-of-fit tests, there are 
+several more exotic varieties to be found. For further generalizations of tests see the extended study done in \cite{dewev1973}.
 
@@
 
-\address{
-  Taylor B. Arnold \\
+\address{Taylor B. Arnold \\
+Yale University\\
 24 Hillhouse Ave. \\
@@
 
-\address{
-  John W. Emerson \\
+\address{John W. Emerson \\
+Yale University\\
 24 Hillhouse Ave. \\

### 2010-11-07 pkg/cvm.test/R/cvm.test.R

@@ -1,6 +1,27 @@
 cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
 
-  cvm.pval <- function(STAT, lambda) {
+  cvm.stat <- function(x,y,type) {
+    fdata <- ecdf(x)
+    if(type != 'A2') {
+      igrand <- function(z) return((fdata(z) - y(z))^2)
+    } else {
+      igrand <- function(z) return((fdata(z) - y(z))^2 / (y(z) - y(z)^2))
+    }
+    stat <- integrate(igrand, -Inf, Inf)[1]
+    if(type != 'U2') {
+      return(stat)
+    } else {
+      igrand <- function(z) return((fdata(z) - y(z)-stat)^2) 
+      return(integrate(igrand, -Inf, Inf)[1])
+    }    
+  }
+
+  cvm.pval <- function(stat) {
+
+    return()
+  }  
+
+  cvm.pval.disc <- function(STAT, lambda) {
 
     x <- STAT
 
@@ -33,7 +54,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
     }
   } # End cvm.pval()
 
-  cvm.stat <- function(x,y, type=c("W2", "U2", "A2")) {
+  cvm.stat.disc <- function(x,y, type=c("W2", "U2", "A2")) {
     type <- match.arg(type)
     I <- knots(y)
     N <- length(x)
@@ -74,17 +95,32 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
 
   type <- match.arg(type)
   DNAME <- deparse(substitute(x))
-  if(length(setdiff(x, knots(y))) != 0) {
-    stop("Data are incompatable with null distribution")
+  if(is.stepfun(y)) {
+    if(length(setdiff(x, knots(y))) != 0) {
+      stop("Data are incompatable with null distribution")
+    }
+    tempout <- cvm.stat.disc(x,y,type=type)
+    STAT <- tempout[1]
+    lambda <- tempout[2:length(tempout)]
+    PVAL <- cvm.pval.disc(STAT, lambda)
+    METHOD <- paste("Cramer-von Mises -", type)
+    names(STAT) <- as.character(type)
+    RVAL <- list(statistic = STAT, p.value = PVAL, alternative = "Two.sided",
+                 method = METHOD, data.name=DNAME)
+  } else {
+    y <- y[!is.na(y)]
[…]

### 2010-11-09 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@22 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \abstract{
-   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
-   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
-   to users. The packages ks.test and cvm.test serve to fill this need in the R language for
-   the two most popular nonparametric tests. The structure of these two packages are explained and examples
-   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
-   in their implementation.
+A general theory for extending nonparametric goodness-of-fit tests 
+to discrete null distributions has existed
+for several decades. Despite this, modern statistical software 
+has generally failed to provide this methodology
+to users. We offer a revision of R's \code{ks.test()} function
+and a new \code{cvm.test()} function that serve to fill this need
+in the R language for two of the most popular nonparametric tests. 
+This paper describes the the structure of these contributions and
+provides examples of their usage. Particular attention is given
+to various numerical issues that arise in their implementation.
 }
@@
 Given a particular null distribution, goodness-of-fit tests 
-are used in order to test whether it is likely that the data were generated via the null distribution.
+are used in order to test whether it is likely that the data were
+generated via the null distribution.
 
 %Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
-While almost any hypothesis test can be viewed as a variant on a goodness-of-fit test, the term is typically
-applied to those tests which are nonparametric in nature. That is, they do not conduct the hypothesis testing 
+While almost any hypothesis test can be viewed as a variant on a
+goodness-of-fit test, the term ``goodness-of-fit test'' is typically
+reserved for those tests which are nonparametric in nature. 
+That is, they do not conduct the hypothesis testing 
 through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
@@
 
-Hmm... I thought the package had a trivial example where the standard
-\code{ks.test()} was a problem?
-\begin{Schunk}
-\begin{Sinput}
-> library(ks.test)
-> x <- sample(1:10, 25, replace = TRUE)
-> x
-\end{Sinput}
-\begin{Soutput}
- [1]  3  4  6 10  3  9 10  7  7  1  3  2  7  4  8  5  8 10
-[19]  4  8 10  3  7  2  3
-\end{Soutput}
-\begin{Sinput}
-> ks.test(x, ecdf(1:10))
-\end{Sinput}
-\begin{Soutput}
-	One-sample Kolmogorov-Smirnov test
-
-data:  x 
-D = 0.08, p-value = 1
-alternative hypothesis: two-sided 
-\end{Soutput}
-\begin{Sinput}
-> ks.test(x, ecdf(1:10), alternative = "g")
-\end{Sinput}
-\begin{Soutput}
-	One-sample Kolmogorov-Smirnov test
-
-data:  x 
-D^+ = 0.04, p-value = 0.7731
-alternative hypothesis: the CDF of x lies above the null hypothesis 
-\end{Soutput}
-\begin{Sinput}
-> ks.test(x, ecdf(1:10), alternative = "g", exact = FALSE)
-\end{Sinput}
-\begin{Soutput}
-	One-sample Kolmogorov-Smirnov test
-
-data:  x 
-D^+ = 0.04, p-value = 0.9231
-alternative hypothesis: the CDF of x lies above the null hypothesis 
-\end{Soutput}
-\end{Schunk}
-
-Discuss this, nice example.  Relate to new material that is about to go into
-the earlier section.
+Consider a toy example, with observed data of length 2 (specifically, the
+values 0 and 1) and a hypothized null distribution that places equal
+probability on the values 0 and 1.  With the current \code{ks.test()} function
+in R (which, admittedly, doesn't claim to handle discrete distributions),
+the reported p-value, 0.5, is clearly incorrect:
 \begin{Schunk}
@@
 \end{Soutput}
+\end{Schunk}
+Instead, the value of $D$ given in equation (1)
+should be 0 and the associated p-value should be 1.  Our revision of \code{ks.test()}
+fixes this problem when the user provides a discrete distribution:
+\begin{Schunk} 
 \begin{Sinput}
@@
 
-JAY: simulation showing original ks.test() was conservative, Conover exact.
+Next, we simulate a sample of 25 from the discrete uniform distribution on
+the integers $\{1, 2, \ldots, 10\}$ and show several variants of the
+new \code{ks.test()} implementation.  The first is the default two-sided test,
+where the reported p-value is a conservative upper bound for the actual
+p-value.  In this case, the approximation may not be that tight, but this is
+irrelevant for such large p-values (for more interesting p-values, the
+upper bound is very close to the true p-value).
+\begin{Schunk}
+\begin{Sinput}
+> library(ks.test)
+> set.seed(1)
+> x <- sample(1:10, 25, replace = TRUE)
+> x
+\end{Sinput}
+\begin{Soutput}
+ [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
+[14]  4  8  5  8 10  4  8 10  3  7  2  3
+\end{Soutput}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10))
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D = 0.08, p-value = 1
+alternative hypothesis: two-sided 
+\end{Soutput}
+\end{Schunk}
+Next, we conduct the default one-sided test, where Conover's method
+provides the exact p-value (up to the numerical precision of the
+implementation):
+\begin{Schunk}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10), alternative = "g")
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D^+ = 0.04, p-value = 0.7731
+alternative hypothesis:
+the CDF of x lies above the null hypothesis 
+\end{Soutput}
+\end{Schunk}
+In contrast, the option \code{exact=FALSE} results in the
+p-value obtained by applying the classical Kolmogorov-Smirnov
+test, resulting in a conservative p-value:
+\begin{Schunk}
+\begin{Sinput}
+> ks.test(x, ecdf(1:10), alternative = "g", 
++         exact = FALSE)
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D^+ = 0.04, p-value = 0.9231
+alternative hypothesis:
+the CDF of x lies above the null hypothesis 
+\end{Soutput}
+\end{Schunk}
+
+JAY: reconsider simulation.  Not really needed, if we have citations for
+the conservatism points, etc...?  The code for this is in the JSS folder.
+if we want to revisit it.
+
+Finally, we employ two of the Cram\'{e}r-von Mises tests.  Taylor, are these
+your ``cynical example'' that you commented on, or did you have something
+in addition to these?  What should be discussed here?
 

### 2010-11-09 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@23 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 and a new \code{cvm.test()} function that serve to fill this need
-in the R language for two of the most popular nonparametric tests. 
-This paper describes the the structure of these contributions and
+in the R language for two of the most popular nonparametric
+goodness-of-fit tests. This paper describes these contributions and
 provides examples of their usage. Particular attention is given
@@
 %the underlying process from which it came. 
-Given a particular null distribution, goodness-of-fit tests 
-are used in order to test whether it is likely that the data were
-generated via the null distribution.
+Goodness-of-fit tests are used to assess whether data are consistent
+with a hypothesized null distribution.  The $\chi^2$ test the best-known
+parametric goodness-of-fit test, while the most popular nonparametric tests
+are the classic test proposed by Kolmogorov and Smirnov followed closely by
+several variants on an estimation procedure proposed by
+Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
+tests, or simply Cram\'{e}r-von Mises tests).
+
+In their most basic form, the nonparametric goodness-of-fit
+tests are intended for continuous null distributions, but they have
+also been adapted for use with discrete null distributions.  Unfortunately,
+most modern statistical software has failed to incorporate these discrete
+versions, leaving researchers with the $\chi^2$ test or in the unfortunate
+position of applying a nonparametric test designed for continuous null
+distributions and hoping for the best.
+As we will see, the latter choice can be particularly dangerous
+in the small sample setting. This paper presents a revision of R's
+\code{ks.test()} function and a new \code{cvm.test()} function to
+fill this void for researchers and practitioners in the R environment.
+
+%Given a particular null distribution, goodness-of-fit tests 
+%are used in order to test whether it is likely that the data were
+%generated via the null distribution.
 
 %Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
-While almost any hypothesis test can be viewed as a variant on a
-goodness-of-fit test, the term ``goodness-of-fit test'' is typically
-reserved for those tests which are nonparametric in nature. 
-That is, they do not conduct the hypothesis testing 
-through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
-calculate differences between the empirical distribution of the observed data and the overall distribution 
-of the null model. In many cases these tests are preferential since they tend to have an increased power for 
-interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
-factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
-Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.
+%While almost any hypothesis test can be viewed as a variant on a
+%goodness-of-fit test, the term ``goodness-of-fit test'' is typically
+%reserved for those tests which are nonparametric in nature. 
+%That is, they do not conduct the hypothesis testing 
+%through the explicit calculation of the parameters generating the data. Instead, they %generally attempt to
+%calculate differences between the empirical distribution of the observed data and the overall distribution 
+%of the null model. In many cases these tests are preferential since they tend to have an increased power for 
+%interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
+%factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
+%Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.
 
 % Discrete versions exist; just no code
-While the original aim of nonparametric goodness of fit tests was meant for continuous null distributions, discrete versions
-have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
-This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
-apply the functions meant for continuous distributions. As we will see, the latter can be particularly dangerous
-in the small sample setting. The functions \pkg{ks.test} and \pkg{cvm.test} presented here serve to fill this void for users
-operating in the R environment.
+%While the original aim of nonparametric goodness of fit tests was meant for continuous null distributions, discrete versions
+%have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
+%This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
+%apply the functions meant for continuous distributions.
 
@@
 
-\begin{align*}
+\begin{align}
 \sup_x &\left| f(x)- g(x) \right|   \\
@@
 &=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg)
-\end{align*}
+\end{align}
 
@@
 is also a step function, we could replace the formula for some small $\epsilon$ by:
-\begin{align*}
+\begin{align}
 \sup_x &\left| f(x)- g(x) \right| =   \\
         &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
-\end{align*}
+\end{align}
 Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function and knowing something
@@
 \cite{Watson1961}. The test statistics are, respectively, given as:
-\begin{align*}
+\begin{align}
 W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
@@
 U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
-\end{align*}
+\end{align}
 Where $F$ is either the cumulative distribution of the null model or the empirical cumulative distribution of the observed data.
@@
 sum of chi squares, is:
-\begin{align*}
+\begin{align}
 \mathbb{P}\{Q \geq x \} = \frac{1}{2} + \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
-\end{align*}
+\end{align}
 For continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ which depend on the actual weights used.
@@
 following inequality:
-\begin{align*}
+\begin{align}
 \mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
 &= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
-\end{align*}
+\end{align}
 We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
@@
 \end{center}
-\caption{Plot of p-value for given test statistics using numerical integration (pink) versus a conservative
-         chi-squared bound (blue). }
+\caption{Plot of p-value for given test statistics using numerical integration
+(pink) versus a conservative chi-squared bound (blue). }
 \label{cvmissues}
@@
 
-Package \pkg{ks.test} contains function \code{ks.test()}, a revision of
+Functions \code{ks.test()} and \code{cvm.test()} are provided for
+convenience in packages \pkg{ks.test} and \pkg{cvm.test}, respectively.
+Function \code{ks.test()} offers a revision of
 R's Kolmogorov-Smirnov function \code{ks.test()} from base
-package \pkg{stats}.
-Package \pkg{cvm.test} contains function \code{cvm.test()}, a new
-function for Cram\'{e}r-von Mises tests.  Both are available from the authors
-as of October, 2010,
+package \pkg{stats}, while \code{cvm.test()} is a new
+function for Cram\'{e}r-von Mises tests.
+Both are available from the authors
 or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});

### 2010-11-09 rjournal/codeforgraphics.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@24 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,5 +1,44 @@
 #setwd('/home/tba3/Desktop/today/ks-test/JSS')
 
+  cvm.stat.disc <- function(x,y, type=c("W2", "U2", "A2")) {
+    type <- match.arg(type)
+    I <- knots(y)
+    N <- length(x)
+    e <- diff(c(0,N*y(I)))
+    obs <- rep(0, length(I))
+    for(j in 1:length(I)) {
+      obs[j] <- length(which(x == I[j]))
+    }
+    S <- cumsum(obs)
+    T <- cumsum(e)
+    H <- T/N
+    p <- e/N
+    t <- (p + p[c(2:length(p), 1)])/2
+    Z <- S - T
+    Zbar <- sum(Z*t)
+
+    S0 <- diag(p) - p %*% t(p)
+    A <- matrix(1, length(p), length(p))
+    A <- apply(row(A) >= col(A),2, as.numeric)
+    E <- diag(t)
+    One <- rep(1, nrow(E))
+    K <- diag(0, length(H))
+    diag(K)[-length(H)] <- 1/(H[-length(H)]*(1-H[-length(H)]))
+    Sy <- A %*% S0 %*% t(A)
+    M <- switch(type, W2 = E,
+                U2 = (diag(1, nrow(E)) - 
+                      E%*%One%*%t(One))%*%E%*%(diag(1, nrow(E)) -
+                      One%*%t(One)%*%E),
+                A2 = E%*%K)
+    lambda <- eigen(M%*%Sy)$values
+
+    STAT <- switch(type, W2 = sum(Z^2*t)/N, U2 = sum((Z-Zbar)^2*t)/N,
+                   A2 = sum((Z^2*t/(H*(1-H) ))[-length(I)])/N)
+
+    return(c(STAT, lambda))
+  } # End cvm.stat()
+
+
   cvm.pval <- function(STAT, lambda, good=TRUE) {
 
     x <- STAT
@@ -34,10 +73,13 @@
   } # End cvm.pval()
 
 #lambda <- sort(runif(10))
-lambda <- sort(c(0.020837 , 0.342575 , 0.531838 , 0.604825 , 0.645453 , 0.706874 , 0.750545 , 0.776914 , 0.810034 , 0.880271))
[…]

### 2010-11-10 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@26 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \abstract{
-A general theory for extending nonparametric goodness-of-fit tests
+Methodology extending nonparametric goodness-of-fit tests
 to discrete null distributions has existed
@@
 
-In their most basic form, the nonparametric goodness-of-fit
+*** Citations in the paragraph above?  Also check the body of paper. ***
+
+In their most basic forms, the nonparametric goodness-of-fit
 tests are intended for continuous null distributions, but they have
 also been adapted for use with discrete null distributions.  Unfortunately,
-most modern statistical software has failed to incorporate these discrete
+most modern statistical software have failed to incorporate these discrete
 versions, leaving researchers with the $\chi^2$ test or in the unfortunate
@@
 As we will see, the latter choice can be particularly dangerous
-in the small sample setting. This paper presents a revision of R's
+with small sample sizes. This paper presents a revision of R's
 \code{ks.test()} function and a new \code{cvm.test()} function to
@@
 \end{Schunk}
+A different toy example shows the dangers of using R's existing
+\code{ks.test()} function:
+\begin{Schunk}
+\begin{Sinput}
+> ks.test(rep(1, 3), ecdf(1:3))
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
 
-JAY: reconsider simulation.  Not really needed, if we have citations for
-the conservatism points, etc...?  The code for this is in the JSS folder.
-if we want to revisit it.
+data:  rep(1, 3) 
+D = 0.6667, p-value = 0.04938
+alternative hypothesis: two-sided 
+\end{Soutput}
+\end{Schunk}
+If, instead, either \code{exact=FALSE} is used with the new
+\code{ks.test()} function, or if the original \code{stats::ks.test()}
+is used, the reported p-value is 0.1389.
 
-Finally, we employ two of the Cram\'{e}r-von Mises tests.  Taylor, are these
-your ``cynical example'' that you commented on, or did you have something
-in addition to these?  What should be discussed here?
-
+Finally, we employ two of the Cram\'{e}r-von Mises tests.  Need short discussion,
+and we could use a good cyclical example:
 \begin{Schunk}

### 2010-11-15 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@27 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Goodness-of-fit tests are used to assess whether data are consistent
-with a hypothesized null distribution.  The $\chi^2$ test the best-known
+with a hypothesized null distribution.  The $\chi^2$ test is the best-known
 parametric goodness-of-fit test, while the most popular nonparametric tests
@@
 
+*** why "estimation procedure" above?
 *** Citations in the paragraph above?  Also check the body of paper. ***
 
-In their most basic forms, the nonparametric goodness-of-fit
+In their most basic forms, these nonparametric goodness-of-fit
 tests are intended for continuous null distributions, but they have
@@
 versions, leaving researchers with the $\chi^2$ test or in the unfortunate
-position of applying a nonparametric test designed for continuous null
-distributions and hoping for the best.
-As we will see, the latter choice can be particularly dangerous
+position of applying a nonparametric test designed for a continuous null
+distribution and hoping for the best.
+As we will see, this latter course of action can be particularly dangerous
 with small sample sizes. This paper presents a revision of R's
@@
 fill this void for researchers and practitioners in the R environment.
+
+*** Later, make sure we point out how using one of the ``old'' tests and
+hoping for the best is a problem. ***
 
@@
 
-Of all the methods for nonparametric goodness-of-fit tests, the most popular is the method devised by Kolmogorov and Smirnov.
-It is the only such test built into the base of R. The idea behind it is fairly simple. Given the cumulative distribution
-function $F_0(x)$ of the continuous null distribution, and the empirical distribution function $F_{data}(x)$ of the observed data, 
-one constructs the statistic:
+The most popular nonparametric goodness-of-fit test is the
+Kolmogorov-Smirnov test. The idea is fairly simple:
+Given the cumulative distribution
+function $F_0(x)$ of the continuous null distribution
+and the empirical distribution function $F_{data}(x)$ of the observed data,  the test statistic is given by
 \begin{align}
-D = \sup_x \left| F_0(x)- F_{data}(x) \right|
+D = \sup_x \left| F_0(x)- F_{data}(x) \right|.
 \end{align}
-The distribution of $D$ under the null model does not depend on which null distribution is being used, making this a computationally
-attractive method. For a standard treatment of the test and its performance relative to other algorithms see \cite{slakter65}. Two common
-alternatives of the above test statistic exist. The absolute value is discarded and the rest is either left alone (the `greater' testing 
-alternative) or the supremum is replaced with a negative infimum (the `lesser' hypothesis alternative). These can be helpful depending 
-on the nature of the alternative hypotheses for which the test is desired to be powerful towards.
+The distribution of $D$ does not depend on the hypothesized
+distribution, making this a computationally
+attractive method. \cite{slakter65} offers a standard presentation
+of the test and its
+performance relative to other algorithms. 
+The test statistic (1) is easily adapted for one-sided tests.
+For these, the absolute value is discarded and the tests are based
+on either the supremum of the remaining difference (the `greater' testing
+alternative) or by replacing the supremum with a negative infimum
+(the `lesser' hypothesis alternative).  Tabulated p-values have been
+available for these tests since (*** year ****).
 
-The extension of this result to non-continuous null distributions does not have such a clean solution. The formula of
-the test statistic $D$ remains unchanged, however the distribution of the testing statistic is much more difficult. Unlike in the 
-continuous case, it ultimately depends on which null model was chosen; this makes it impossible to simply read p-values directly 
-off of a fixed table. It was known since at least the 1950's that using the tables for continuous distributions resulted in 
-conservative p-values; it was not until \citet{Conover1972} that a method for computing exact p-values in this case was developed.
+**** check this above, both do away with the absolute value, I think? ***
+
+The extension of the Kolmogorov-Smirnov test to non-continuous 
+null distributions does not have such a clean solution. The formula of
+the test statistic $D$ remains unchanged, but the distribution
+of the test statistic is much more difficult to obtain. Unlike the 
+continuous case, the null distribution depends on the null model.
+It has been known since at least the 1950's that using the tables associated
+with continuous null distributions results in conservative p-values
+(**** citation?) when the null distribution is discontinuous 
+(*** or should this be discrete? ***).  In the early 1970's, 
+\citet{Conover1972} developed the method implemented here
+for computing exact p-values
+in the case of discrete null distributions.
 
@@
 
-The implementation of the discrete Kolmogorov-Smirnov function consists of two parts. First the particular test statistic needs to be calculated,
-and then the p-value for that particular statistic must be computed.
+The implementation of the discrete Kolmogorov-Smirnov test is considered
+in two parts. First, the particular test statistic needs to be calculated
+(corresponding to the desired one-sided or two-sided test).
+Then, the p-value for that particular test statistic may be computed.
 
-Given that the test statistic is, theoretically, the same as in the
-continuous case it would seem that the first part could be directly taken from the existing procedures. This is, unfortunately, often not the case.
+The test statistic is, in terms of the theory, the same as in the
+continuous case; it would seem that no additional work would be required,
+but unfortunately this is not the case.
 Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
-is continuous. If we want to determine the supremum of the difference between these two functions notice that:
+is continuous (the classical Kolmogorov-Smirnov situation).
+In order to determine the supremum of the difference between
+these two functions, notice that
+\begin{align}
+\sup_x &\left| f(x)- g(x) \right|  \nonumber  \\
+       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
+&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg).
+\end{align}
+Computing the maximum over these $2N$ (*** why 2N? ***)
+values (with $f$ equal to
+$F_{data}(x)$ and $g$ equal to $F_0(x)$ as defined above) is clearly the 
+most efficient way to compute the Kolmogorov-Smirnov test statistic for
+a continuous null distribution. When the function $g$ is not
+continuous, however, equality (3) does not hold in general because 
+we cannot replace $\lim_{x\rightarrow x_e} g(x)$ with the value $g(x_i)$.
 
+If it is known that $g$ is a step function, it follows that
+for some small $\epsilon$,
 \begin{align}
-\sup_x &\left| f(x)- g(x) \right|   \\
-       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
-&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg)
-\end{align}
-
-Computing the maximum over these $2N$ values (with $f$ equal to $F_{data}(x)$ and $g$ equal to $F_0(x)$ as defined above) is clearly the 
-most efficient way to compute the Kolmogorov-Smirnov test statistic when given a continuous null distribution. When the function $g$ is not
-continuous, notice that this formula no longer works since in general we cannot replace the limit of $g$ with its value at $x_i$. If it is known that $g$
-is also a step function, we could replace the formula for some small $\epsilon$ by:
-\begin{align}
-\sup_x &\left| f(x)- g(x) \right| =   \\
+\sup_x &\left| f(x)- g(x) \right| =  \nonumber \\
         &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
 \end{align}
-Where the discontinuities in $g$ are at least a distance $\epsilon$ apart. But this requires knowing that $g$ is a step function and knowing something
-about its particular break-points. In the case of not knowing,
-or having a $g$ which is in neither a step function nor a continuous function, the only method for computing the supremum is to take the numerical limit in (1). 
-This clearly takes more computational time than simply cycling over $2N$ values.
+where the discontinuities in $g$ are at least a distance $\epsilon$ apart. 
+This, however, requires knowledge that $g$ is a step function as well as of
+the nature of its support (or break-points).
+%
+%*** Hmm... do we care about time? ***  When the break-points of $g$ are
+%unknown, or when $g$ is neither a step function nor a continuous distribution
+%function, computing the supremum requires taking the numerical limit in (1***?). 
+%This is computational involved and risks problems with numerical tolerances.
+%
+We implement the Kolmogorov-Smirnov test statistic for discrete null
+distributions by forcing the user to completely specify the null distribution.
 
-Therefore, in order to implement the discrete Kolmogorov-Smirnov test statistic, we have forced the user to indicate the points of discontinuity
-of the null distribution's cumulative distribution function. It is often the case that the test is used inside of a long simulation, 
-and the added computational time would likely be prohibitive.
+%**** Hmmm... It is often the case that the test is used inside of a
+%long simulation, 
+%and the added computational time would likely be prohibitive.
 
-Once the test statistic is determined, the p-value for this value needs to be computed. For high sample sizes, the test distribution becomes the same as in the
-continuous case. When the user requires exact p-values,
-the methodology in \citet{Conover1972} needs to be used. Code for this procedure in the R language, or in any other open source options, did not previously
-exist and is included in the new function \pkg{ks.test}. The calculations are complex but straightforward; 
-the full details are contained in the package source and in the original Conover paper.
+Having obtained the test statistic, the p-value must then be calculated. 
+For larger sample sizes, the null distribution coincides 
+(***should this be
+stated precisely as an asymptotic result???)
+with the standard null distribution (CITATION???). 
+When an exact p-value is required for
+smaller sample sizes, the methodology in \citet{Conover1972} is used. 
+Full details of the calculations are contained in source code
+of our revised function \code{ks.test()} and in Conover's original paper.
+
+** code hasn't previously existed?  I deleted this stuff, not sure if
+it fits here.  Is it obvious (since we bothered to do this in the first
+place)?  Or move it?
+
+%Code for this procedure in the R language, or in any other open source options, did not previously
+%exist and is included in the new function \pkg{ks.test}.
 
@@
 %distribution for the testing statistic.
-While the Kolmogorov-Smirnov test is the most well known of the non-parametric goodness of fit tests, there is another family of 
-tests which has been shown to be more powerful to a large class of alternatives distributions. The original was developed jointly by
-Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928}, and further adapted by \cite{anderson1952}, and 
-\cite{Watson1961}. The test statistics are, respectively, given as:
+
+While the Kolmogorov-Smirnov test may be the most popular of
+the nonparametric goodness-of-fit tests, there is another family of 
+tests which has been shown to be more powerful against a large class
+of alternatives hypotheses. 
+The original test was developed jointly (*** or was it simultaneously? ***) by
+Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
+The original test statistic, $W^2$, Anderson's $A^2$, and Watson's
+$U^2$ are:
 \begin{align}
@@
 \end{align}
-Where $F$ is either the cumulative distribution of the null model or the empirical cumulative distribution of the observed data.
-As in the Kolmogorov-Smirnov test statistic, these all have distribution free null distributions in the continuous case.
+As with the original Kolmogorov-Smirnov test statistic, these all have 
+null distributions of their test statistics which are independent of the
+specified null models.
 
-The relative powers of these tests to different alternatives are studied in depth in \cite{stephens1974}. In general, the $W^2$
-statistic is used unless there is a good reason not to. The $A^2$ statistic was developed by Anderson in order to be
-generalized to the two-sample case, but was shown to perform very similarly to the original statistic in the one-sample
-setting. Watson's $U^2$ statistic was developed for distributions which are cyclically distributed; that is they have some
-order to them but no natural starting point. A common example would be the months of the year. His statistic will be same
-if the data are cyclically reordered, since giving preference to one ordering would be unnatural.
+The relative powers of these tests to different alternatives are
+studied in depth in \cite{stephens1974}. In general, the $W^2$
+statistic is recommended unless there is a compelling reason to
+consider one of the others. The $A^2$ statistic was developed by
+Anderson in the process of generalizing the test for the two-sample case,
+but was shown to perform very similarly to the original statistic
+in the one-sample setting (*** so why, then would we ever use it??? ***).
+Watson's $U^2$ statistic was developed for distributions
+which are cyclic (with an ordering to the support but
+no natural starting point  *** how best to describe this? ***). 
+For example, a distribution on the months of the year could be considered
+cyclic. Watson's statistic is invariant to cyclical reordering (giving preference to one ordering would be unnatural).  **** is this last point necessary? ***
+
+*** Is a second paragraph on Stephen's results really necessary?  Can we
+pick and choose and get away with one paragraph? ***
 
@@
 %Discuss the discrete version. How to calculate W2, U2, and A2.
-Generalizations of the Cram\'{e}r-von Mises tests were developed in \cite{choulakian1994}. Much like for the Kolmogorov-Smirnov test,
-the theoretical form of the test statistics are unchanged; although the discreteness allows for a slightly simpler representation.
-The null distribution of the test statistics are again distribution dependent, unlike the continuous version. The methods do not
-suggest finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
-chi-squared variables (the weights depending on the particular distribution). This asymptotic distribution is what we implement here;
-the original papers shows that this approximation is conservative and asymptotically equivalent to the true null distribution.
+
+*** I edited, below, make sure you check it. ****
+
+Generalizations of the Cram\'{e}r-von Mises tests to discrete
+distributions were developed in \cite{choulakian1994}. As with the
+Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, although the discreteness allows for slightly simpler representations.
+The null distribution of the test statistics are again hypothesis-dependent.
+The methods do not offer finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
+chi-squared variables (with the weights depending on the particular
+null distribution). (*** is this true for all 3? ***)
+We implement this asymptotic distribution here;
+the original papers (*** really, all of them? ***)
+show that this approximation is conservative 
+and asymptotically equivalent to the true null distribution.
+
+*** What does it mean for a distribution to be equal to a weighted sum of
+independent chi-squared variables?  I think this needs to be stated more
+precisely, above.  A formal definition of $Q$ is needed someplace, too,
+as I'll probably note below. ***
 
@@
 
-Calculation of the three test statistics is done by straightforward matrix algebra as given in \cite{choulakian1994}. Determining 
-the form of the asymptotic null distribution is also easy using the built-in eigenvalue decomposition functions. The only 
-difficulty in the process involves actually calculating the percentiles for these weighted chi-squares.
+Calculation of the three test statistics is done using the
+matrix algebra given by \cite{choulakian1994}. The form of the
+asymptotic null distribution is obtained using R's
+eigenvalue decomposition functions.
 
-%Show how to compute the distribution of a weighted sum of chi squared's.
-The method used for calculating the distribution of a weighted sum of independent chi-squared variables is given in \cite{imhof1961}.
-A general method for computing any quadratic form of normals is presented, which is easily adapted for our case since
-each chi-squared variable has only one degree of freedom. The exact formula given for the distribution function of $Q$, the weighted
-sum of chi squares, is:
+The only notable difficulty in the process involves calculating the percentiles
+of the weighted sum of chi-squares,
 \begin{align}
-\mathbb{P}\{Q \geq x \} = \frac{1}{2} + \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
+Q = \sum_{i=1}^{p} \lambda_i \chi^2_{i,1df}
 \end{align}
-For continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ which depend on the actual weights used.
+\cite{imhof1961} provides a method for obtaining the distribution of
+a weighted sum of independent chi-squared variables.  It is
+easily adapted for our case because
+the chi-squared variables have only one degree of freedom.
+The exact formula given for the distribution function of $Q$
+is given by
+\begin{align}
+\mathbb{P}\{Q \geq x \} = \frac{1}{2} + 
+\frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
+\end{align}
+for continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ depending on the actual weights used.
 
-There is no analytic solution to the integral, but integration can be carried out using numerical techniques. This seems
-fine in most situations, but numerical issues do become a problem in the regime of large test statistics $x$. The function
-$\theta(\cdot, x)$ is linear in $x$, and thus as the the test statistic grows the corresponding period of the integrand 
-decreases. As the function acquires too many inflection points, the approximation becomes unstable. This is further
-magnified by this occurring when $p$-values should be very small; thus tiny fluctuations which would be undetectable
-elsewhere are quite prominent.
-
-There is fortunately a simple conservative approximation which can get around this numerical problem. Given the
+There is no analytic solution to the integral, so the integration is
+accomplished numerically. This seems fine in most situations we considered,
+but numerical issues appear in the regime of large test statistics $x$
+(or, equivalently, small p-values).
+The function $\theta(\cdot, x)$ is linear in $x$; as the test statistic 
+grows the corresponding periodicity of the integrand decreases and
+the approximation becomes unstable. 
+%This is further magnified in situations where the 
+%$p$-values should be very small, where tiny fluctuations (undetectable
+%elsewhere) become quite prominent. 
+We resolve this problem by using a simple conservative approximation to
+avoid the numerical instability. Consider the
 following inequality:
@@
 \end{align}
-We see that the values for the weighted sum can be bounded by a simple transformation and a chi-squared distribution
-of a higher degree of freedom. While for higher p-values it is better to use the original formulation, for small
-p-values (we picked a cut-off of 0.001) this correction is useful and, given the small values, it should not
-greatly affect the interpretation of the results.
+The values for the weighted sum can be bounded using a simple transformation
+and a chi-squared distribution of a higher degree of freedom. 
+The original formulation is preferable for most p-values, while
+this approximation is useful for smaller p-values (smaller than 0.001,
+based on our observations of the numerical instablity of the original
+formulaton); it should not greatly affect the interpretation of the results.
 
-Figure \ref{cvmissues} shows the non-monotonicity of the function as the test statistic grows, as well as how the
-conservative approximation performs. While it is slightly conservative, it has the nice property of being strictly
-monotone and not exhibiting other odd noisy behavior.
+*** How is the cutoff established in the code? ***
+
+Figure \ref{cvmissues} shows the non-monotonicity of the function 
+(*** which function, reference it? ***) as the test statistic grows, 
+compared to the conservative approximation.
+
+*** stuff deleted, redundant ***.
 
@@
 \caption{Plot of p-value for given test statistics using numerical integration
-(pink) versus a conservative chi-squared bound (blue). }
+(pink) versus a conservative chi-squared bound (dashed blue). }
 \label{cvmissues}
@@
 
-The revised \pkg{ks.test()} function supports one-sample tests for discrete
+The revised \code{ks.test()} function supports one-sample tests for discrete
 null distributions by allowing the second argument, \code{y}, to be
@@
 
-The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
-It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
-number of data points observed at each value and the expected number of data points that should be observed at each point. 
-This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
-at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
-have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
-test is generally not used unless there are at least five data points in each element of the support. On the other hand,
-chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
-(e.g. a set of ethnicities).
-For a more complete comparison of the two tests' power see \cite{slakter65}.
+*** I commented out a bunch of stuff.  I think the discussion can be short
+and sweet. ***
 
-The computational capabilities of modern computers provide an alternative to using a complex formula to
-calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
-cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
-This can be quite useful in some cases, but using the exact formulas from our methodology has several
-distinct advantages. First of all, if the function is called many times (say, within another simulation
-study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
-for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
-similar threshold). Additionally, while
-simulations are generally a well received accepted method, it is often important in applied data analysis
-to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
-run of a simulation.
+%The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
+%It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
+%number of data points observed at each value and the expected number of data points that should be observed at each point. 
+%This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
+%at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
+%have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
+%test is generally not used unless there are at least five data points in each element of the support. On the other hand,
+%chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
+%(e.g. a set of ethnicities).
+%For a more complete comparison of the two tests' power see \cite{slakter65}.
+
+%The computational capabilities of modern computers provide an alternative to using a complex formula to
+%calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
+%cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
+%This can be quite useful in some cases, but using the exact formulas from our methodology has several
+%distinct advantages. First of all, if the function is called many times (say, within another simulation
+%study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
+%for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
+%similar threshold). Additionally, while
+%simulations are generally a well received accepted method, it is often important in applied data analysis
+%to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
+%run of a simulation.
+
+*** Hmm... if you claim the following, you want to make sure it is true.
+Did you read it somewhere? ***
 
@@
 
-In the continuous setting, both the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two sample analogues. Here data are
-observed from two processes, and the hypothesis tested is whether they came from the same (but unspecified) distribution. There
-does not exist an analogous theory for discrete distributions. This comes from the fact that the discrete null distributions
-of the test statistics depend on the exact null distribution; therefore the two sample case would surely have to depend
-on the exact distributions used as well, which are generally not even stated in the two-sample case.
+In the continuous setting, both of 
+the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two-sample analogues. Here data are
+observed from two processes, and the hypothesis tested is 
+whether they came from the same (but unspecified) distribution. There
+does not exist an analogous theory for discrete distributions. 
+This comes from the fact that the discrete null distributions
+of the test statistics depend on the exact null distribution; 
+therefore the two-sample case would surely have to depend
+on the exact distributions used as well (*** is this true? ***), 
+which are generally not even stated in the two-sample case.
 
-While we have implemented the two most popular variants of goodness-of-fit tests, there are 
-several more exotic varieties to be found. For further generalizations of tests see the extended study done in \cite{dewev1973}.
+While we have implemented the two most popular variants of 
+goodness-of-fit tests, there are 
+several more exotic varieties to be found. For further 
+generalizations of tests see the extended study done in \cite{dewev1973}.
 

### 2010-11-17 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@28 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 are the classic test proposed by Kolmogorov and Smirnov followed closely by
-several variants on an estimation procedure proposed by
+several variants on a procedure proposed by
 Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
-tests, or simply Cram\'{e}r-von Mises tests).
+tests, or simply Cram\'{e}r-von Mises tests) .
 
-*** why "estimation procedure" above?
-*** Citations in the paragraph above?  Also check the body of paper. ***
+%*** why "estimation procedure" above?
+%*** Citations in the paragraph above?  Also check the body of paper. ***
 
@@
 
-*** Later, make sure we point out how using one of the ``old'' tests and
-hoping for the best is a problem. ***
+%*** Later, make sure we point out how using one of the ``old'' tests and
+%hoping for the best is a problem. ***
 
@@
 (the `lesser' hypothesis alternative).  Tabulated p-values have been
-available for these tests since (*** year ****).
+available for these tests since 1933. %(*** year ****).
 
-**** check this above, both do away with the absolute value, I think? ***
+%**** check this above, both do away with the absolute value, I think? ***
 
@@
 with continuous null distributions results in conservative p-values
-(**** citation?) when the null distribution is discontinuous 
-(*** or should this be discrete? ***).  In the early 1970's, 
+when the null distribution is discontinuous 
+\citep{stephens1974}.  In the early 1970's, 
 \citet{Conover1972} developed the method implemented here
@@
 these two functions, notice that
-\begin{align}
-\sup_x &\left| f(x)- g(x) \right|  \nonumber  \\
-       &= \max_i \left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right) \\
-&=  \max_i \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg).
-\end{align}
-Computing the maximum over these $2N$ (*** why 2N? ***)
+\begin{align*}
+&\sup_x \left| f(x)- g(x) \right|  \nonumber  \\
+       &= \max_i \left( \max\left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right)\right) \\
+&=  \max_i \bigg(\max \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg) \bigg).
+\end{align*}
+Computing the maximum over these $2N$
 values (with $f$ equal to
@@
 Having obtained the test statistic, the p-value must then be calculated. 
-For larger sample sizes, the null distribution coincides 
-(***should this be
-stated precisely as an asymptotic result???)
-with the standard null distribution (CITATION???). 
+For larger sample sizes, the null distribution can be approximated
+with the standard null distribution. 
 When an exact p-value is required for
@@
 
-** code hasn't previously existed?  I deleted this stuff, not sure if
-it fits here.  Is it obvious (since we bothered to do this in the first
-place)?  Or move it?
+%** code hasn't previously existed?  I deleted this stuff, not sure if
+%it fits here.  Is it obvious (since we bothered to do this in the first
+%place)?  Or move it?
 
@@
 of alternatives hypotheses. 
-The original test was developed jointly (*** or was it simultaneously? ***) by
+The original test was developed by
 Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
@@
 The relative powers of these tests to different alternatives are
-studied in depth in \cite{stephens1974}. In general, the $W^2$
-statistic is recommended unless there is a compelling reason to
-consider one of the others. The $A^2$ statistic was developed by
-Anderson in the process of generalizing the test for the two-sample case,
-but was shown to perform very similarly to the original statistic
-in the one-sample setting (*** so why, then would we ever use it??? ***).
+studied in depth in \cite{stephens1974}. The $W^2$
+statistic was the original test statistic; it has the benefit of having
+the simplest form.
+The $A^2$ statistic was developed by
+Anderson in the process of generalizing the test for the two-sample case.
 Watson's $U^2$ statistic was developed for distributions
 which are cyclic (with an ordering to the support but
-no natural starting point  *** how best to describe this? ***). 
+no natural starting point), as it is invariant to cyclic reordering of
+the support. 
 For example, a distribution on the months of the year could be considered
-cyclic. Watson's statistic is invariant to cyclical reordering (giving preference to one ordering would be unnatural).  **** is this last point necessary? ***
-
-*** Is a second paragraph on Stephen's results really necessary?  Can we
-pick and choose and get away with one paragraph? ***
+cyclic.
 
@@
 
-*** I edited, below, make sure you check it. ****
+%*** I edited, below, make sure you check it. ****
 
@@
 Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, although the discreteness allows for slightly simpler representations.
-The null distribution of the test statistics are again hypothesis-dependent.
-The methods do not offer finite sample results, but rather show that the asymptotic null distribution is equal to a weighted sum of independent
+The null distribution of the test statistics are again hypothesis-dependent. The methods do not offer finite sample results, but rather 
+show that the asymptotic distribution of the test statistic under the null hypothesis is equal to the distribution of a weighted sum of independent
 chi-squared variables (with the weights depending on the particular
-null distribution). (*** is this true for all 3? ***)
+null distribution). 
 We implement this asymptotic distribution here;
-the original papers (*** really, all of them? ***)
+the original papers %(*** really, all of them? ***)
 show that this approximation is conservative 
@@
 
-*** What does it mean for a distribution to be equal to a weighted sum of
-independent chi-squared variables?  I think this needs to be stated more
-precisely, above.  A formal definition of $Q$ is needed someplace, too,
-as I'll probably note below. ***
+%*** What does it mean for a distribution to be equal to a weighted sum of
+%independent chi-squared variables?  I think this needs to be stated more
+%precisely, above.  A formal definition of $Q$ is needed someplace, too,
+%as I'll probably note below. ***
 
@@
 this approximation is useful for smaller p-values (smaller than 0.001,
-based on our observations of the numerical instablity of the original
-formulaton); it should not greatly affect the interpretation of the results.
+based on our observations of the numerical instability of the original
+formulation); it should not greatly affect the interpretation of the results.
 
-*** How is the cutoff established in the code? ***
+%*** How is the cutoff established in the code? ***
 
-Figure \ref{cvmissues} shows the non-monotonicity of the function 
-(*** which function, reference it? ***) as the test statistic grows, 
-compared to the conservative approximation.
-
-*** stuff deleted, redundant ***.
+Figure \ref{cvmissues} shows the non-monotonicity of the numerical
+evaluation of equation 7 for the null distribution which is
+uniform on the set $\{1,2,3\}$, as well as the behavior of the
+conservative approximation in equation 9.
 
@@
 \caption{Plot of p-value for given test statistics using numerical integration
-(pink) versus a conservative chi-squared bound (dashed blue). }
+(pink) versus a conservative chi-squared bound (dashed blue). The null distribution
+is unform on the set $\{1,2,3\}$. }
 \label{cvmissues}
@@
 
-*** Discussion of what we might have broken: cases where the user
-provided a discrete distribution to the original ks.test() even though
-it wasn't intended to be supported. ***
+%*** Discussion of what we might have broken: cases where the user
+%provided a discrete distribution to the original ks.test() even though
+%it wasn't intended to be supported. ***
 
@@
 
-TAYLOR: good cyclical example, relating to new material about to go into
-earlier section.
+Cyclical example:
+
+\begin{Schunk}
+\begin{Sinput}
+> library(cvm.test)
+> x <- sample(1:4, 20, replace=T)
+> cvm.test(x, ecdf(1:4), type='U2')
+\end{Sinput}
+\begin{Soutput}
+	Cramer-von Mises - U2
+
+data:  x 
+U2 = 0.075, p-value = 0.4027
+alternative hypothesis: Two.sided 
+\end{Soutput}
+\begin{Sinput}
+> y <- x%%4 + 1
+> cvm.test(y, ecdf(1:4), type = 'U2')
+\end{Sinput}
+\begin{Soutput}
+	Cramer-von Mises - U2
+
+data:  y 
+U2 = 0.075, p-value = 0.4027
+alternative hypothesis: Two.sided 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-*** I commented out a bunch of stuff.  I think the discussion can be short
-and sweet. ***
+%*** I commented out a bunch of stuff.  I think the discussion can be short
+%and sweet. ***
 
@@
 
-*** Hmm... if you claim the following, you want to make sure it is true.
-Did you read it somewhere? ***
+%*** Hmm... if you claim the following, you want to make sure it is true.
+%Did you read it somewhere? ***
 
@@
 therefore the two-sample case would surely have to depend
-on the exact distributions used as well (*** is this true? ***), 
+on the exact distributions used as well %(*** is this true? ***), 
 which are generally not even stated in the two-sample case.

### 2010-11-18 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@29 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 to discrete null distributions has existed
-for several decades. Despite this, modern statistical software
+for several decades. However, modern statistical software
 has generally failed to provide this methodology
 to users. We offer a revision of R's \code{ks.test()} function
-and a new \code{cvm.test()} function that serve to fill this need
+and a new \code{cvm.test()} function that fill this need
 in the R language for two of the most popular nonparametric
@@
 are the classic test proposed by Kolmogorov and Smirnov followed closely by
-several variants on a procedure proposed by
-Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
-tests, or simply Cram\'{e}r-von Mises tests) .
+several variants on Cram\'{e}r-von Mises tests.
+
+%procedure proposed by
+%Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
+%tests, or simply Cram\'{e}r-von Mises tests).
 
@@
 tests are intended for continuous null distributions, but they have
-also been adapted for use with discrete null distributions.  Unfortunately,
-most modern statistical software have failed to incorporate these discrete
-versions, leaving researchers with the $\chi^2$ test or in the unfortunate
-position of applying a nonparametric test designed for a continuous null
-distribution and hoping for the best.
-As we will see, this latter course of action can be particularly dangerous
-with small sample sizes. This paper presents a revision of R's
+also been adapted for discrete null distributions.  Unfortunately,
+most modern statistical software packages and programming environments
+have failed to incorporate these discrete versions.  As a result,
+researchers would typically rely upon the $\chi^2$ test or
+use of a nonparametric test designed for a continuous null distribution.
+For smaller sample sizes, in particular, both of these choices can produce
+misleading inferences.
+
+This paper presents a revision of R's
 \code{ks.test()} function and a new \code{cvm.test()} function to
 fill this void for researchers and practitioners in the R environment. 
+We first present overviews of the theory and general implementation of the
+discrete Kolmogorov-Smirnov and Cram\'{e}r-von Mises tests.  We discuss
+the particular implementation of tests in R and provide examples.  We
+conclude with a short discussion, including the state of existing continuous
+and two-sample Cram\'{e}r-von Mises testing in R.
 
@@
 The most popular nonparametric goodness-of-fit test is the
-Kolmogorov-Smirnov test. The idea is fairly simple:
+Kolmogorov-Smirnov test.
 Given the cumulative distribution
@@
 \begin{align}
-D = \sup_x \left| F_0(x)- F_{data}(x) \right|.
+D = \sup_x \left| F_0(x)- F_{data}(x) \right|.    \label{eqD}
 \end{align}
@@
 performance relative to other algorithms. 
-The test statistic (1) is easily adapted for one-sided tests.
-For these, the absolute value is discarded and the tests are based
+The test statistic is easily adapted for one-sided tests.
+For these, the absolute value in (\ref{eqD}) is discarded and the tests are based
 on either the supremum of the remaining difference (the `greater' testing
@@
 The extension of the Kolmogorov-Smirnov test to non-continuous 
-null distributions does not have such a clean solution. The formula of
-the test statistic $D$ remains unchanged, but the distribution
-of the test statistic is much more difficult to obtain. Unlike the 
-continuous case, the null distribution depends on the null model.
+null distributions is not straightforward. The formula of
+the test statistic $D$ remains unchanged, but its distribution
+is much more difficult to obtain; unlike the 
+continuous case, its distribution depends on the null model.
 It has been known since at least the 1950's that using the tables associated
@@
 when the null distribution is discontinuous 
-\citep{stephens1974}.  In the early 1970's, 
+(see *** or *** or \citet{stephens1974}, for example).  In the early 1970's, 
 \citet{Conover1972} developed the method implemented here
@@
 
-The implementation of the discrete Kolmogorov-Smirnov test is considered
-in two parts. First, the particular test statistic needs to be calculated
+The implementation of the discrete Kolmogorov-Smirnov test involves
+two steps. First, the particular test statistic is calculated
 (corresponding to the desired one-sided or two-sided test).
@@
 
-The test statistic is, in terms of the theory, the same as in the
-continuous case; it would seem that no additional work would be required,
-but unfortunately this is not the case.
+The form of the test statistic is the same as in the
+continuous case; it would seem that no additional work would be required
+for the implemenetation, but this is not the case.
 Consider two non-decreasing functions $f$ and $g$, where the function $f$ is a step function with jumps on the set $\{x_1, \ldots x_N \}$ and $g$
@@
 In order to determine the supremum of the difference between
-these two functions, notice that
-\begin{align*}
+these two functions, notice that **** FORMAT THESE AT END ****
+\begin{align}
 &\sup_x \left| f(x)- g(x) \right|  \nonumber  \\
-       &= \max_i \left( \max\left( \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right)\right) \\
-&=  \max_i \bigg(\max \bigg( \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg) \bigg).
-\end{align*}
+       &= \max_i \left( \max\left\{ \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right\} \right) \label{SUP1}\\
+&=  \max_i \bigg(\max \bigg\{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg\} \bigg). \label{SUP2}
+\end{align}
 Computing the maximum over these $2N$
@@
 a continuous null distribution. When the function $g$ is not
-continuous, however, equality (3) does not hold in general because 
-we cannot replace $\lim_{x\rightarrow x_e} g(x)$ with the value $g(x_i)$.
+continuous, however, equality (\ref{SUP2}) does not hold in general because 
+we cannot replace $\lim_{x\rightarrow x_i} g(x)$ with the value $g(x_i)$.
 
@@
 \sup_x &\left| f(x)- g(x) \right| =  \nonumber \\
-        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \}
+        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \} \label{epsilon}
 \end{align}
-where the discontinuities in $g$ are at least a distance $\epsilon$ apart. 
+where the discontinuities in $g$ are more than some distance $\epsilon$ apart. 
 This, however, requires knowledge that $g$ is a step function as well as of
@@
 For larger sample sizes, the null distribution can be approximated
-with the standard null distribution. 
+with the null distribution from the classical Kolmogorov-Smirnov test. 
 When an exact p-value is required for
@@
 \begin{align}
-W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \\
-A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \\
-U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x)
+W^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) \right]^2 dF_0(x) \label{W2} \\
+A^2 &= n \cdot \int_{-\infty}^{\infty} \frac{\left[F_{data}(x)- F_{0}(x) \right]^2}{F_0(x) -F_0(x)^2} dF_0(x) \label{A2} \\
+U^2 &= n \cdot \int_{-\infty}^{\infty} \left[ F_{data}(x)- F_{0}(x) - W^2 \right]^2 dF_0(x) \label{U2}
 \end{align}
@@
 
-The relative powers of these tests to different alternatives are
-studied in depth in \cite{stephens1974}. The $W^2$
+The $W^2$
 statistic was the original test statistic; it has the benefit of having
@@
 
-It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations. As they all involve
-integration over the whole range of data, rather than one supremum, it is not surprising that they are generally best when the
-true alternative distribution deviates a little over the whole range of data rather than deviating a lot over a small range. For
-a complete analysis of the relative powers of these tests see \cite{stephens1974}.
+It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
+As they all involve
+integration over the whole range of data, rather than use of a supremum.
+Thus, they are best-suited for situations where the
+true alternative distribution deviates a little over the whole support
+rather than having large deviations over a small section of the support. For
+a complete analysis of the relative powers of these tests, see 
+\cite{stephens1974}.
 
@@
 distributions were developed in \cite{choulakian1994}. As with the
-Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, although the discreteness allows for slightly simpler representations.
-The null distribution of the test statistics are again hypothesis-dependent. The methods do not offer finite sample results, but rather 
-show that the asymptotic distribution of the test statistic under the null hypothesis is equal to the distribution of a weighted sum of independent
+Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, 
+and the null distributions of the test statistics are again hypothesis-dependent. The methods do not offer finite-sample results, but rather 
+show that the asymptotic distributions of the test statistics under the null hypothesis each involve consideration of a weighted sum of independent
 chi-squared variables (with the weights depending on the particular
 null distribution). 
-We implement this asymptotic distribution here;
-the original papers %(*** really, all of them? ***)
-show that this approximation is conservative 
-and asymptotically equivalent to the true null distribution.
-
-%*** What does it mean for a distribution to be equal to a weighted sum of
-%independent chi-squared variables?  I think this needs to be stated more
-%precisely, above.  A formal definition of $Q$ is needed someplace, too,
-%as I'll probably note below. ***
+We implement the asymptotic distributions here; \cite{stephens1974}
+showed that these approximations are conservative 
+and asymptotically equivalent to the true null distributions.
 
@@
 Calculation of the three test statistics is done using the
-matrix algebra given by \cite{choulakian1994}. The form of the
-asymptotic null distribution is obtained using R's
-eigenvalue decomposition functions.
-
-The only notable difficulty in the process involves calculating the percentiles
+matrix algebra given by \cite{choulakian1994}. 
+The only notable difficulty in the implementation of the discrete form
+of the tests involves calculating the percentiles
 of the weighted sum of chi-squares,
 \begin{align}
-Q = \sum_{i=1}^{p} \lambda_i \chi^2_{i,1df}
+Q = \sum_{i=1}^{p} \lambda_i \chi^2_{i,1df},   \label{eqQ}
 \end{align}
-\cite{imhof1961} provides a method for obtaining the distribution of
-a weighted sum of independent chi-squared variables.  It is
+where $p$ is the number of elements in the support of the hypothesized
+distribution.
+\cite{imhof1961} provides a method for obtaining the distribution of $Q$,
 easily adapted for our case because
@@
 \frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
+\label{distQ}
 \end{align}
-for continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ depending on the actual weights used.
+for continuous functions $\theta(\cdot, x)$ and $\rho(\cdot)$ depending on the  weights $\lambda_i$.
 
-There is no analytic solution to the integral, so the integration is
+There is no analytic solution to the integral in (\ref{distQ}), 
+so the integration is
 accomplished numerically. This seems fine in most situations we considered,
@@
 \begin{align}
-\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \\
+\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \label{ineq1} \\
 &= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
+\label{ineq2}
 \end{align}
@@
 The original formulation is preferable for most p-values, while
-this approximation is useful for smaller p-values (smaller than 0.001,
+this approximation is used for smaller p-values (smaller than 0.001,
 based on our observations of the numerical instability of the original
-formulation); it should not greatly affect the interpretation of the results.
-
-%*** How is the cutoff established in the code? ***
+formulation).
 
 Figure \ref{cvmissues} shows the non-monotonicity of the numerical
-evaluation of equation 7 for the null distribution which is
+evaluation of equation (\ref{distQ}) for the null distribution which is
 uniform on the set $\{1,2,3\}$, as well as the behavior of the
-conservative approximation in equation 9.
+conservative upper bound given in equation (\ref{ineq2}).
 
@@
 \end{center}
-\caption{Plot of p-value for given test statistics using numerical integration
-(pink) versus a conservative chi-squared bound (dashed blue). The null distribution
-is unform on the set $\{1,2,3\}$. }
+\caption{Plot of calculated p-values for given test statistics
+using numerical integration
+(pink) compared to the conservative chi-squared bound (dashed blue). 
+The null distribution is uniform on the set $\{1,2,3\}$ in this example.
+The instability in the calculated p-values is a result of numerical
+instabilities.}
 \label{cvmissues}

### 2010-11-18 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@30 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 two-sided alternative hypotheses, but the reported p-values will be
-conservative.  For one-sided tests,
+conservative; very tight bounds for interesting (small) p-values are
+presented in \cite{Conover1972}.  For one-sided tests,
 exact p-values are calculated using Conover's
@@
 When \code{exact = FALSE} the asymptotic distribution is used which is known
-to be imprecise but conservative, even for small samples (CITATION).
+to be conservative though imprecise for small samples (CITATION).
 
@@
 
-The function \pkg{cvm.test()} is similar to \pkg{ks.test()}.  Its first two
+The function \pkg{cvm.test()} is similar in design
+to \pkg{ks.test()}.  Its first two
 arguments specify the data and null distribution; the only extra option,
@@
 As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
-\code{htest}.
+\code{htest}.  Although we designed \code{cmv.test()} primarily for cases with
+discrete distributions \code{y}, we implemented the continuous forms of the
+test for variants \code{W2} and \code{A2} (as noted in the Discussion, below).
 
@@
 
-Consider a toy example, with observed data of length 2 (specifically, the
+Consider a toy example with observed data of length 2 (specifically, the
 values 0 and 1) and a hypothized null distribution that places equal
@@
 \begin{Sinput}
+> library(ks.test)
 > ks.test(c(0, 1), ecdf(c(0, 1)))
@@
 
-Next, we simulate a sample of 25 from the discrete uniform distribution on
+Next, we simulate a sample of size 25 from the discrete uniform distribution on
 the integers $\{1, 2, \ldots, 10\}$ and show several variants of the
@@
 p-value.  In this case, the approximation may not be that tight, but this is
-irrelevant for such large p-values (for more interesting p-values, the
-upper bound is very close to the true p-value).
+irrelevant for such large p-values (for more interesting (small) p-values, the
+upper bound is very close to the true p-value as shown in \cite{Conover1972}).
 \begin{Schunk}
 \begin{Sinput}
-> library(ks.test)
 > set.seed(1)
@@
 
-Finally, we employ two of the Cram\'{e}r-von Mises tests.  Need short discussion,
-and we could use a good cyclical example:
+Finally, we demonstrate the Cram\'{e}r-von Mises tests with the same
+simulated data. 
 \begin{Schunk}
@@
 
-Cyclical example:
+*** Cyclical example? ***
 
@@
 
-In the end, for methods such as those presented in this paper where it is possible,
-there are enough positives to suggest that we attempt to implement exact $p$-value calculations. This does lead
-to an interesting direction for future research in non-parametric goodness of fit tests: the methods presented
-here were historically chosen because they could be shown to have fixed null distributions which could easily
-calculated in an era without fast computers for carrying out permutation tests. It is quite possible that
-variants without the property, which could be easily used today, have greater power to certain alternative 
-distributions.
+This paper presents the implementation of a range of nonparametric
+goodness-of-fit tests for discrete null distributions.  In some cases,
+the p-values are known to be exact.  In others, conservativeness 
+and tightness of bounds have been established in the literature.  As a result,
+no simulations are necessary for these methods (which were developed during
+an era when extensive simulations may have been prohibitively expensive
+or time-consuming).
+However, this does raise the possibility that alternative tests relying upon
+modern computational abilities could provide even greater power in certain
+situations, a possible avenue for future work.
 
@@
 In the continuous setting, both of 
-the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests have two-sample analogues. Here data are
-observed from two processes, and the hypothesis tested is 
-whether they came from the same (but unspecified) distribution. There
-does not exist an analogous theory for discrete distributions. 
+the Kolmogorov-Smirnov and the Cram\'{e}r-von Mises tests
+have two-sample analogues. When data are
+observed from two processes or sampled from two populations, 
+the hypothesis tested is 
+whether they came from the same (unspecified) distribution. An
+analogous theory for discrete distributions does not exist.
 This comes from the fact that the discrete null distributions
@@
 which are generally not even stated in the two-sample case. 
+*** I don't think I buy this. Couldn't this just be done as a
+chi-squared test on a 2xp table, at least? *** Also,
+see \url{http://ideas.repec.org/p/mtl/montde/2001-23.html}.
 
 While we have implemented the two most popular variants of 
-goodness-of-fit tests, there are 
-several more exotic varieties to be found. For further 
-generalizations of tests see the extended study done in \cite{dewev1973}.
+goodness-of-fit tests for discrete distributions, further 
+generalizations of tests are described in the
+extended study of \cite{dewev1973}.  We also provide general one-sample
+Cram\'{e}r-von Mises testing for continuous distributions simply
+to help complement the abilities provided by several other
+R packages, although these tests are not the focus of this paper.
+In particular, functions implemented in package \pkg{nortest} focus on the
+composite hypothesis of normality, while
+package \pkg{ADGofTest} provides the Anderson-Darling variant
+of the test for continuous distributions.  Packages
+\pkg{CvM2SL1Test} and \pkg{CvM2SL1Test} provide support for
+two-sample Cram\'{e}r-von Mises tests with continuous distributions.
+ *** Need to be added
+to the bibliography. ***
 

### 2010-11-18 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@31 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 two-sample Cram\'{e}r-von Mises tests with continuous distributions.
+Package \pkg{cramer} offers a multivariate Cram\'{e}r test for
+the two-sample problem.  Possible reference, though it isn't clear
+if this particular one is really part of this literature segment?
+\url{http://en.scientificcommons.org/39157548}
+
  *** Need to be added
-to the bibliography. ***
+to the bibliography, whatever we choose to do ***
 

### 2010-11-19 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@32 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \code{ks.test()} function and a new \code{cvm.test()} function to
-fill this void for researchers and practitioners in the R environment. 
+fill this void for researchers and practitioners in the R environment
+\citep{R2010}. 
 We first present overviews of the theory and general implementation of the
 discrete Kolmogorov-Smirnov and Cram\'{e}r-von Mises tests.  We discuss
-the particular implementation of tests in R and provide examples.  We
+the particular implementation of the tests in R and provide examples.  We
 conclude with a short discussion, including the state of existing continuous
@@
 \citet{Conover1972} developed the method implemented here
-for computing exact p-values
+for computing p-values
 in the case of discrete null distributions.
@@
 In order to determine the supremum of the difference between
-these two functions, notice that **** FORMAT THESE AT END ****
+these two functions, notice that
 \begin{align}
-&\sup_x \left| f(x)- g(x) \right|  \nonumber  \\
-       &= \max_i \left( \max\left\{ \left|g(x_i) - f(x_i) \right|, \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1}) \right| \right\} \right) \label{SUP1}\\
-&=  \max_i \bigg(\max \bigg\{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i) - f(x_{i-1}) \right| \bigg\} \bigg). \label{SUP2}
+\sup_x \left| f(x) - g(x) \right| &  \nonumber  \\
+       =  \max_i \bigg[ \max\bigg(  & \left|g(x_i) - f(x_i) \right|, \nonumber \\
+& \lim_{x \rightarrow x_i} \left| g(x) - f(x_{i-1})
+   \right| \bigg) \bigg] \label{SUP1}\\
+=  \max_i \bigg[\max \bigg( & \left|g(x_i) - f(x_i) \right|, \nonumber \\
+& \left| g(x_i) - f(x_{i-1}) \right| \bigg) \bigg]. \label{SUP2}
 \end{align}
@@
 \sup_x &\left| f(x)- g(x) \right| =  \nonumber \\
-        &                    \max_i \{ \left|g(x_i) - f(x_i) \right|, \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \} \label{epsilon}
+        & \max_i \left( \left|g(x_i) - f(x_i) \right|, 
+    \left| g(x_i - \epsilon) - f(x_{i-1}) \right| \right) \label{epsilon}
 \end{align}
@@
 This, however, requires knowledge that $g$ is a step function as well as of
-the nature of its support (or break-points).
-%
-%*** Hmm... do we care about time? ***  When the break-points of $g$ are
-%unknown, or when $g$ is neither a step function nor a continuous distribution
-%function, computing the supremum requires taking the numerical limit in (1***?). 
-%This is computational involved and risks problems with numerical tolerances.
-%
-We implement the Kolmogorov-Smirnov test statistic for discrete null
-distributions by forcing the user to completely specify the null distribution.
-
-%**** Hmmm... It is often the case that the test is used inside of a
-%long simulation, 
-%and the added computational time would likely be prohibitive.
+the nature of its support (or break-points).  As a result,
+we implement the Kolmogorov-Smirnov test statistic for discrete null
+distributions by requiring the complete specification of the null distribution.
 
@@
 of our revised function \code{ks.test()} and in Conover's original paper.
-
-%** code hasn't previously existed?  I deleted this stuff, not sure if
-%it fits here.  Is it obvious (since we bothered to do this in the first
-%place)?  Or move it?
-
-%Code for this procedure in the R language, or in any other open source options, did not previously
-%exist and is included in the new function \pkg{ks.test}.
 
@@
 While the Kolmogorov-Smirnov test may be the most popular of
-the nonparametric goodness-of-fit tests, there is another family of 
-tests which has been shown to be more powerful against a large class
+the nonparametric goodness-of-fit tests, Cram\'{e}r-von Mises
+tests have been shown to be more powerful against a large class
 of alternatives hypotheses. 
@@
 
-The $W^2$
-statistic was the original test statistic; it has the benefit of having
-the simplest form.
+The $W^2$ statistic was the original test statistic.
 The $A^2$ statistic was developed by
@@
 which are cyclic (with an ordering to the support but
-no natural starting point), as it is invariant to cyclic reordering of
-the support. 
-For example, a distribution on the months of the year could be considered
-cyclic.
+no natural starting point); it is invariant to cyclic reordering of
+the support.   For example, a distribution on the months of the year
+could be considered cyclic.
 
 It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
-As they all involve
-integration over the whole range of data, rather than use of a supremum.
-Thus, they are best-suited for situations where the
+They all involve integration over the whole range of data, rather than
+use of a supremum, so they are best-suited for situations where the
 true alternative distribution deviates a little over the whole support
-rather than having large deviations over a small section of the support. For
-a complete analysis of the relative powers of these tests, see 
-\cite{stephens1974}.
+rather than having large deviations over a small section of the support. 
+\cite{stephens1974} offers a complete analysis of the relative powers
+of these tests.
 
@@
 Kolmogorov-Smirnov test, the forms of the test statistics are unchanged, 
-and the null distributions of the test statistics are again hypothesis-dependent. The methods do not offer finite-sample results, but rather 
-show that the asymptotic distributions of the test statistics under the null hypothesis each involve consideration of a weighted sum of independent
+and the null distributions of the test statistics are again hypothesis-dependent. \cite{choulakian1994} does not offer finite-sample results, but rather 
+shows that the asymptotic distributions of the test statistics under the null hypothesis each involve consideration of a weighted sum of independent
 chi-squared variables (with the weights depending on the particular
@@
 and asymptotically equivalent to the true null distributions.
+
+**** Why would Stephens (1974) have this, when they were developed in 1994? ****
 
@@
 the approximation becomes unstable. 
-%This is further magnified in situations where the 
-%$p$-values should be very small, where tiny fluctuations (undetectable
-%elsewhere) become quite prominent. 
+As an example of this numerical instability, the red plotted in
+Figure \ref{cvmissues} shows the non-monotonicity of the numerical
+evaluation of equation (\ref{distQ}) for a null distribution that is
+uniform on the set $\{1,2,3\}$.
+
 We resolve this problem by using a simple conservative approximation to
@@
 The original formulation is preferable for most p-values, while
-this approximation is used for smaller p-values (smaller than 0.001,
+this upper bound is used for smaller p-values (smaller than 0.001,
 based on our observations of the numerical instability of the original
-formulation).
-
-Figure \ref{cvmissues} shows the non-monotonicity of the numerical
-evaluation of equation (\ref{distQ}) for the null distribution which is
-uniform on the set $\{1,2,3\}$, as well as the behavior of the
-conservative upper bound given in equation (\ref{ineq2}).
+formulation).  Although it would be preferable to determine the use of
+the bound based on values of the test statistic rather than the p-value,
+the range of ``extreme'' values of the test statistic varies with the
+hypothesized distribution.
 
@@
 using numerical integration
-(pink) compared to the conservative chi-squared bound (dashed blue). 
+(red) compared to the conservative chi-squared bound (dashed blue). 
 The null distribution is uniform on the set $\{1,2,3\}$ in this example.
@@
 or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
-they will be proposed for inclusion in \pkg{stats} in late 2010.
+\code{ks.test()} will be proposed for inclusion in \pkg{stats} in late 2010.
 

### 2010-11-19 rjournal/codeforgraphics.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@34 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -91,7 +91,7 @@ for(i in 1:length(I)) {
 
 #pdf('fig1.pdf', height=6, width=8)
 coff <- 1.2
-coff2 <- 1.2
+coff2 <- 1.3
 plot(I[I > coff], z[I > coff], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
 abline(h=0, lty=2)
 lines(I[I > coff2], y[I > coff2], lty=2, col='blue', lwd=2)

### 2010-11-19 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@33 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@35 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-%General idea of goodness-of-fit tests (hypothesis test about underlying distribution).
-%Given an observed sequence of data, the typical question of statistical inference is to determine
-%the underlying process from which it came. 
 Goodness-of-fit tests are used to assess whether data are consistent
@@
 several variants on Cram\'{e}r-von Mises tests.
-
-%procedure proposed by
-%Cram\'{e}r and von Mises (tests sometimes called Cram\'{e}r-von Mises-Smirnov
-%tests, or simply Cram\'{e}r-von Mises tests).
-
-%*** why "estimation procedure" above?
-%*** Citations in the paragraph above?  Also check the body of paper. ***
 
@@
 
-%*** Later, make sure we point out how using one of the ``old'' tests and
-%hoping for the best is a problem. ***
-
-%Given a particular null distribution, goodness-of-fit tests 
-%are used in order to test whether it is likely that the data were
-%generated via the null distribution.
-
-%Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
-%While almost any hypothesis test can be viewed as a variant on a
-%goodness-of-fit test, the term ``goodness-of-fit test'' is typically
-%reserved for those tests which are nonparametric in nature. 
-%That is, they do not conduct the hypothesis testing 
-%through the explicit calculation of the parameters generating the data. Instead, they %generally attempt to
-%calculate differences between the empirical distribution of the observed data and the overall distribution 
-%of the null model. In many cases these tests are preferential since they tend to have an increased power for 
-%interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
-%factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
-%Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.
-
-% Discrete versions exist; just no code
-%While the original aim of nonparametric goodness of fit tests was meant for continuous null distributions, discrete versions
-%have existed since the early 1970s. Unfortunately, current statistical software has largely failed to incorporate them.
-%This has left the end-user to either use only parametric tests, such as Pearson's Chi-Squared test, or to incorrectly
-%apply the functions meant for continuous distributions.
-
 \section{Kolmogorov-Smirnov Test}
-
-% Most popular choice of test; explain the basic algorithm and purpose for continuous distributions (maybe a few P's).
 
@@
 (the `lesser' hypothesis alternative).  Tabulated p-values have been
-available for these tests since 1933. %(*** year ****).
-
-%**** check this above, both do away with the absolute value, I think? ***
+available for these tests since 1933.
 
@@
 when the null distribution is discontinuous 
-(see *** or *** or \citet{stephens1974}, for example).  In the early 1970's, 
+(see \cite{slakter1965}, \cite{goodman1954}, and \cite{massey1951}).  
+In the early 1970's, 
 \citet{Conover1972} developed the method implemented here
@@
 Full details of the calculations are contained in source code
-of our revised function \code{ks.test()} and in Conover's original paper.
+of our revised function \code{ks.test()} and in Conover's paper.
 
@@
 
-%Discuss the discrete version. How to calculate W2, U2, and A2.
-
-%*** I edited, below, make sure you check it. ****
-
 Generalizations of the Cram\'{e}r-von Mises tests to discrete
@@
 null distribution).
-We implement the asymptotic distributions here; \cite{stephens1974}
-showed that these approximations are conservative 
-and asymptotically equivalent to the true null distributions.
-
-**** Why would Stephens (1974) have this, when they were developed in 1994? ****
 
@@
 based on our observations of the numerical instability of the original
-formulation).  Although it would be preferable to determine the use of
+formulation).  Figure \ref{cvmissues} shows this bound with the blue dashed
+line; values in red exceeding the bound are a result of the numerical
+instability.  Although it would be preferable to determine the use of
 the bound based on values of the test statistic rather than the p-value,
@@
 The null distribution is uniform on the set $\{1,2,3\}$ in this example.
-The instability in the calculated p-values is a result of numerical
-instabilities.}
+The sharp variations in the calculated p-values are a result of numerical
+instabilities, and the true p-values are bounded by the blue dashed curve.}
 \label{cvmissues}
@@
 
-The function \pkg{cvm.test()} is similar in design
-to \pkg{ks.test()}.  Its first two
+The function \code{cvm.test()} is similar in design
+to \code{ks.test()}.  Its first two
 arguments specify the data and null distribution; the only extra option,
@@
 
-Finally, we demonstrate the Cram\'{e}r-von Mises tests with the same
+We demonstrate the Cram\'{e}r-von Mises tests with the same
 simulated data. 
@@
 
-*** Cyclical example? ***
-
+We conclude with a toy cyclical example showing that the test is
+invariant to cyclic reordering of the support.
 \begin{Schunk}
@@
 > library(cvm.test)
-> x <- sample(1:4, 20, replace=T)
-> cvm.test(x, ecdf(1:4), type='U2')
-\end{Sinput}
-\begin{Soutput}
-	Cramer-von Mises - U2
-
-data:  x 
-U2 = 0.075, p-value = 0.4027
-alternative hypothesis: Two.sided 
-\end{Soutput}
-\begin{Sinput}
-> y <- x%%4 + 1
-> cvm.test(y, ecdf(1:4), type = 'U2')
+> set.seed(1)
+> y <- sample(1:4, 20, replace=T)
+> cvm.test(y, ecdf(1:4), type='U2')
 \end{Sinput}
@@
 data:  y 
-U2 = 0.075, p-value = 0.4027
+U2 = 0.0094, p-value = 0.945
+alternative hypothesis: Two.sided 
+\end{Soutput}
+\begin{Sinput}
+> z <- y%%4 + 1
+> cvm.test(z, ecdf(1:4), type = 'U2')
+\end{Sinput}
+\begin{Soutput}
+	Cramer-von Mises - U2
+
+data:  z 
+U2 = 0.0094, p-value = 0.945
 alternative hypothesis: Two.sided 
@@
 \end{Schunk}
+In contrast, the Kolmogorov-Smirnov or the standard Cram\'{e}r-von Mises tests
+produce different results after such a reordering. For example, the default
+Cram\'{e}r-von Mises yields p-values of 
+*** we have a problem, or maybe got really unlucky? ***.
 
@@
 the hypothesis tested is 
-whether they came from the same (unspecified) distribution. An
-analogous theory for discrete distributions does not exist.
+whether they came from the same (unspecified) distribution. 
 This comes from the fact that the discrete null distributions
@@
 therefore the two-sample case would surely have to depend
-on the exact distributions used as well %(*** is this true? ***), 
-which are generally not even stated in the two-sample case. 
-*** I don't think I buy this. Couldn't this just be done as a
-chi-squared test on a 2xp table, at least? *** Also,
-see \url{http://ideas.repec.org/p/mtl/montde/2001-23.html}.
+on the exact distributions used as well
+which are generally not even stated in the two-sample case
+\cite{newpaper}.  **** explored goodness-of-fit tests for
+discrete distributions using a permutation test approach.
 
-While we have implemented the two most popular variants of 
-goodness-of-fit tests for discrete distributions, further 
-generalizations of tests are described in the
+Further 
+generalizations of goodness-of-fit tests for discrete distributions
+are described in the
 extended study of \cite{dewev1973}.  We also provide general one-sample
@@
 R packages, although these tests are not the focus of this paper.
-In particular, functions implemented in package \pkg{nortest} focus on the
+In particular, functions implemented in package \pkg{nortest} 
+\citep{nortest} focus on the
 composite hypothesis of normality, while
-package \pkg{ADGofTest} provides the Anderson-Darling variant
+package \pkg{ADGofTest} \citep{ADGofTest} provides the Anderson-Darling variant
 of the test for continuous distributions.  Packages
-\pkg{CvM2SL1Test} and \pkg{CvM2SL1Test} provide support for
+\pkg{CvM2SL1Test} \citep{CvM2SL1Test} and \pkg{CvM2SL2Test} 
+\citep{CvM2SL2Test} provide support for
 two-sample Cram\'{e}r-von Mises tests with continuous distributions.
-Package \pkg{cramer} offers a multivariate Cram\'{e}r test for
-the two-sample problem.  Possible reference, though it isn't clear
-if this particular one is really part of this literature segment?
-\url{http://en.scientificcommons.org/39157548}
-
- *** Need to be added
-to the bibliography, whatever we choose to do ***
+Package \pkg{cramer} \citep{cramer} offers a multivariate Cram\'{e}r test for
+the two-sample problem.
 

### 2010-11-19 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@37 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 distribution, making this a computationally
-attractive method. \cite{slakter65} offers a standard presentation
+attractive method. \cite{slakter1965} offers a standard presentation
 of the test and its
@@
 if the sample size is less than or equal to 30; otherwise, asymptotic distributions
-are used which are reliable in such cases (CITATION? Is this correct?).
-When \code{exact = FALSE} the asymptotic distribution is used which is known
-to be conservative though imprecise for small samples (CITATION).
+are used which are reliable in such cases.
+When \code{exact = FALSE} the asymptotic distribution is used and known
+to be conservative though imprecise for small samples (see \cite{Conover1972}
+for details).
 
@@
 produce different results after such a reordering. For example, the default
-Cram\'{e}r-von Mises yields p-values of 
-*** we have a problem, or maybe got really unlucky? ***.
+Cram\'{e}r-von Mises test yields p-values of 0.8237 and 0.9577 with the original
+and transformed data $y$ and $z$, respectively.
 
@@
 
-%Possibly go into a brief discussion of classes and class inheritance in R.
-The use of `htest' in our functions; benefits and issues with using it.
-Particularly troublesome in ks.test where we have a range of p-values.
-More precisely, explain that we
-%could provide upper and lower bounds for the p-value but are limited by class
-%\code{htest}.
-
 In the continuous setting, both of 
@@
 whether they came from the same (unspecified) distribution. 
-This comes from the fact that the discrete null distributions
-of the test statistics depend on the exact null distribution; 
-therefore the two-sample case would surely have to depend
-on the exact distributions used as well
-which are generally not even stated in the two-sample case
-\cite{newpaper}.  **** explored goodness-of-fit tests for
+With the discrete case, however, the null distribution of the
+test statistic in the discrete case depends on the underlying
+probability model, as discussed by \cite{walsh1963}.
+Such an extension would require the specification of a
+null distribution, which generally goes unstated in two-sample
+goodness-of-fit tests.  We note that
+\cite{Mdufour2001} explored goodness-of-fit tests for
 discrete distributions using a permutation test approach.
@@
 are described in the
-extended study of \cite{dewev1973}.  We also provide general one-sample
+extended study of \cite{dewev1973}.  Our \code{cvm.test()} function
+also provides general one-sample
 Cram\'{e}r-von Mises testing for continuous distributions simply
-to help complement the abilities provided by several other
+to help complement the functionality provided by other
 R packages, although these tests are not the focus of this paper.
-In particular, functions implemented in package \pkg{nortest} 
+Functions implemented in package \pkg{nortest} 
 \citep{nortest} focus on the
@@
 package \pkg{ADGofTest} \citep{ADGofTest} provides the Anderson-Darling variant
-of the test for continuous distributions.  Packages
+of the test for general continuous distributions.  Packages
 \pkg{CvM2SL1Test} \citep{CvM2SL1Test} and \pkg{CvM2SL2Test} 
-\citep{CvM2SL2Test} provide support for
+\citep{CvM2SL2Test} provide
 two-sample Cram\'{e}r-von Mises tests with continuous distributions.

### 2010-11-19 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@38 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Next, we simulate a sample of size 25 from the discrete uniform distribution on
-the integers $\{1, 2, \ldots, 10\}$ and show several variants of the
+the integers $\{1, 2, \ldots, 10\}$ and show usage of the
 new \code{ks.test()} implementation.  The first is the default two-sided test,
@@
 p-value.  In this case, the approximation may not be that tight, but this is
-irrelevant for such large p-values (for more interesting (small) p-values, the
-upper bound is very close to the true p-value as shown in \cite{Conover1972}).
+irrelevant for such large p-values. For more interesting (small) p-values, the
+upper bound is very close to the true p-value (shown in \cite{Conover1972}).
 \begin{Schunk}
@@
 \end{Schunk}
+
 A different toy example shows the dangers of using R's existing
@@
 Cram\'{e}r-von Mises test yields p-values of 0.8237 and 0.9577 with the original
-and transformed data $y$ and $z$, respectively.
+and transformed data \code{y} and \code{z}, respectively.
 
@@
 
-This paper presents the implementation of a range of nonparametric
+This paper presents the implementation of several nonparametric
 goodness-of-fit tests for discrete null distributions.  In some cases,
@@
 and tightness of bounds have been established in the literature.  As a result,
-no simulations are necessary for these methods (which were developed during
-an era when extensive simulations may have been prohibitively expensive
-or time-consuming).
+no simulations are necessary for these methods, which were generally
+developed during an era when extensive simulations may have been
+prohibitively expensive or time-consuming).
 However, this does raise the possibility that alternative tests relying upon
@@
 With the discrete case, however, the null distribution of the
-test statistic in the discrete case depends on the underlying
+test statistic depends on the underlying
 probability model, as discussed by \cite{walsh1963}.
@@
 goodness-of-fit tests.  We note that
-\cite{Mdufour2001} explored goodness-of-fit tests for
+\cite{Mdufour2001} explored two-sample goodness-of-fit tests for
 discrete distributions using a permutation test approach.
+
+*** possibly change following paragraph depending on what we
+decide to include in \code{cvm.test()}. ***
 

### 2010-11-29 pkg/cvm.test/R/cvm.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@39 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,25 +1,20 @@
 cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
 
-  cvm.stat <- function(x,y,type) {
-    fdata <- ecdf(x)
-    if(type != 'A2') {
-      igrand <- function(z) return((fdata(z) - y(z))^2)
-    } else {
-      igrand <- function(z) return((fdata(z) - y(z))^2 / (y(z) - y(z)^2))
-    }
-    stat <- integrate(igrand, -Inf, Inf)[1]
-    if(type != 'U2') {
-      return(stat)
-    } else {
-      igrand <- function(z) return((fdata(z) - y(z)-stat)^2) 
-      return(integrate(igrand, -Inf, Inf)[1])
-    }    
-  }
-
-  cvm.pval <- function(stat) {
-
-    return()
-  }  
+  #cvm.stat <- function(x,y,type) {
+  #  fdata <- ecdf(x)
+  #  if(type != 'A2') {
+  #    igrand <- function(z) return((fdata(z) - y(z))^2)
+  #  } else {
+  #    igrand <- function(z) return((fdata(z) - y(z))^2 / (y(z) - y(z)^2))
+  #  }
+  #  stat <- integrate(igrand, -Inf, Inf)[1]
+  #  if(type != 'U2') {
+  #    return(stat)
+  #  } else {
+  #    igrand <- function(z) return((fdata(z) - y(z)-stat)^2) 
+  #    return(integrate(igrand, -Inf, Inf)[1])
+  #  }    
+  #}
 
   cvm.pval.disc <- function(STAT, lambda) {
 
@@ -52,7 +47,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
       df <- sum(lambda != 0)
       return(dchisq(STAT/max(lambda),df))
     }
-  } # End cvm.pval()
+  } # End cvm.pval.disc()
 
   cvm.stat.disc <- function(x,y, type=c("W2", "U2", "A2")) {
     type <- match.arg(type)
@@ -90,7 +85,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
                    A2 = sum((Z^2*t/(H*(1-H) ))[-length(I)])/N)
 
     return(c(STAT, lambda))
-  } # End cvm.stat()
+  } # End cvm.stat.disc()
 
 
   type <- match.arg(type)
@@ -108,16 +103,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
[…]

### 2010-11-29 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@40 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 In their most basic forms, these nonparametric goodness-of-fit
-tests are intended for continuous null distributions, but they have
-also been adapted for discrete null distributions.  Unfortunately,
+tests are intended for continuous hypothesized distributions, but they have
+also been adapted for discrete distributions.  Unfortunately,
 most modern statistical software packages and programming environments
@@
 researchers would typically rely upon the $\chi^2$ test or
-use of a nonparametric test designed for a continuous null distribution.
+a nonparametric test designed for a continuous null distribution.
 For smaller sample sizes, in particular, both of these choices can produce
@@
 Given the cumulative distribution
-function $F_0(x)$ of the continuous null distribution
-and the empirical distribution function $F_{data}(x)$ of the observed data,  the test statistic is given by
+function $F_0(x)$ of the hypothesized distribution
+and the empirical distribution function $F_{data}(x)$ of the
+observed data,  the test statistic is given by
 \begin{align}
@@
 \end{align}
-The distribution of $D$ does not depend on the hypothesized
+When $F_0$ is continuous, 
+the distribution of $D$ does not depend on the hypothesized
 distribution, making this a computationally
@@
 (the `lesser' hypothesis alternative).  Tabulated p-values have been
-available for these tests since 1933.
+available for these tests since 1933.  **** CITATION? ***
 
@@
 is much more difficult to obtain; unlike the 
-continuous case, its distribution depends on the null model.
-It has been known since at least the 1950's that using the tables associated
-with continuous null distributions results in conservative p-values
+continuous case, it depends on the null model.
+Use of the tables associated
+with continuous hypothesized distributions results in conservative p-values
 when the null distribution is discontinuous 
@@
 
+*** Would a picture help at a point of discontinuity? ***
+
 If it is known that $g$ is a step function, it follows that
@@
 This, however, requires knowledge that $g$ is a step function as well as of
-the nature of its support (or break-points).  As a result,
+the nature of its support (specifically, the break-points).  As a result,
 we implement the Kolmogorov-Smirnov test statistic for discrete null
@@
 As with the original Kolmogorov-Smirnov test statistic, these all have 
-null distributions of their test statistics which are independent of the
-specified null models.
+test statistic null distributions which are independent of the
+hypothesized continuous models.
 
@@
 As with \code{ks.test()}, \code{cvm.test()} returns an object of class 
-\code{htest}.  Although we designed \code{cmv.test()} primarily for cases with
-discrete distributions \code{y}, we implemented the continuous forms of the
-test for variants \code{W2} and \code{A2} (as noted in the Discussion, below).
+\code{htest}.
 
@@
 
-*** possibly change following paragraph depending on what we
-decide to include in \code{cvm.test()}. ***
-
 Further 
 generalizations of goodness-of-fit tests for discrete distributions
-are described in the
-extended study of \cite{dewev1973}.  Our \code{cvm.test()} function
-also provides general one-sample
-Cram\'{e}r-von Mises testing for continuous distributions simply
-to help complement the functionality provided by other
-R packages, although these tests are not the focus of this paper.
+are described in the extended study of \cite{dewev1973}.  There are
+existing R packages for certain type of Cram\'{e}r-von Mises 
+goodness-of-fit tests for continuous distributions.
 Functions implemented in package \pkg{nortest} 

### 2010-11-29 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@41 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 (the `lesser' hypothesis alternative).  Tabulated p-values have been
-available for these tests since 1933.  **** CITATION? ***
+available for these tests since 1933 \citep{kol33}.
 
@@
 hypothesized continuous models.
-
 The $W^2$ statistic was the original test statistic.
@@
 
-It has been shown that these tests can be more powerful than Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
+It has been shown that these tests can be more powerful than 
+Kolmogorov-Smirnov tests to certain deviations from the hypothesized distribution. 
 They all involve integration over the whole range of data, rather than
@@
 rather than having large deviations over a small section of the support. 
-\cite{stephens1974} offers a complete analysis of the relative powers
+\cite{stephens1974} offers a comprehensive analysis of the relative powers
 of these tests.
@@
 \code{ks.test()} function, or if the original \code{stats::ks.test()}
-is used, the reported p-value is 0.1389.
+is used, the reported p-value is 0.1389 even though the test statistic
+is the same.
 

### 2010-11-29 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@42 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 alternative) or by replacing the supremum with a negative infimum
-(the `lesser' hypothesis alternative).  Tabulated p-values have been
+(the `lesser' hypothesis alternative).  Tabulated $p$-values have been
 available for these tests since 1933 \citep{kol33}.
@@
 Use of the tables associated
-with continuous hypothesized distributions results in conservative p-values
+with continuous hypothesized distributions results in conservative $p$-values
 when the null distribution is discontinuous 
@@
 \citet{Conover1972} developed the method implemented here
-for computing p-values
+for computing $p$-values
 in the case of discrete null distributions.
@@
 (corresponding to the desired one-sided or two-sided test).
-Then, the p-value for that particular test statistic may be computed.
+Then, the $p$-value for that particular test statistic may be computed.
 
@@
 
-*** Would a picture help at a point of discontinuity? ***
-
 If it is known that $g$ is a step function, it follows that
@@
 
-Having obtained the test statistic, the p-value must then be calculated. 
+Having obtained the test statistic, the $p$-value must then be calculated. 
 For larger sample sizes, the null distribution can be approximated
 with the null distribution from the classical Kolmogorov-Smirnov test. 
-When an exact p-value is required for
+When an exact $p$-value is required for
 smaller sample sizes, the methodology in \citet{Conover1972} is used. 
@@
 %The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that 
-%arise when calculated the p-values for larger sample sizes.
+%arise when calculated the $p$-values for larger sample sizes.
 
@@
 but numerical issues appear in the regime of large test statistics $x$
-(or, equivalently, small p-values).
+(or, equivalently, small $p$-values).
 The function $\theta(\cdot, x)$ is linear in $x$; as the test statistic 
@@
 and a chi-squared distribution of a higher degree of freedom. 
-The original formulation is preferable for most p-values, while
-this upper bound is used for smaller p-values (smaller than 0.001,
+The original formulation is preferable for most $p$-values, while
+this upper bound is used for smaller $p$-values (smaller than 0.001,
 based on our observations of the numerical instability of the original
@@
 instability.  Although it would be preferable to determine the use of
-the bound based on values of the test statistic rather than the p-value,
+the bound based on values of the test statistic rather than the $p$-value,
 the range of ``extreme'' values of the test statistic varies with the
@@
 \end{center}
-\caption{Plot of calculated p-values for given test statistics
+\caption{Plot of calculated $p$-values for given test statistics
 using numerical integration
@@
 The null distribution is uniform on the set $\{1,2,3\}$ in this example.
-The sharp variations in the calculated p-values are a result of numerical
-instabilities, and the true p-values are bounded by the blue dashed curve.}
+The sharp variations in the calculated $p$-values are a result of numerical
+instabilities, and the true $p$-values are bounded by the blue dashed curve.}
 \label{cvmissues}
@@
 warning unless \code{y} describes a discrete distribution.  When a discrete
-distribution is specified, exact p-values are not available for 
-two-sided alternative hypotheses, but the reported p-values will be
-conservative; very tight bounds for interesting (small) p-values are
+distribution is specified, exact $p$-values are not available for 
+two-sided alternative hypotheses, but the reported $p$-values will be
+conservative; very tight bounds for interesting (small) $p$-values are
 presented in \cite{Conover1972}.  For one-sided tests,
-exact p-values are calculated using Conover's
+exact $p$-values are calculated using Conover's
 method (when \code{exact = NULL} or \code{exact = TRUE})
@@
 in R (which, admittedly, doesn't claim to handle discrete distributions),
-the reported p-value, 0.5, is clearly incorrect:
+the reported $p$-value, 0.5, is clearly incorrect:
 \begin{Schunk}
@@
 data:  c(0, 1) 
-D = 0.5, p-value = 0.5
+D = 0.5, $p$-value = 0.5
 alternative hypothesis: two-sided 
@@
 Instead, the value of $D$ given in equation (1)
-should be 0 and the associated p-value should be 1.  Our revision of \code{ks.test()}
+should be 0 and the associated $p$-value should be 1.  Our revision of \code{ks.test()}
 fixes this problem when the user provides a discrete distribution:
@@
 data:  c(0, 1) 
-D = 0, p-value = 1
+D = 0, $p$-value = 1
 alternative hypothesis: two-sided 
@@
 new \code{ks.test()} implementation.  The first is the default two-sided test,
-where the reported p-value is a conservative upper bound for the actual
-p-value.  In this case, the approximation may not be that tight, but this is
-irrelevant for such large p-values. For more interesting (small) p-values, the
-upper bound is very close to the true p-value (shown in \cite{Conover1972}).
+where the reported $p$-value is a conservative upper bound for the actual
+$p$-value.  In this case, the approximation may not be that tight, but this is
+irrelevant for such large $p$-values. For more interesting (small) $p$-values, the
+upper bound is very close to the true $p$-value (shown in \cite{Conover1972}).
 \begin{Schunk}
@@
 data:  x 
-D = 0.08, p-value = 1
+D = 0.08, $p$-value = 1
 alternative hypothesis: two-sided 
@@
 Next, we conduct the default one-sided test, where Conover's method
-provides the exact p-value (up to the numerical precision of the
+provides the exact $p$-value (up to the numerical precision of the
 implementation):
@@
 data:  x 
-D^+ = 0.04, p-value = 0.7731
+D^+ = 0.04, $p$-value = 0.7731
 alternative hypothesis:
@@
 In contrast, the option \code{exact=FALSE} results in the
-p-value obtained by applying the classical Kolmogorov-Smirnov
-test, resulting in a conservative p-value:
+$p$-value obtained by applying the classical Kolmogorov-Smirnov
+test, resulting in a conservative $p$-value:
 \begin{Schunk}
@@
 data:  x 
-D^+ = 0.04, p-value = 0.9231
+D^+ = 0.04, $p$-value = 0.9231
 alternative hypothesis:
@@
 data:  rep(1, 3) 
-D = 0.6667, p-value = 0.04938
+D = 0.6667, $p$-value = 0.04938
 alternative hypothesis: two-sided 
@@
 \code{ks.test()} function, or if the original \code{stats::ks.test()}
-is used, the reported p-value is 0.1389 even though the test statistic
+is used, the reported $p$-value is 0.1389 even though the test statistic
 is the same.
@@
 data:  x 
-W2 = 0.057, p-value = 0.8114
+W2 = 0.057, $p$-value = 0.8114
 alternative hypothesis: Two.sided 
@@
 data:  x 
-A2 = 0.3969, p-value = 0.75
+A2 = 0.3969, $p$-value = 0.75
 alternative hypothesis: Two.sided 
@@
 data:  y 
-U2 = 0.0094, p-value = 0.945
+U2 = 0.0094, $p$-value = 0.945
 alternative hypothesis: Two.sided 
@@
 data:  z 
-U2 = 0.0094, p-value = 0.945
+U2 = 0.0094, $p$-value = 0.945
 alternative hypothesis: Two.sided 
@@
 produce different results after such a reordering. For example, the default
-Cram\'{e}r-von Mises test yields p-values of 0.8237 and 0.9577 with the original
+Cram\'{e}r-von Mises test yields $p$-values of 0.8237 and 0.9577 with the original
 and transformed data \code{y} and \code{z}, respectively.
@@
 %The computational capabilities of modern computers provide an alternative to using a complex formula to
-%calculate the p-value for a test statistic. By drawing random samples from the null distribution, in many
-%cases the p-values for a given statistic can be calculated accurately in a relatively short time span.
+%calculate the $p$-value for a test statistic. By drawing random samples from the null distribution, in many
+%cases the $p$-values for a given statistic can be calculated accurately in a relatively short time span.
 %This can be quite useful in some cases, but using the exact formulas from our methodology has several
@@
 goodness-of-fit tests for discrete null distributions.  In some cases,
-the p-values are known to be exact.  In others, conservativeness 
+the $p$-values are known to be exact.  In others, conservativeness 
 and tightness of bounds have been established in the literature.  As a result,
@@
 developed during an era when extensive simulations may have been
-prohibitively expensive or time-consuming).
+prohibitively expensive or time-consuming.
 However, this does raise the possibility that alternative tests relying upon

### 2011-04-23 pkg/ks.test/R/ks.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@54 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -69,33 +69,22 @@ ks.test <- function(x, y, ...,
             knots.y <- knots(y) 
             eps <- min(tol, min(diff(knots.y)) * tol)
             eps2 <- min(tol, min(diff(y(knots.y))) * tol)
-
-            a <- c()
-            b <- c()
-            f_a <- c()
-
+            a <- rep(0, n); b <- a; f_a <- a
             for (i in 1:n) {
-                a.1 <- which(y(knots.y) + S >= i/n + eps2)
-                if (sum(a.1) > 0) a <- c(a, knots.y[a.1[1]])
-                else a <- c(a, Inf) 
-    
-                b.1 <- which(y(knots.y) - S > (i-1)/n - eps2)
-                if (sum(b.1) > 0) b <- c(b, knots.y[b.1[1]])
-                else b <- c(b, Inf)
-    
+                a[i] <- min(c(knots.y[which(y(knots.y) + S >= i/n + eps2)[1]],
+                              Inf), na.rm=TRUE)
+                b[i] <- min(c(knots.y[which(y(knots.y) - S > (i-1)/n - eps2)[1]],
+                              Inf), na.rm=TRUE)
                 # Calculate F(a_i-)
                 # If a_i is not a knot of F_0,
                 # then simply return the value of F_0(a_i).
                 # Otherwise, evaluate F_0(a_i - eps)
-                if (!(a[i] %in% knots.y)) f_a <- c(f_a, y(a[i]))
-                else {
-                    f_a <- c(f_a, y(a[i]-eps))
-                }
+                f_a[i] <- ifelse(!(a[i] %in% knots.y),
+                                 y(a[i]), y(a[i]-eps))
             } 
             f_b <- y(b)
   
             # NOW HAVE f_a and f_b which are Niederhausen u, v
-
             p <- rep(1, n+1)
             for (i in 1:n) {
                 tmp <- 0

### 2011-08-24 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@58 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-11-29 rjournal/RJtemplate.tex
@@
 fill this void for researchers and practitioners in the R environment
-\citep{R2010}. 
+\citep{R2010}. This
+work was motivated by the need for such goodness-of-fit testing in a study
+of Olympic figure skating scoring \citep{emersonarnold2010}.  
 We first present overviews of the theory and general implementation of the
@@
 \citet{Conover1972} developed the method implemented here
-for computing $p$-values
+for computing exact one-sided $p$-values
 in the case of discrete null distributions.
+The method developed in \citet{gleser85} is used provide exact p-values
+for two-sided tests.
 
@@
 When an exact $p$-value is required for
-smaller sample sizes, the methodology in \citet{Conover1972} is used. 
+smaller sample sizes, the methodology in \citet{Conover1972} is used in
+for one-sided tests.  For two-sided tests, the methods presented in
+\citet{gleser85} lead to exact two-sided $p$-values.  This requires the
+calculation of rectangular probabilities for uniform order statistics
+as discussed by \citet{nieder81}.
 Full details of the calculations are contained in source code
-of our revised function \code{ks.test()} and in Conover's paper.
-
-%Numerical issues arise from the summation; need to do something else for high $n$. Simulation? Asymptotic approximation? 
-%The interesting part of the implementation of Conover's method, from a computational standpoint, are the difficult numerical issues that 
-%arise when calculated the $p$-values for larger sample sizes.
+of our revised function \code{ks.test()} and in the papers of
+Conover and Gleser.
 
@@
 \subsection{Overview}
-
-%General idea for continuous distributions. In many cases actually more powerful of the two. There again exists a distribution free 
-%distribution for the testing statistic.
 
@@
 The original test was developed by
-Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
+Harald Cram\'{e}r and Richard von Mises \citep{cramer1928, vonmises1928} 
+and further adapted by \cite{anderson1952}, and  \cite{Watson1961}. 
 The original test statistic, $W^2$, Anderson's $A^2$, and Watson's
@@
 \mathbb{P}\{Q \geq x \} = \frac{1}{2} + 
-\frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du
-\label{distQ}
+\frac{1}{\pi} \int_{0}^{\infty} \frac{\sin\theta(u,x)}{u \rho(u) } du \label{distQ}
 \end{align}
@@
 uniform on the set $\{1,2,3\}$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.4]{fig1.pdf}
+\end{center}
+\caption{Plot of calculated $p$-values for given test statistics
+using numerical integration
+(red) compared to the conservative chi-squared bound (dashed blue). 
+The null distribution is uniform on the set $\{1,2,3\}$ in this example.
+The sharp variations in the calculated $p$-values are a result of numerical
+instabilities, and the true $p$-values are bounded by the blue dashed curve.}
+\label{cvmissues}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.4]{fig1.pdf}
-\end{center}
-\caption{Plot of calculated $p$-values for given test statistics
-using numerical integration
-(red) compared to the conservative chi-squared bound (dashed blue). 
-The null distribution is uniform on the set $\{1,2,3\}$ in this example.
-The sharp variations in the calculated $p$-values are a result of numerical
-instabilities, and the true $p$-values are bounded by the blue dashed curve.}
-\label{cvmissues}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Kolmogorov-Smirnov and Cram\'{e}r-von Mises Tests in R}
@@
 or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
-\code{ks.test()} will be proposed for inclusion in \pkg{stats} in late 2010.
+\code{ks.test()} will be proposed for inclusion in recommended
+package \pkg{stats}.
 
@@
 the presence of ties in the data (the first argument, \code{x}) generates a
-warning unless \code{y} describes a discrete distribution.  When a discrete
-distribution is specified, exact $p$-values are not available for 
-two-sided alternative hypotheses, but the reported $p$-values will be
-conservative; very tight bounds for interesting (small) $p$-values are
-presented in \cite{Conover1972}.  For one-sided tests,
-exact $p$-values are calculated using Conover's
-method (when \code{exact = NULL} or \code{exact = TRUE})
-if the sample size is less than or equal to 30; otherwise, asymptotic distributions
+warning unless \code{y} describes a discrete distribution.  
+If the sample size is less than or equal to 30, exact $p$-values are provided;
+otherwise, asymptotic distributions
 are used which are reliable in such cases.
@@
 for details).
-
-%*** Discussion of what we might have broken: cases where the user
-%provided a discrete distribution to the original ks.test() even though
-%it wasn't intended to be supported. ***
 
@@
 
-%%%%%%%%%%%%%%%%%%
 \section{Examples}
@@
 Instead, the value of $D$ given in equation (1)
-should be 0 and the associated $p$-value should be 1.  Our revision of \code{ks.test()}
+should be 0 and the associated $p$-value should be 1.  
+Our revision of \code{ks.test()}
 fixes this problem when the user provides a discrete distribution:
@@
 new \code{ks.test()} implementation.  The first is the default two-sided test,
-where the reported $p$-value is a conservative upper bound for the actual
-$p$-value.  In this case, the approximation may not be that tight, but this is
-irrelevant for such large $p$-values. For more interesting (small) $p$-values, the
-upper bound is very close to the true $p$-value (shown in \cite{Conover1972}).
+where the exact $p$-value is obtained using the methods of \citet{gleser85}.
 \begin{Schunk}
@@
 data:  x 
-D = 0.08, $p$-value = 1
+D = 0.08, $p$-value = 0.9354
 alternative hypothesis: two-sided 
@@
 data:  rep(1, 3) 
-D = 0.6667, $p$-value = 0.04938
+D = 0.6667, $p$-value = 0.07407
 alternative hypothesis: two-sided 
@@
 
-%*** I commented out a bunch of stuff.  I think the discussion can be short
-%and sweet. ***
-
-%The chi-squared test is a popular alternative to the methods presented here for doing goodness-of-fit tests for discrete data.
-%It differs by using no information about the geometry of the distribution, instead relying only the difference between the 
-%number of data points observed at each value and the expected number of data points that should be observed at each point. 
-%This generally has the effect of having a greater power towards alternatives with sharply different densities on a few points,
-%at the expense of losing power for alternatives with slight deviations over all of the points. Also, the non-parametric tests
-%have an advantage of being useful when the sample size is lower compared to the support of the distribution. The chi-squared
-%test is generally not used unless there are at least five data points in each element of the support. On the other hand,
-%chi-squared can be used in situations when there is not a numerical interpretation of the observation space 
-%(e.g. a set of ethnicities).
-%For a more complete comparison of the two tests' power see \cite{slakter65}.
-
-%The computational capabilities of modern computers provide an alternative to using a complex formula to
-%calculate the $p$-value for a test statistic. By drawing random samples from the null distribution, in many
-%cases the $p$-values for a given statistic can be calculated accurately in a relatively short time span.
-%This can be quite useful in some cases, but using the exact formulas from our methodology has several
-%distinct advantages. First of all, if the function is called many times (say, within another simulation
-%study), the computational benefits of using a formula can quickly become substantial. There is also a greater need
-%for supervision in a simulation study, with the exact number of runs needed to reach convergence (or other 
-%similar threshold). Additionally, while
-%simulations are generally a well received accepted method, it is often important in applied data analysis
-%to be sure that the calculated $p$-values are not the result of a numerical oddity in a particular
-%run of a simulation.
-
-%*** Hmm... if you claim the following, you want to make sure it is true.
-%Did you read it somewhere? ***
-
 This paper presents the implementation of several nonparametric
-goodness-of-fit tests for discrete null distributions.  In some cases,
-the $p$-values are known to be exact.  In others, conservativeness 
-and tightness of bounds have been established in the literature.  As a result,
+goodness-of-fit tests for discrete null distributions.  
+In some cases the $p$-values are known to be exact.  In others,
+conservativeness in special cases with small $p$-values
+has been established.  As a result,
 no simulations are necessary for these methods, which were generally
@@
 
-%\bibliography{example}
+\section{Acknowledgement}
 
-\bibliography{paper}
+We are grateful to Le-Minh Ho for his implementation of the algorithm
+described in \citet{nieder81} for calculating rectangular probabilities
+of uniform order statistics.
+
+\bibliography{ArnoldEmerson}
 

### 2011-08-24 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@59 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \code{ks.test()} function and a new \code{cvm.test()} function to
-fill this void for researchers and practitioners in the R environment
-\citep{R2010}. This
+fill this void for researchers and practitioners in the R environment. This
 work was motivated by the need for such goodness-of-fit testing in a study
@@
 Having obtained the test statistic, the $p$-value must then be calculated. 
-For larger sample sizes, the null distribution can be approximated
+For larger sample sizes, the null distribution
+of the test statistic can be approximated
 with the null distribution from the classical Kolmogorov-Smirnov test. 
@@
 A different toy example shows the dangers of using R's existing
-\code{ks.test()} function:
+\code{ks.test()} function with discrete data:
 \begin{Schunk}

### 2011-08-24 pkg/ks.test/R/ks.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@60 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -56,11 +56,12 @@ ks.test <- function(x, y, ...,
     }
 
     #############################################################
-    ## TBA, JWE: conover.pval is a new function implementing the
-    ## p-value calculations presented in Conover (1972), which is
-    ## cited in our revision of ks.test.Rd.
+    ## TBA, JWE: exact.pval is a new function implementing the
+    ## p-value calculations presented in Conover (1972) and
+    ## Gleser (1985), which are cited in our revision of
+    ## ks.test.Rd.
 
-    conover.pval <-  function(alternative, STATISTIC, x, z, n, y, tol) {
+    exact.pval <-  function(alternative, STATISTIC, x, z, n, y, tol) {
     
         ts.pval <- function(S, x, z, n, H, tol) {
             # The exact two-sided p-value from Gleser (1985)
@@ -85,6 +86,8 @@ ks.test <- function(x, y, ...,
             f_b <- y(b)
   
             # NOW HAVE f_a and f_b which are Niederhausen u, v
+            # which uses the Noe result.
+
             p <- rep(1, n+1)
             for (i in 1:n) {
                 tmp <- 0
@@ -190,15 +193,19 @@ ks.test <- function(x, y, ...,
 
     } else if (is.stepfun(y)) {
         z <- knots(y)
+
+    ## JE NOTE: could allow larger n for two-sided case???
+    ## revisit.
+
         if(is.null(exact) || exact) exact <- (n <= 30)
         METHOD <- "One-sample Kolmogorov-Smirnov test"
-        dev <- c(0, ecdf(x)(z) - y(z))      # JE had been blowing away data
+        dev <- c(0, ecdf(x)(z) - y(z))
         STATISTIC <- switch(alternative,
                             "two.sided" = max(abs(dev)),
                             "greater" = max(dev),
                             "less" = max(-dev))
         PVAL <- switch(exact,
-                       "TRUE" = conover.pval(alternative, STATISTIC,
+                       "TRUE" = exact.pval(alternative, STATISTIC,
                                              x, z, n, y, tol),
                        "FALSE" = NULL)
         nm_alternative <- switch(alternative,

### 2011-08-24 pkg/ks.test/R/ks.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@61 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -40,7 +40,7 @@ ks.test <- function(x, y, ...,
                     tol=1e-8)
 
     ## END TBA, JWE
-    ###############
+    ################
 
 {
 

### 2011-08-24 pkg/ks.test/R/ks.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@62 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -37,7 +37,11 @@ ks.test <- function(x, y, ...,
     ## for equality (a==b).  We also use it in establishing a small epsilon
     ## for some other calculations involving the step functions.
 
-                    tol=1e-8)
+    ## Additional arguments, simulate.p.value and B, follow the usage of
+    ## fisher.test, but are only used for the discrete distribution case.
+
+                    tol=1e-8,
+                    simulate.p.value=FALSE, B=2000)
 
     ## END TBA, JWE
     ################
@@ -141,6 +145,29 @@ ks.test <- function(x, y, ...,
         return(p)
     }
 
+    sim.pval <-  function(alternative, STATISTIC, x, z, n, y, tol, B) {
+        # Simulate B samples from given stepfun y
+        knots.y <- knots(y)
+        fknots.y <- y(knots.y)
+        u <- runif(B*length(x))
+        u <- sapply(u, function(a) return(knots.y[sum(a>fknots.y)+1]))
+        dim(u) <- c(B, length(x))
+        
+        # Calculate B values of the test statistic
+        getks <- function(a, knots.y, fknots.y) {
+            dev <- c(0, ecdf(a)(knots.y) - fknots.y)
+            STATISTIC <- switch(alternative,
+                              "two.sided" = max(abs(dev)),
+                              "greater" = max(dev),
+                              "less" = max(-dev))
+            return(STATISTIC)
+        }
+        s <- apply(u, 1, getks, knots.y, fknots.y)
+
+        # Produce the estimated p-value
+        return(sum(s >= STATISTIC-tol) / B)
+    }
+
     ## END TBA, JWE
     ###############
 
@@ -191,6 +218,8 @@ ks.test <- function(x, y, ...,
     ## a new possibility for y: an object of class 'stepfun', of
     ## which 'ecdf' is a subclass we expect most users to have.
 
+    ## Don't recalculate knots(z) more than necessary.
+
     } else if (is.stepfun(y)) {
         z <- knots(y)
 
@@ -204,10 +233,17 @@ ks.test <- function(x, y, ...,
                             "two.sided" = max(abs(dev)),
                             "greater" = max(dev),
                             "less" = max(-dev))
-        PVAL <- switch(exact,
-                       "TRUE" = exact.pval(alternative, STATISTIC,
-                                             x, z, n, y, tol),
-                       "FALSE" = NULL)
+        if (simulate.p.value) {
+            PVAL <- switch(exact,
+                           "TRUE" = sim.pval(alternative, STATISTIC,
[…]

### 2011-08-24 rjournal/code_taylor.R

@@ -1,0 +1,115 @@
+# Function to get lambdas:
+  cvm.stat.disc <- function(x,y, type=c("W2", "U2", "A2")) {
+    type <- match.arg(type)
+    I <- knots(y)
+    N <- length(x)
+    e <- diff(c(0,N*y(I)))
+    obs <- rep(0, length(I))
+    for(j in 1:length(I)) {
+      obs[j] <- length(which(x == I[j]))
+    }
+    S <- cumsum(obs)
+    T <- cumsum(e)
+    H <- T/N
+    p <- e/N
+    t <- (p + p[c(2:length(p), 1)])/2
+    Z <- S - T
+    Zbar <- sum(Z*t)
+
+    S0 <- diag(p) - p %*% t(p)
+    A <- matrix(1, length(p), length(p))
+    A <- apply(row(A) >= col(A),2, as.numeric)
+    E <- diag(t)
+    One <- rep(1, nrow(E))
+    K <- diag(0, length(H))
+    diag(K)[-length(H)] <- 1/(H[-length(H)]*(1-H[-length(H)]))
+    Sy <- A %*% S0 %*% t(A)
+    M <- switch(type, W2 = E,
+                U2 = (diag(1, nrow(E)) - 
+                      E%*%One%*%t(One))%*%E%*%(diag(1, nrow(E)) -
+                      One%*%t(One)%*%E),
+                A2 = E%*%K)
+    lambda <- eigen(M%*%Sy)$values
+
+    STAT <- switch(type, W2 = sum(Z^2*t)/N, U2 = sum((Z-Zbar)^2*t)/N,
+                   A2 = sum((Z^2*t/(H*(1-H) ))[-length(I)])/N)
+
+    return(c(STAT, lambda))
+  } # End cvm.stat.disc()
+
+
+logf <- function(t, x, lambda) {
+  ans <- -t*x
+  ans <- ans - 0.5*sum( log(1-2*t*lambda) )
+  return(ans)
+}
+
+x <- 15
+lambda <- rep(1,3)
+tvals <- seq(0.00001, 1/(2*max(lambda)), length.out=100)
+vals <- exp(sapply(tvals, logf, x, lambda))
+plot(tvals, vals)
+ans1 <- min(vals[is.finite(vals)])
+df <- sum(lambda != 0)
+ans2 <- dchisq(x/max(lambda),df)
+
[…]

### 2011-08-25 pkg/cvm.test/R/cvm.test.R
this work
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@63 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -45,7 +45,14 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
     if(pval > 0.001) return(pval)
     if(pval <= 0.001) {
       df <- sum(lambda != 0)
-      return(dchisq(STAT/max(lambda),df))
+      est1 <- dchisq(STAT/max(lambda),df)      
+      logf <- function(t) {
+        ans <- -t*STAT
+        ans <- ans - 0.5*sum( log(1-2*t*lambda) )
+        return(ans)
+      }
+      est2 <- optimize(logf, interval=c(0,1/(2*max(lambda))))[[1]]
+      return(min(est1,est2))
     }
   } # End cvm.pval.disc()
 

### 2011-08-25 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@62 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@64 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Having obtained the test statistic, the $p$-value must then be calculated. 
-For larger sample sizes, the null distribution
-of the test statistic can be approximated
-with the null distribution from the classical Kolmogorov-Smirnov test. 
 When an exact $p$-value is required for
@@
 Conover and Gleser.
+
+For larger sample sizes, the classical Kolmogorov-Smirnov test is used
+and is known to produce conservative p-values for discrete distributions;
+the revised \code{ks.test()} supports estimation
+of p-values via simulation if desired.
 
@@
 
-We resolve this problem by using a simple conservative approximation to
-avoid the numerical instability. Consider the
+We resolve this problem by using a combination of
+two simple conservative approximations to
+avoid the numerical instability. First, consider the
 following inequality:
@@
 The values for the weighted sum can be bounded using a simple transformation
-and a chi-squared distribution of a higher degree of freedom. 
-The original formulation is preferable for most $p$-values, while
-this upper bound is used for smaller $p$-values (smaller than 0.001,
+and a chi-squared distribution of a higher degree of freedom.  Second,
+consider the Markov inequality: 
+\begin{align}
+\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{E}
+  \left[\exp\left(t\sum_{i=1}^{p}\lambda_iZ_i^2\right)\right]\exp(-tx) \\
+&= \frac{\exp(-tx)}{\sqrt{\prod_{i=1}^{p}(1-2t\lambda_i)}}
+\label{ineq3}
+\end{align}
+where the bound can be
+minimized over $t \in (0,1/2\lambda_{max})$.
+The upper bounds for the $p$-value given by (\ref{ineq2}) and (\ref{ineq3})
+are both calculated and the smaller is used in cases where the numerical
+instability of (\ref{distQ}) is a concern.
+
+The original formulation, numerical integration of (\ref{distQ}),
+ is preferable for most $p$-values, while
+the upper bound described above
+is used for smaller $p$-values (smaller than 0.001,
 based on our observations of the numerical instability of the original
-formulation).  Figure \ref{cvmissues} shows this bound with the blue dashed
-line; values in red exceeding the bound are a result of the numerical
+formulation).  Figure \ref{cvmissues} shows the bounds with the blue and green
+dashed lines; values in red exceeding the bounds are a result of the numerical
 instability.  Although it would be preferable to determine the use of
@@
 \section{Examples}
+
+NOTE: Add an example or two with simulated p-values.  Remember to add
+stats:: everywhere appropriate for clarity, and ks.test::, etc...
 

### 2011-08-25 pkg/ks.test/R/ks.test.R

@@ -65,13 +65,12 @@ ks.test <- function(x, y, ...,
     ## Gleser (1985), which are cited in our revision of
     ## ks.test.Rd.
 
-    exact.pval <-  function(alternative, STATISTIC, x, z, n, y, tol) {
+    exact.pval <-  function(alternative, STATISTIC, x, n, y, knots.y, tol) {
     
-        ts.pval <- function(S, x, z, n, H, tol) {
+        ts.pval <- function(S, x, n, y, knots.y, tol) {
             # The exact two-sided p-value from Gleser (1985)
             # and Niederhausen (1981)
             f_n <- ecdf(x)
-            knots.y <- knots(y) 
             eps <- min(tol, min(diff(knots.y)) * tol)
             eps2 <- min(tol, min(diff(y(knots.y))) * tol)
             a <- rep(0, n); b <- a; f_a <- a
@@ -89,8 +88,8 @@ ks.test <- function(x, y, ...,
             } 
             f_b <- y(b)
   
-            # NOW HAVE f_a and f_b which are Niederhausen u, v
-            # which uses the Noe result.
+            # NOW HAVE f_a and f_b which are Niederhausen u, v, and this
+            # uses the Noe result.
 
             p <- rep(1, n+1)
             for (i in 1:n) {
@@ -101,11 +100,16 @@ ks.test <- function(x, y, ...,
                 }
                 p[i+1] <- tmp
             }	
-            p <- max(0, 1 - p[n+1])          # Use Niederhausen result
+            p <- max(0, 1 - p[n+1])          # Niederhausen result
+            if (p>1) {
+                warning("numerical instability in p-value calculation.")
+                p <- 1
+            }
             return(p)
         }
 
-        less.pval <- function(S, z, n, H, tol) {
+        # Conover
+        less.pval <- function(S, n, H, z, tol) {
             m <- ceiling(n*(1-S))
             c <- S+(1:m-1)/n
             CDFVAL <- H(sort(z))
@@ -122,7 +126,8 @@ ks.test <- function(x, y, ...,
             return(p)        
         }
 
-        greater.pval <- function(S, z, n, H, tol) {
+        # Conover
+        greater.pval <- function(S, n, H, z, tol) {
             m <- ceiling(n*(1-S))
             c <- 1-(S+(1:m-1)/n)
             CDFVAL <- c(0,H(sort(z)))
@@ -139,15 +144,14 @@ ks.test <- function(x, y, ...,
         }
 
         p <- switch(alternative,
[…]

### 2011-08-25 pkg/cvm.test/R/cvm.test.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@65 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -51,7 +51,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
         ans <- ans - 0.5*sum( log(1-2*t*lambda) )
         return(ans)
       }
-      est2 <- optimize(logf, interval=c(0,1/(2*max(lambda))))[[1]]
+      est2 <- exp(nlm(logf, 1/(4*max(lambda)))$minimum)
       return(min(est1,est2))
     }
   } # End cvm.pval.disc()

### 2011-08-25 rjournal/codeforgraphics.R

@@ -74,12 +74,13 @@
 
 #lambda <- sort(runif(10))
 #lambda <- sort(c(0.020837 , 0.342575 , 0.531838 , 0.604825 , 0.645453 , 0.706874 , 0.750545 , 0.776914 , 0.810034 , 0.880271))
-y <- ecdf(1:4)
-x <- sample(1:4)
+set.seed(42)
+y <- ecdf(1:3)
+x <- sample(1:3, replace=TRUE)
 lambda <- cvm.stat.disc(x,y,'W2')[-1]
 cvm.pval(1, lambda)
 
-I <- seq(0,5,length.out=500)
+I <- seq(0,5,length.out=30)
 y <- rep(NA, length(I))
 for(i in 1:length(I)) {
   y[i] <- cvm.pval(I[i], lambda, good=TRUE)
@@ -88,6 +89,20 @@ z <- rep(NA, length(I))
 for(i in 1:length(I)) {
   z[i] <- cvm.pval(I[i], lambda, good=FALSE)
 }
+w <- rep(NA, length(I))
+for(i in 1:length(I)) {
+  STAT <- I[i]
+  logf <- function(t) {
+    ans <- -t*STAT
+    ans <- ans - 0.5*sum( log(1-2*t*lambda) )
+    return(ans)
+  }
+  est2 <- exp(nlm(logf, 1/(4*max(lambda)))$minimum)
+  #est2 <- exp(optimize(logf, interval=c(0,1/(2*max(lambda))))[[1]])
+  w[i] <- est2
+}
+
+
 
 #pdf('fig1.pdf', height=6, width=8)
 coff <- 1.2
@@ -95,7 +110,7 @@ coff2 <- 1.3
 plot(I[I > coff], z[I > coff], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
 abline(h=0, lty=2)
 lines(I[I > coff2], y[I > coff2], lty=2, col='blue', lwd=2)
-lines(I[I > coff], z[I > coff], col='salmon', lwd=2)
+lines(I[I > coff], z[I > coff], col='green', lwd=2)
 #dev.off()
 
 

### 2011-08-25 rjournal/codeforgraphics.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@66 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -80,7 +80,7 @@ x <- sample(1:3, replace=TRUE)
 lambda <- cvm.stat.disc(x,y,'W2')[-1]
 cvm.pval(1, lambda)
 
-I <- seq(0,5,length.out=30)
+I <- seq(0,5,length.out=500)
 y <- rep(NA, length(I))
 for(i in 1:length(I)) {
   y[i] <- cvm.pval(I[i], lambda, good=TRUE)
@@ -110,7 +110,7 @@ coff2 <- 1.3
 plot(I[I > coff], z[I > coff], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
 abline(h=0, lty=2)
 lines(I[I > coff2], y[I > coff2], lty=2, col='blue', lwd=2)
-lines(I[I > coff], z[I > coff], col='green', lwd=2)
+lines(I[I > coff], w[I > coff], col='green', lwd=2)
 #dev.off()
 
 

### 2011-08-25 code.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@67 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,7 +1,7 @@
 N <- seq(10, 500, by = 20)
 P <- seq(10,200, by=10)
 
-N <- seq(60, 85, by = 2)
+N <-    seq(60, 85, by = 2)
 P <- seq(10,30, by=1)
 P <- rep(25, 15)
 

### 2011-08-25 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@68 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 warning unless \code{y} describes a discrete distribution.  
-If the sample size is less than or equal to 30, exact $p$-values are provided;
-otherwise, asymptotic distributions
-are used which are reliable in such cases.
-When \code{exact = FALSE} the asymptotic distribution is used and known
-to be conservative though imprecise for small samples (see \cite{Conover1972}
-for details).
+If the sample size is less than or equal to 30, or when \code{exact=TRUE},
+exact $p$-values are provided (a warning is issued when the sample size is
+greater than 30 due to possible numerical instabilities).
+When \code{exact = FALSE} (or when \code{exact} is unspecified and the sample
+size is greater than 30) the classical Kolmogorov-Smirnov null distribution
+of the test statistic is used and resulting $p$-values are known
+to be conservative though imprecise (see \cite{Conover1972}
+for details).  In such cases, simulated $p$-values may be desired, produced
+by the \code{simulate.p.value=TRUE} option.
 
@@
 \section{Examples}
-
-NOTE: Add an example or two with simulated p-values.  Remember to add
-stats:: everywhere appropriate for clarity, and ks.test::, etc...
 
@@
 > library(ks.test)
-> ks.test(c(0, 1), ecdf(c(0, 1)))
+> ks.test::ks.test(c(0, 1), ecdf(c(0, 1)))
 \end{Sinput}
@@
 \begin{Sinput}
-> ks.test(x, ecdf(1:10))
+> ks.test::ks.test(x, ecdf(1:10))
 \end{Sinput}
@@
 \begin{Sinput}
-> ks.test(x, ecdf(1:10), alternative = "g")
+> ks.test::ks.test(x, ecdf(1:10),
++   alternative = "g")
 \end{Sinput}
@@
 \begin{Sinput}
-> ks.test(x, ecdf(1:10), alternative = "g", 
-+         exact = FALSE)
+> ks.test::ks.test(x, ecdf(1:10),
++   alternative = "g", exact = FALSE)
 \end{Sinput}
@@
 \end{Schunk}
+The $p$-value may also be estimated via a Monte Carlo simulation:
+\begin{Schunk}
+\begin{Sinput}
+> ks.test::ks.test(x, ecdf(1:10),
++   alternative = "g",
++   simulate.p.value=TRUE, B=10000)
+\end{Sinput}
+\begin{Soutput}
+	One-sample Kolmogorov-Smirnov test
+
+data:  x 
+D^+ = 0.04, p-value = 0.7717
+alternative hypothesis: 
+the CDF of x lies above the null hypothesis 
+\end{Soutput}
+\end{Schunk}
 
@@
 \begin{Sinput}
-> ks.test(rep(1, 3), ecdf(1:3))
+> ks.test::ks.test(rep(1, 3), ecdf(1:3))
 \end{Sinput}
@@
 conservativeness in special cases with small $p$-values
-has been established.  As a result,
-no simulations are necessary for these methods, which were generally
+has been established.  Although we provide for Monte Carlo simulated
+$p$-values with the new \code{ks.test()},
+no simulations may be necessary necessary for these methods; they
+were generally
 developed during an era when extensive simulations may have been
@@
 described in \citet{nieder81} for calculating rectangular probabilities
-of uniform order statistics.
+of uniform order statistics.  We also appreciate the feedback from two
+reviewers of this paper as well as from a reviewer of 
+\cite{emersonarnold2010}.
 

### 2011-08-25 pkg/dgof/R/cvm.test.R

@@ -1,0 +1,120 @@
+cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
+
+  #cvm.stat <- function(x,y,type) {
+  #  fdata <- ecdf(x)
+  #  if(type != 'A2') {
+  #    igrand <- function(z) return((fdata(z) - y(z))^2)
+  #  } else {
+  #    igrand <- function(z) return((fdata(z) - y(z))^2 / (y(z) - y(z)^2))
+  #  }
+  #  stat <- integrate(igrand, -Inf, Inf)[1]
+  #  if(type != 'U2') {
+  #    return(stat)
+  #  } else {
+  #    igrand <- function(z) return((fdata(z) - y(z)-stat)^2) 
+  #    return(integrate(igrand, -Inf, Inf)[1])
+  #  }    
+  #}
+
+  cvm.pval.disc <- function(STAT, lambda) {
+
+    x <- STAT
+
+    theta <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+         VAL <- VAL + 0.5*atan(lambda[i]*u)
+      }
+      return(VAL - 0.5*x*u)
+    }
+
+    rho <- function(u) {
+      VAL <- 0
+      for(i in 1:length(lambda)) {
+        VAL <- VAL + log(1 + lambda[i]^2*u^2)     
+      }
+      VAL <- exp(VAL*0.25)
+      return(VAL)
+    }
+
+    fun <- function(u) return(sin(theta(u))/(u*rho(u)))
+
+    pval <- 0
+    try(pval <- 0.5 + integrate(fun, 0, Inf, subdivisions=1e6)$value/pi,
+        silent=TRUE)
+    if(pval > 0.001) return(pval)
+    if(pval <= 0.001) {
+      df <- sum(lambda != 0)
+      est1 <- dchisq(STAT/max(lambda),df)      
+      logf <- function(t) {
+        ans <- -t*STAT
+        ans <- ans - 0.5*sum( log(1-2*t*lambda) )
+        return(ans)
+      }
+      est2 <- exp(nlm(logf, 1/(4*max(lambda)))$minimum)
+      return(min(est1,est2))
+    }
+  } # End cvm.pval.disc()
+
+  cvm.stat.disc <- function(x,y, type=c("W2", "U2", "A2")) {
+    type <- match.arg(type)
+    I <- knots(y)
+    N <- length(x)
+    e <- diff(c(0,N*y(I)))
[…]

### 2011-08-25 pkg/dgof/R/ks.test.R

@@ -1,0 +1,352 @@
+#  Original file .../src/library/stats/R/ks.test.R
+#  Part of the R package, http://www.R-project.org
+#
+#  This program is free software; you can redistribute it and/or modify
+#  it under the terms of the GNU General Public License as published by
+#  the Free Software Foundation; either version 2 of the License, or
+#  (at your option) any later version.
+#
+#  This program is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#
+#  A copy of the GNU General Public License is available at
+#  http://www.r-project.org/Licenses/
+#
+# Modification by Taylor Arnold and Jay Emerson, 2010:
+#
+# Any changes to the original R Core source is noted by
+# the following comment convension to aid code review:
+#
+#   ####################
+#   ## TBA, JWE: comment
+#
+#   ## END TBA, JWE
+#   ###############
+
+
+
+ks.test <- function(x, y, ...,
+                    alternative = c("two.sided", "less", "greater"),
+                    exact = NULL,
+
+    #########################################################################
+    ## TBA, JWE: an additional argument, tol, is used as an upper bound for
+    ## possible rounding error in values (say, a and b) when needing to check
+    ## for equality (a==b).  We also use it in establishing a small epsilon
+    ## for some other calculations involving the step functions.
+
+    ## Additional arguments, simulate.p.value and B, follow the usage of
+    ## fisher.test, but are only used for the discrete distribution case.
+
+                    tol=1e-8,
+                    simulate.p.value=FALSE, B=2000)
+
+    ## END TBA, JWE
+    ################
+
+{
+
+    pkolmogorov1x <- function(x, n) {
+        ## Probability function for the one-sided one-sample Kolmogorov
+        ## statistics, based on the formula of Birnbaum & Tingey (1951).
+        if(x <= 0) return(0)
+        if(x >= 1) return(1)
+        j <- seq.int(from = 0, to = floor(n * (1 - x)))
+        1 - x * sum(exp(lchoose(n, j)
+                        + (n - j) * log(1 - x - j / n)
+                        + (j - 1) * log(x + j / n)))
+    }
+
[…]

### 2011-08-25 rjournal/codeforgraphics.R
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@69 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -106,11 +106,15 @@ for(i in 1:length(I)) {
 
 #pdf('fig1.pdf', height=6, width=8)
 coff <- 1.2
-coff2 <- 1.3
-plot(I[I > coff], z[I > coff], type='l', xlab='test statistic', ylab='p-value', col='salmon', lwd=2)
+coff2 <- 3.0
+plot(I[I > coff & I < coff2], z[I > coff & I < coff2], type='l', xlim=c(coff, coff2),
+    xlab='test statistic', ylab='p-value', col='tomato1', lwd=2)
 abline(h=0, lty=2)
-lines(I[I > coff2], y[I > coff2], lty=2, col='blue', lwd=2)
-lines(I[I > coff], w[I > coff], col='green', lwd=2)
+lines(I[I > 1.3], y[I > 1.3], lty=2, col='steelblue', lwd=3)
+lines(I[I > coff], w[I > coff], lty=2, col='olivedrab4', lwd=3)
+legend(2.25, 0.015, legend=c('Asymptotic p-value (9)',
+                             'Bound given by (11)','Bound given by (13)' ), 
+     col=c('tomato1','steelblue', 'olivedrab4'), lty=c(1,2,2))
 #dev.off()
 
 

### 2011-08-25 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@72 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 using numerical integration
-(red) compared to the conservative chi-squared bound (dashed blue). 
+(red) compared to the conservative chi-squared bound (dashed blue) and
+the Markov inequality bound (dashed green). 
 The null distribution is uniform on the set $\{1,2,3\}$ in this example.
 The sharp variations in the calculated $p$-values are a result of numerical
-instabilities, and the true $p$-values are bounded by the blue dashed curve.}
+instabilities, and the true $p$-values are bounded by the dashed curves.}
 \label{cvmissues}
@@
 Functions \code{ks.test()} and \code{cvm.test()} are provided for
-convenience in packages \pkg{ks.test} and \pkg{cvm.test}, respectively.
+convenience in package \pkg{dgof}, available on CRAN.
 Function \code{ks.test()} offers a revision of
-R's Kolmogorov-Smirnov function \code{ks.test()} from base
-package \pkg{stats}, while \code{cvm.test()} is a new
+R's Kolmogorov-Smirnov function \code{ks.test()} from recommended
+package \pkg{stats}; \code{cvm.test()} is a new
 function for Cram\'{e}r-von Mises tests.
-Both are available from the authors
-or from R-Forge (\url{https://r-forge.r-project.org/R/?group_id=802});
-\code{ks.test()} will be proposed for inclusion in recommended
-package \pkg{stats}.
 

### 2011-08-25 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@73 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-For larger sample sizes, the classical Kolmogorov-Smirnov test is used
+For larger sample sizes (or when requested for smaller sample sizes),
+the classical Kolmogorov-Smirnov test is used
 and is known to produce conservative p-values for discrete distributions;
@@
 \section{Cram\'{e}r-von Mises Tests}
+
+NOTE: NEED TO ADD SIMUALTED P-VALUES.
 
@@
 > library(ks.test)
-> ks.test::ks.test(c(0, 1), ecdf(c(0, 1)))
+> dgof::ks.test(c(0, 1), ecdf(c(0, 1)))
 \end{Sinput}
@@
 \begin{Sinput}
-> ks.test::ks.test(x, ecdf(1:10))
+> dgof::ks.test(x, ecdf(1:10))
 \end{Sinput}
@@
 \begin{Sinput}
-> ks.test::ks.test(x, ecdf(1:10),
+> dgof::ks.test(x, ecdf(1:10),
 +   alternative = "g")
@@
 \begin{Sinput}
-> ks.test::ks.test(x, ecdf(1:10),
+> dgof::ks.test(x, ecdf(1:10),
 +   alternative = "g", exact = FALSE)
@@
 \begin{Sinput}
-> ks.test::ks.test(x, ecdf(1:10),
+> dgof::ks.test(x, ecdf(1:10),
 +   alternative = "g",
@@
 \begin{Sinput}
-> ks.test::ks.test(rep(1, 3), ecdf(1:3))
+> dgof::ks.test(rep(1, 3), ecdf(1:3))
 \end{Sinput}
@@
 Package \pkg{cramer} \citep{cramer} offers a multivariate Cram\'{e}r test for
-the two-sample problem.
+the two-sample problem.  Finally, we note that the discrete goodness-of-fit tests
+discussed in this paper do not allow the estimation of parameters of the
+hypothesized null distributions (see \cite{LSS2007} for example).
 

### 2011-08-27 rjournal/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@74 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 data:  c(0, 1) 
-D = 0.5, $p$-value = 0.5
+D = 0.5, p-value = 0.5
 alternative hypothesis: two-sided 
@@
 data:  c(0, 1) 
-D = 0, $p$-value = 1
+D = 0, p-value = 1
 alternative hypothesis: two-sided 
@@
 data:  x 
-D = 0.08, $p$-value = 0.9354
+D = 0.08, p-value = 0.9354
 alternative hypothesis: two-sided 
@@
 data:  x 
-D^+ = 0.04, $p$-value = 0.7731
+D^+ = 0.04, p-value = 0.7731
 alternative hypothesis:
@@
 data:  x 
-D^+ = 0.04, $p$-value = 0.9231
+D^+ = 0.04, p-value = 0.9231
 alternative hypothesis:
@@
 data:  rep(1, 3) 
-D = 0.6667, $p$-value = 0.07407
+D = 0.6667, p-value = 0.07407
 alternative hypothesis: two-sided 
@@
 data:  x 
-W2 = 0.057, $p$-value = 0.8114
+W2 = 0.057, p-value = 0.8114
 alternative hypothesis: Two.sided 
@@
 data:  x 
-A2 = 0.3969, $p$-value = 0.75
+A2 = 0.3969, p-value = 0.75
 alternative hypothesis: Two.sided 
@@
 data:  y 
-U2 = 0.0094, $p$-value = 0.945
+U2 = 0.0094, p-value = 0.945
 alternative hypothesis: Two.sided 
@@
 data:  z 
-U2 = 0.0094, $p$-value = 0.945
+U2 = 0.0094, p-value = 0.945
 alternative hypothesis: Two.sided 

### 2011-08-27 pkg/cvm.test/R/cvm.test.R

@@ -118,3 +118,27 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
   return(RVAL)
 }
 
+
+x <- rbinom(8,10,3/4)
+
+y <- ecdf(rbinom(10000,10,3/4))
+cvm.test(x,y,type='U2')
+x <- 10-x
+y <- ecdf(rbinom(10000,10,1/4))
+cvm.test(x,y, type='U2')
+
+
+x <- rbinom(8,10,3/4)
+
+y <- ecdf(rbinom(10000,10,3/4))
+cvm.test(x,y,type='A2')
+x <- 10-x
+y <- ecdf(rbinom(10000,10,1/4))
+cvm.test(x,y, type='A2')
+
+
+
+
+
+
+

### 2011-08-27 pkg/dgof/R/cvm.test.R

@@ -1,20 +1,6 @@
-cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
-
-  #cvm.stat <- function(x,y,type) {
-  #  fdata <- ecdf(x)
-  #  if(type != 'A2') {
-  #    igrand <- function(z) return((fdata(z) - y(z))^2)
-  #  } else {
-  #    igrand <- function(z) return((fdata(z) - y(z))^2 / (y(z) - y(z)^2))
-  #  }
-  #  stat <- integrate(igrand, -Inf, Inf)[1]
-  #  if(type != 'U2') {
-  #    return(stat)
-  #  } else {
-  #    igrand <- function(z) return((fdata(z) - y(z)-stat)^2) 
-  #    return(integrate(igrand, -Inf, Inf)[1])
-  #  }    
-  #}
+cvm.test <- function(x,y, type=c("W2", "U2", "A2"),
+                    simulate.p.value=FALSE,
+                    B=2000, tol=1e-8) {
 
   cvm.pval.disc <- function(STAT, lambda) {
 
@@ -51,7 +37,8 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
         ans <- ans - 0.5*sum( log(1-2*t*lambda) )
         return(ans)
       }
-      est2 <- exp(nlm(logf, 1/(4*max(lambda)))$minimum)
+      est2 <- 1
+      try( est2 <- exp(nlm(logf, 1/(4*max(lambda)))$minimum), silent=TRUE)
       return(min(est1,est2))
     }
   } # End cvm.pval.disc()
@@ -85,7 +72,7 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
                 U2 = (diag(1, nrow(E)) - 
                       E%*%One%*%t(One))%*%E%*%(diag(1, nrow(E)) -
                       One%*%t(One)%*%E),
-                A2 = E%*%K)
+                      A2 = E%*%K)
     lambda <- eigen(M%*%Sy)$values
 
     STAT <- switch(type, W2 = sum(Z^2*t)/N, U2 = sum((Z-Zbar)^2*t)/N,
@@ -94,6 +81,21 @@ cvm.test <- function(x,y, type=c("W2", "U2", "A2")) {
     return(c(STAT, lambda))
   } # End cvm.stat.disc()
 
+
+  cvm.pval.disc.sim <-  function(STATISTIC, lambda, y, type, tol, B) {
[…]

### 2011-08-27 pkg/dgof/R/ks.test.R

@@ -14,10 +14,10 @@
 #  A copy of the GNU General Public License is available at
 #  http://www.r-project.org/Licenses/
 #
-# Modification by Taylor Arnold and Jay Emerson, 2010:
+#  Modification by Taylor Arnold and Jay Emerson, 2010:
 #
-# Any changes to the original R Core source is noted by
-# the following comment convension to aid code review:
+#  Any changes to the original R Core source is noted by
+#  the following comment convension to aid code review:
 #
 #   ####################
 #   ## TBA, JWE: comment

### 2011-08-27 rjournal/codeforgraphics.R

@@ -112,9 +112,9 @@ plot(I[I > coff & I < coff2], z[I > coff & I < coff2], type='l', xlim=c(coff, co
 abline(h=0, lty=2)
 lines(I[I > 1.3], y[I > 1.3], lty=2, col='steelblue', lwd=3)
 lines(I[I > coff], w[I > coff], lty=2, col='olivedrab4', lwd=3)
-legend(2.25, 0.015, legend=c('Asymptotic p-value (9)',
+legend(2.07, 0.0013, legend=c('Unstable asymptotic p-value from (9)',
                              'Bound given by (11)','Bound given by (13)' ), 
-     col=c('tomato1','steelblue', 'olivedrab4'), lty=c(1,2,2))
+     col=c('tomato1','steelblue', 'olivedrab4'), lty=c(1,2,2), lwd=c(2,3,3))
 #dev.off()
 
 

### 2011-09-01 rjournal/OldRJS/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@75 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-11-29 rjournal/RJtemplate.tex
@@
 
-%\bibliography{example}
+\section{Acknowledgement}
+
+We are grateful to Le-Minh Ho for his implementation of the algorithm
+described in \citet{nieder81} for calculating rectangular probabilities
+of uniform order statistics.
 

### 2011-09-01 rjournal/ArnoldEmerson/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@77 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2011-08-27 rjournal/AE.tex
@@
 
-NOTE: NEED TO ADD SIMUALTED P-VALUES.
-
 \subsection{Overview}
@@
 We resolve this problem by using a combination of
-two simple conservative approximations to
+two conservative approximations to
 avoid the numerical instability. First, consider the
@@
 \begin{align}
-\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{E}
-  \left[\exp\left(t\sum_{i=1}^{p}\lambda_iZ_i^2\right)\right]\exp(-tx) \\
-&= \frac{\exp(-tx)}{\sqrt{\prod_{i=1}^{p}(1-2t\lambda_i)}}
+\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) & \leq \nonumber \\
+\mathbb{E}\biggl[\exp\biggl( & t\sum_{i=1}^{p} 
+    \lambda_iZ_i^2\biggr)\biggr] \exp(-tx) \\
+& = \frac{\exp(-tx)}{\sqrt{\prod_{i=1}^{p}(1-2t\lambda_i)}}
 \label{ineq3}
@@
 are both calculated and the smaller is used in cases where the numerical
-instability of (\ref{distQ}) is a concern.
+instability of (\ref{distQ}) may be a concern.
 
@@
 \begin{Sinput}
-> library(ks.test)
+> library(dgof)
 > dgof::ks.test(c(0, 1), ecdf(c(0, 1)))
@@
 \begin{Sinput}
-> library(cvm.test)
 > cvm.test(x, ecdf(1:10))
@@
 \begin{Sinput}
-> library(cvm.test)
 > set.seed(1)
